### 2011-01-05 proof.tex
Annotated proof of output_input_at_update_times
Includes new technology for annotating proofs.
Collects proof states along with string descriptions using
side-effecting tactics that don't change the proof state.
The presentation side doesn't work too well at the moment, mostly
because LaTeX and the pretty-printer don't get along...

Also reorder simplifiedTheory so everything before
output_input_at_update_times is a dependency of it.

\documentclass{article}
\usepackage{holtexbasic}
\usepackage{alltt}
\usepackage{fixltx2e}
\begin{document}
\input{output-input-at-update-times.tex}
\end{document}

### 2011-01-05 proof.tex
Use smaller margins and no paragraph indents
Start annotating proofs for more theorems
Add a write_proof function that writes the last proof annotated, adding
extra rules to the grammars beforehand and removing them afterwards.
Make the proof tex files end with Proof.tex and gitignore them
Add introduction, model summary, and sectioning to the proof document

@@
 \usepackage{alltt}
-\usepackage{fixltx2e}
+\usepackage[a4paper,margin=2cm]{geometry}
+\usepackage{parskip}
 \begin{document}
-\input{output-input-at-update-times.tex}
+\section*{Introduction}
+First, a warning about one quirky convention.
+When I write $\sum_{m=0}^b (f m)$ I mean for the variable $m$ to range from $0$ up to but \emph{not including} $b$.
+
+This file contains the statements of a few key lemmas, as well as proof sketches for the important lemmas.
+How to read the proofs?
+Imagine you are trying to prove the statement yourself.
+There are snapshots of the ``proof state'' interspersed with commentary on what to do next to get to the next proof state.
+A proof state consists of what is currently remaining to be proved, then a line of dashes (\texttt{----------}) then a numbered list of assumptions that are currently being assumed.
+(Note: sometimes we might be proving a subsidiary goal, and the proof state of the whole goal isn't printed until that subgoal is finished; I try to make this clear with the commentary.)
+
+\section*{The Model}
+\begin{description}
+\item[Global Parameters:]\hfill
+\begin{itemize}
+\item
+     the width of every slice, encoded as the index, w, of the last shift register
+\item
+     the input series, input.
+\end{itemize}
+\item[Each slice has:]\hfill
+\begin{itemize}
+\item
+     w+1 shift registers, named SR 0, SR 1, ..., SR w;
+\item
+     an output register.
+\end{itemize}
+\item[Each SR has a source]\hfill
+\begin{itemize}
+\item
+     The source for SR 0 of slice 0 is the global input.
+\item
+     The source for SR 0 of slice (SUC n) is the output of slice n.
+\item
+     The source for SR (SUC m) is SR m of the same slice.
+\end{itemize}
+\item[Each slice updates periodically]\hfill
+\begin{itemize}
+\item
+     Slice n updates at time t iff ∃x. n + ((SUC x) * ((SUC w) ** n))
+\item
+   At time 0, all shift registers have value 0.
+\item
+   At each update time for its slice, a SR gets the value of its source at the previous time.
+\item
+   The output of a slice at some time is the sum of its shift registers at the same time.
+\end{itemize}
+\end{description}
+\section*{Some Small Lemmas}
+\input{SRLastUpdateProof.tex}
+\input{SRFirstProof.tex}
+\input{SR0UntilProof.tex}
+\section*{Major Lemma 1}
+\input{OutputFirstProof.tex}
+\section*{Major Lemma 2}
+\input{OutputSourceProof.tex}
+\section*{Main Theorem}
+\input{SumInChunksProof.tex}
+\input{OutputInputProof.tex}
 \end{document}

### 2011-01-05 proof.pre.tex
Start using the munger; make overrides a separate file
Use munger to improve the model description

--- 2011-01-05 proof.tex
@@
 First, a warning about one quirky convention.
-When I write $\sum_{m=0}^b (f m)$ I mean for the variable $m$ to range from $0$ up to but \emph{not including} $b$.
+When I write \HOLtm{SIGMA (\m. f m) (count b)} I mean for the variable $m$ to range from $0$ up to but \emph{not including} $b$.
 
@@
 \item
-     the width of every slice, encoded as the index, w, of the last shift register
+     the width of every slice, encoded as the index, \HOLtm{p.w}, of the last shift register;\\
+     the total number of shift registers is \HOLtm{SUC p.w = p.w + 1};
 \item
-     the input series, input.
+     the input series, \HOLtm{p.input}.
 \end{itemize}
@@
 \item
-     w+1 shift registers, named SR 0, SR 1, ..., SR w;
+     \HOLtm{SUC p.w} shift registers, named \HOLtm{SR p n 0}, \HOLtm{SR p n 1}, ..., \HOLtm{SR p n p.w};
 \item
-     an output register.
+     an output register \HOLtm{output n}.
 \end{itemize}
@@
 \item
-     The source for SR 0 of slice 0 is the global input.
+     The source for \HOLtm{SR p 0 0} of slice \HOLtm{0} is the global input.
 \item
-     The source for SR 0 of slice (SUC n) is the output of slice n.
+     The source for \HOLtm{SR p (SUC n) 0} of slice \HOLtm{SUC n} (the output of slice \HOLtm{n}).
 \item
-     The source for SR (SUC m) is SR m of the same slice.
+     The source for \HOLtm{SR p (SUC n) (SUC m)} is \HOLtm{SR p (SUC n ) m} (of the same slice).
 \end{itemize}
@@
 \item
-     Slice n updates at time t iff ∃x. n + ((SUC x) * ((SUC w) ** n))
+     Slice \HOLtm{n} updates at time \HOLtm{t} if and only if \HOLtm{∃x. t = n + ((SUC x) * ((SUC p.w) ** n))}
 \item
-   At time 0, all shift registers have value 0.
+   At time \HOLtm{0}, all shift registers have value \HOLtm{0}.
 \item
@@
 \end{description}
+
+This is all summarized by the folowing theorems.
+\begin{alltt}
+\HOLthm[alltt]{simplified.update_time_def}
+\HOLthm[alltt]{simplified.Slice_def}
+\end{alltt}
 \section*{Some Small Lemmas}

### 2011-01-06 proof.pre.tex
Make SIGMA TeX notation use an inclusive upper bound

@@
 \section*{Introduction}
-First, a warning about one quirky convention.
-When I write \HOLtm{SIGMA (\m. f m) (count b)} I mean for the variable $m$ to range from $0$ up to but \emph{not including} $b$.
+Update: Contrary to my earlier warning when I write \HOLtm{SIGMA (\m.
+f m) (count (SUC b))} I now mean for the variable $m$ to range from $0$ up to \emph{and including} $b$, as per the usual convention.
 

### 2011-01-07 proof.pre.tex
Delegate formatting details to TeX commands

@@
 \usepackage{alltt}
+\newcommand{\HOLAsmSep}{\texttt{----------}}
 \usepackage[a4paper,margin=2cm]{geometry}
@@
 There are snapshots of the ``proof state'' interspersed with commentary on what to do next to get to the next proof state.
-A proof state consists of what is currently remaining to be proved, then a line of dashes (\texttt{----------}) then a numbered list of assumptions that are currently being assumed.
+A proof state consists of what is currently remaining to be proved, then a line of dashes (\HOLAsmSep{}) then a numbered list of assumptions that are currently being assumed.
 (Note: sometimes we might be proving a subsidiary goal, and the proof state of the whole goal isn't printed until that subgoal is finished; I try to make this clear with the commentary.)

### 2011-03-18 fmics/paper.tex
Add FMICS paper to repo

\documentclass{llncs}

\usepackage{times,graphicx,epsfig,boxedminipage,url,xspace,array}

\begin{document}

\thispagestyle{empty}
\title{Formal Verification of Real-Time Data Processing of the LHC Beam Loss Monitoring System: A Case Study}
\author{Naghmeh Ghafari \inst{1} \and Ramana Kumar \inst{2} \and Jeff Joyce \inst{1}  
\and  Bernd Dehning \inst{3} \and   Christos Zamantzas \inst{3} }
\institute{Critical Systems Labs, Vancouver, BC, Canada
  \and University of Cambridge, Cambridge, UK \and CERN, Geneva, Switzerland}

\pagestyle{plain}

\maketitle

\begin{abstract}

This paper describes a collaborative effort in which the HOL4 theorem prover is being used to formally verify properties of a complex structure used in the FGPA implementation of an integral element of the Large Hadron Collider (LHC) machine protection system at the European Organization for Nuclear Research (CERN). This structure, known as the \emph{Successive Running Sums} (SRS), generates the primary input to the decision logic that must initiate a critical action by the LHC machine protection system in response to the detection of a dangerous level of beam particle loss. The use of mechanized logical deduction complements an intensive study of the SRS structure using simulation. We are especially interested in using logical deduction to obtain a generic result that will be applicable to variants of the SRS structure. This collaborative effort has individuals with very diverse backgrounds ranging from theoretical physics to system safety. One interesting result is the extent to which the use of a formal method has compelled the stakeholders to clarify intricate details of the SRS structure and behaviour.

%The strategy for machine protection of the Large Hadron Collider (LHC) at the European Organization for Nuclear Research (CERN) is mainly based on the Beam Loss Monitoring (BLM) system. At each turn of the beam, several thousands of data is recorded and processed in order to decide whether the beam should be permitted to continue circulating or it should be extracted due to a dangerous beam loss that may result in significant equipment damage.Such a decision is based on the analysis of loss patterns in time and the beam energy.  The real-time analysis of the recorded data is mainly performed by the \emph{Successive Running Sums} (SRS) technique which allows many and long integration periods to be maintained for each detector's data.

%In this case study, our eventual goal is to use HOL4 theorem prover to show the correctness of the SRS technique in with respect to BLM specification.

%we use HOL4 theorem prover to show the correctness of the SRS technique in BLM. We present the formulation of a parameterized model of SRS in HOL4 and explain the verification strategy and results. {\bf what did we learn through this exercise from verification perspective?}

\end{abstract}

%\sloppy

\section{Introduction}

The Large Hadron Collider (LHC) at the European Organization for Nuclear Research (CERN) provides a head-on collisions of protons at a center of a mass energy of 14 TeV for high energy particle physics research. %The LHC is installed in a 27 km long LEP tunnel. 
In order to reach the required magnetic field strengths, superconducting magnets cooled with superfluid helium are used. In addition to the energy stored in the LHC, the 700 MJ stored in the circulating beams could initiate quenches (transition of the superconductors to normal conducting state) and cause damages if only a very small fraction of the circulating beam particles deposit their energy in the equipment. The consequences of such event ranges from several hours of down time for cooling down the magnets to reach the superconducting state again, to months of repairs in the case of equipment damage.

The strategy for machine protection of the LHC is mainly based on the Beam Loss Monitoring (BLM) system. At each turn of the beam, BLM records several thousands of data and process them in order to decide to signal the safe extraction of the beam if losses exceed quench level thresholds. The processing involves a proper analysis of the loss pattern in time and the energy of the beam.

{\bf myNote: Add 1) the approach currently taken for the analysis of SRS, 2) the approach we took in this paper for the analysis, 3) the reasons behind that  the goals for this case study 4) the main contributions}

 Rushby's emphasis on the role of mechanized logical deduction as an "implacable skeptic that insists on its human user providing justification" thus rings true.
 %Rushby's description of mechanized logical deduction as an "implacable skeptic that insists on its human user providing justification" rings true for this effort.

The your process of trying to
capture the specification via experiment and refinement by a team with
different backgrounds and expertise is very interesting. the eventual goal is to use HOL4 theorem prover to show
the correctness of the SRS technique in BLM

%The loss of a very small fraction of the circulating beam may induce a quench of the super conducting magnets or even physical damage to machine components. The detection of the lost beam protons allows protection of the equipment against quenches and damage by generating a beam dump trigger when the losses exceed thresholds. At each turn of the beam, there will be several thousands of data to record and process in order to decide if the beams should be permitted to continue circulating or their safe extraction is necessary to be triggered.

%In general, the BLM system converts the signal given by the recording of shower particles into electrical signals and decides if threshold limits are being exceeded. The beam-permit signal is de-asserted if one detector exceeds the limit and consecutively is transferred to the beam dump system  by the Beam Interlock system.

\section{BLM System Overview}

\begin{figure}[t]
  \centering  \scalebox{0.7}{ \includegraphics{BLETC.eps}}
   \caption{Block diagram of the processes assigned to BLETC card.}
  \label{fig:BLETC}
\end{figure}

The loss measurement principle \cite{Dehning-IPAC,Chris-FPGA} is based on the energy deposition detection of secondary shower particles using a specially-designed detector called \emph{ionization chamber}. There are around 4000 ionization chambers located under the magnets all around the tunnel. These detectors convert the signal given by the recording of shower particles into electrical signals and send them to tunnel cards. Tunnel cards, which are located in the tunnel and implemented by radiation-tolerant electronics, acquire and digitize the data from the detectors and transmit those to the surface using optical links. There, the data processing cards, named BLETCs, receive those data and decide whether or not the beam should be permitted to be injected or continue circulating. Each tunnel card receives data from eight detectors and each surface card receives data from two tunnel cards. The surface card provides data to the Logging, the Post Mortem and the Collimation systems that will be used to drive on-line displays in the control room to allow offline analysis of the losses and to setup automatically the collimators. Due to demanding performance requirements, the BLM system is making use of modern field programmable gate arrays (FPGAs), which include the resources needed to design complex processing and can be reprogrammed making them ideal for future upgrades or system specification changes. Figure~\ref{fig:BLETC} shows a block diagram of the processes assigned in BLETC FPGA. In the following, we give a brief overview of each of the four main processing blocks in BLETC card.

\emph{(a) Receive, Check, and Compare (RCC)}: RCC is responsible for facilitating the reception process at the entry stage of the surface FPGA. It ensures the correct reception and detection of erroneous transmissions by redundancy in transmissions and using Cyclic Redundancy Check (CRC) and the 8B/10B algorithms.

\emph{(b) Data Processing}: Quench of magnets that result from proton loss depends on the loss duration and the beam energy. Given the tolerance acceptable for quench prevention, the quench threshold versus loss time is approximated with a minimum number of sliding integration windows (called \emph{Running Sums}) fulfilling the tolerance. In order to achieve the \emph{dynamic range} (i.e. the domain of variation of the beam losses) requested by the specification, the detectors use both Current-to-Frequency Converters (CFCs) and ADC circuitries. The data processing block merges these two types of data coming from the same detector into one value and send it to Successive Running Sums (SRS) block. The verification of SRS block is the focus of this paper and its implementation is discussed in more detail in the next section.

\emph{(c) Threshold Comparator}: Every new calculated running sum need to be compared with the corresponding threshold that was chosen by the beam energy reading given that moment. The comparator will initiate a beam dump request if any of the running sums is higher than its corresponding threshold. The beam dump requests are forwarded to the Beam Interlock System which will initiate the beam dump. There are 12 running sums calculated for each 16 detector channels allocated to a BLETC card. The beam energy information is scaled into 32 levels (0.45 to 7 TeV) and each processing module will hold data only for those 16 detectors connected. That would give a total of 6,144 threshold values needed to be held on each card.

\emph{(d) Logging, Post Mortem and Collimation}:  To be able to trace back the loss signal development, BLM should store the loss measurement data. This data will be sent over the VME-bus for online viewing and storage by the Logging and Post-Mortem systems. For the purpose of supervision, the BLM system will drive an online event display to show error and status information recorded by tunnel electronics and the RCC process as well as the maximum loss rates seen by the running sums. BLETC card also provides data to Collimation system for the correct alignment and setup of the collimators.

\section{Successive Running Sums (SRS)}

\begin{figure}[t]
  \centering \includegraphics{rs.eps}
   \caption{Block diagram showing how to produce and maintain a continuous running sum of arriving values.}
  \label{fig:RS-basic}
\end{figure}

Beam losses can happen either in a single turn of the beam in the tunnel, with a sudden beam loss, or in progressive losses during numerous turns. One-turn failures are called \emph{ultra-fast} losses. Multi-turn losses can be divided between \emph{very fast} losses those which happen in less than 10 ms, \emph{fast} losses which happen in more than 10 ms and \emph{steady} losses, where the beam is lost over one second or more~\cite{Schmidt-ICFA}.

The processing of the data collected by the detectors involves a proper analysis of the loss patterns in time and a proper account of the energy of the beam. The procedure for the data processing is based on the idea that a constantly updated moving window can be kept by adding to a register the incoming newest value and subtracting its oldest value (see Figure~\ref{fig:RS-basic}). The number of values that are kept under the window defines the integration time it represents.  The ideal strategy is to have an infinite number of constantly updated moving windows with various lengths to cover the whole time region from 40 micro-seconds (the data arrival rate to BLETC card)  to 100 seconds for detecting ultra-fast losses to steady losses, respectively. Such an implementation is not feasible due to the limited amount of resources. Therefore, given the tolerance acceptable for quench prevention, the quench threshold versus loss time curve is approximated by a minimum number of steps that fulfill this tolerance.

\begin{figure}[t]
  \centering \includegraphics{SRS-basic.eps}
   \caption{Block diagram showing a configuration for efficient summation of many values.}
  \label{fig:SRS-basic}
\end{figure}

In order to construct long moving windows, it is required to keep long histories of received data. The technique employed to reach long integration periods with relatively small in length shift registers uses consecutive storage of sums of the received values. Instead of storing all values needed for the sum, this technique stores successively parts of the total sum using only a fraction of the otherwise needed memory space. This technique works by feeding the sum of one shift register's contents, every time its contents become completely updated, to the input of another shift register (see Figure~\ref{fig:SRS-basic}). By cascading more of these elements, very long moving windows could be constructed with significantly small memory space. The above scheme is the basis for \emph{Successive Running Sums} (SRS) implementation in BLETC.

SRS implementation minimizes the resource usage by using the already calculated running sums in order to calculate bigger in length running sums without the need of extra summation points. In addition, it makes use of multipoint shift registers that are configured to give intermediate outputs, referred to as \emph{taps}. The taps provide data outputs at certain points in the shift register chain thus contribute towards the efficient use of resources.

As in the SRS implementation the sum of one shift register's content is fed to the input of another shift register, the optimal achievable latency in the response of each stage is equal to the refreshing time of the preceding shift registers, i.e. the time needed to completely update its contents.  The \emph{read delay} signal (see Figure~\ref{fig:SRS-basic}) of each shift register holds a delay equal to this latency to guarantee the correct operation. Thus, the delay is every time equal to the preceding shift register's delay multiplied by the elements planned to be used in the sum.

\begin{figure}[t]
  \centering \scalebox{0.65}{ \includegraphics{srs-modified.eps}}
   \caption{The block diagram showing the Successive Running Sums implementation in BLETC card.}
  \label{fig:srs}
\end{figure}

Figure~\ref{fig:srs} shows the implementation of SRS in BLETC card. It consists of 6 \emph{slices}, where each slice computes two running sums (noted by RS) with the use of a multipoint shift register, two subtractors and two accumulators (see Figure~\ref{fig:slice}).

As shown in Table~\ref{}, by cascading 6 slices, it is enough to reach the 100 seconds upper integration limit requested by the specifications.

\begin{figure}[t]
  \centering \scalebox{0.8}{ \includegraphics{multi-rs.eps}}
   \caption{The block diagram showing the implementation of each slice.}
  \label{fig:slice}
\end{figure}

\begin{figure}[t]
\centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multicolumn{2}{|c}{Range} &\multicolumn{2}{|c|}{Refreshing} & & \\\cline{1-4}
40 us steps & ms & 40 us steps  & ms & slice No & running sum \\\hline\hline
1 & 0.04 & 1 & 0.04 & & RS0 \\\hline
2 & 0.08 & 1 & 0.04 & & RS1 \\\hline
8 & 0.32 & 1 & 0.04 & slice 1 & RS2 \\\hline
16 & 0.64 & 1 & 0.04 & & RS3 \\\hline
64 & 2.56 & 2 & 0.08 & slice 2 & RS4 \\\hline
256 & 10.24 & 2 & 0.08 & & RS5 \\\hline
2048 & 81.92 & 64 & 2.56 &slice  3 & RS6 \\\hline
16384 & 655.36 & 64 & 2.56 &slice 3 & RS7 \\\hline
32768 &1310.72 & 2048 & 81.92 & slice 4 & RS8 \\\hline
131072 & 5242.88 & 2048 & 81.92 & slice 4 & RS9 \\\hline
524288 & 2097.52 & 16384 & 655.36 & slice 5 &RS10 \\\hline
2097152 & 83886.08 & 16384 & 655.36 & slice 5 & RS11 \\\hline

\end{tabular}
\vspace{0.04in}
\label{fig:example-bakery}
\caption{}
\end{figure}

%A similar configuration for the production of the running sums is used in the BLMTC where in order to increase the efficiency in resources, a set of parameterized multipoint shift registers are used. In this configuration, called \emph{Successive Running Sums}, each incoming new value is delayed with a fixed number of cycles by passing it through a shift register and adding the difference of the new and the outputted from the shift register to an accumulator (see Figure ).

\section{Verification of SRS Implementation using HOL}

Calculation of the running sums is a critical component of BLM. In case of no dangerous loss, BLM should not inhabit the beam. If it fails, it generates a false alarm and three hours will be lost to recover the previous situation. Such false alarms compromise the availability of the system. Moreover, in case of a dangerous loss, BLM it has to inhibit the beam permit. If it fails there will be at least 30 days of downtime. 
Any error in this computation would either compromise the availability of the system (unnecessary request for beam dump) or safety of the system (nor requesting a necessary beam dump). The current practice to check the correctness of the SRS is by simulation and testing in which (explanation of how such a test gets performed).

In this paper we present a formal verification approach for the analysis of SRS implementation in BLM.

The formal verification effort described in this paper is based on
 the use of mechanised logical deduction, or theorem proving, to
 demonstrate that desired properties of a system are logically implied
 by a formal model of the system.  Our approach to theorem-proving
 uses the HOL4 open source software tool, which was developed
 initially at the University of Cambridge, but now by an international
 team~\cite{HOL4}.  HOL4 partially automates the task of proving theorems of
 higher-order logic [ref], a formal logic with a similar expressive
 power to set theory that is widely used for formalising hardware and
 software models and statements about them.  Typically, a very large
 portion of the logical deduction is performed automatically by the
 HOL4 system leaving the user to provide the high level proof
 strategy.  The implementation of the HOL4 system is based on Milner's
 LCF approach [ref] and consists of a small “kernel” that provides a
 very high level of confidence in the logical soundness of the
 verification results obtained using the system. The work described
 here could have been done using other systems such as HOL Light
 [ref], Isabelle/HOL [ref], ProofPower [ref], PVS [ref] or Coq
 [ref]. The first three of these support essentially the same version
 of higher-order logic as HOL4, whilst PVS and Coq support more
 powerful logics. <What about ACL2?>

 While offering less “push button” automation than many other kinds of
 formal verification such as model-checking, machine-assisted theorem
 proving is the most appropriate approach to verifying the BLM SRS,
 since it is unlikely that the correctness theorems, which need lemmas
 proved by manually guided mathematical induction, could be generated
 automatically.

The main objective of formalizing the functionality of SRS was to keep the topology parameterized. It is very probable that in the future the configuration of the slices and shift registers changes to accommodate higher precision. If the model is parameterized then, the same model can be used in the future uses of the system. For this, we formalized the definitions of the model by six parameters:

\section{Conclusions and Lessons Learned}

\bibliographystyle{plain} 
\bibliography{refs}

\end{document}

### 2011-03-19 fmics/paper.tex
remove/add some spaces in the tex file
Rewrite the abstract in simpler language

@@
 \institute{Critical Systems Labs, Vancouver, BC, Canada
-  \and University of Cambridge, Cambridge, UK \and CERN, Geneva, Switzerland}
+  \and University of Cambridge, Cambridge, UK
+  \and CERN, Geneva, Switzerland}
 
@@
 
-This paper describes a collaborative effort in which the HOL4 theorem prover is being used to formally verify properties of a complex structure used in the FGPA implementation of an integral element of the Large Hadron Collider (LHC) machine protection system at the European Organization for Nuclear Research (CERN).
-This structure, known as the \emph{Successive Running Sums} (SRS), generates the primary input to the decision logic that must initiate a critical action by the LHC machine protection system in response to the detection of a dangerous level of beam particle loss.
+We describe a collaborative effort in which the HOL4 theorem prover is being used to formally verify properties of a structure within the Large Hadron Collider (LHC) machine protection system at the European Organization for Nuclear Research (CERN).
+The structure, known as \emph{Successive Running Sums} (SRS), provides input to the decision logic for activating the LHC machine protection system in response to the detection of a dangerous level of beam particle loss.
 The use of mechanized logical deduction complements an intensive study of the SRS structure using simulation.
 We are especially interested in using logical deduction to obtain a generic result that will be applicable to variants of the SRS structure.
-This collaborative effort has individuals with very diverse backgrounds ranging from theoretical physics to system safety.
+This collaborative effort has individuals with diverse backgrounds ranging from theoretical physics to system safety.
 One interesting result is the extent to which the use of a formal method has compelled the stakeholders to clarify intricate details of the SRS structure and behaviour.
 
-%The strategy for machine protection of the Large Hadron Collider (LHC) at the European Organization for Nuclear Research (CERN) is mainly based on the Beam Loss Monitoring (BLM) system.
-At each turn of the beam, several thousands of data is recorded and processed in order to decide whether the beam should be permitted to continue circulating or it should be extracted due to a dangerous beam loss that may result in significant equipment damage.Such a decision is based on the analysis of loss patterns in time and the beam energy.
-The real-time analysis of the recorded data is mainly performed by the \emph{Successive Running Sums} (SRS) technique which allows many and long integration periods to be maintained for each detector's data.
-
-%In this case study, our eventual goal is to use HOL4 theorem prover to show the correctness of the SRS technique in with respect to BLM specification.
-
-%we use HOL4 theorem prover to show the correctness of the SRS technique in BLM.
-We present the formulation of a parameterized model of SRS in HOL4 and explain the verification strategy and results.
-{\bf what did we learn through this exercise from verification perspective?}
-
 \end{abstract}
-
-%\sloppy
 

### 2011-03-19 fmics/paper.tex
Reword first two paragraphs of introduction.

@@
 
-The Large Hadron Collider (LHC) at the European Organization for Nuclear Research (CERN) provides a head-on collisions of protons at a center of a mass energy of 14 TeV for high energy particle physics research.
-%The LHC is installed in a 27 km long LEP tunnel.
-In order to reach the required magnetic field strengths, superconducting magnets cooled with superfluid helium are used.
-In addition to the energy stored in the LHC, the 700 MJ stored in the circulating beams could initiate quenches (transition of the superconductors to normal conducting state) and cause damages if only a very small fraction of the circulating beam particles deposit their energy in the equipment.
-The consequences of such event ranges from several hours of down time for cooling down the magnets to reach the superconducting state again, to months of repairs in the case of equipment damage.
+The Large Hadron Collider (LHC) at the European Organization for Nuclear Research (CERN) is a high-energy particle accelerator.
+It collides beams of protons, where each particle has an energy of 14 TeV, and the results of these collisions are used for high-energy physics research.
+In order to reach the required magnetic field strengths, the LHC has superconducting magnets cooled with superfluid helium.
+Because the energy of the beams is so high (700 MJ), if even a small fraction of the beam particles leak into the equipment they can cause the superconductors to transition to their normal conducting state.
+Such a transition is called a quench.
+The consequences of a quench range from several hours of down time (to cool the magnets back down to their superconducting state), to months of repairs (in the case of equipment damage).
 
-The strategy for machine protection of the LHC is mainly based on the Beam Loss Monitoring (BLM) system.
-At each turn of the beam, BLM records several thousands of data and process them in order to decide to signal the safe extraction of the beam if losses exceed quench level thresholds.
-The processing involves a proper analysis of the loss pattern in time and the energy of the beam.
+The main strategy for protecting the LHC from quenches is the Beam Loss Monitoring (BLM) system, which stops the beams safely if the particle loss exceeds thresholds that are likely to result in a quench.
+At each cycle of the beams around the LHC, the BLM system records and processes thousands of data points in order to decide whether to stop the beams.
+The processing includes analysis of the loss pattern over time and of the energy of the beam.
 

### 2011-03-19 fmics/paper.tex
Add another paragraph to the introduction.

@@
 
-The main strategy for protecting the LHC from quenches is the Beam Loss Monitoring (BLM) system, which stops the beams safely if the particle loss exceeds thresholds that are likely to result in a quench.
-At each cycle of the beams around the LHC, the BLM system records and processes thousands of data points in order to decide whether to stop the beams.
+The main strategy for protecting the LHC from quenches is the Beam Loss Monitoring System (BLMS), which stops the beams safely if the particle loss exceeds thresholds that are likely to result in a quench.
+At each cycle of the beams around the LHC, the BLMS records and processes thousands of data points in order to decide whether to stop the beams.
 The processing includes analysis of the loss pattern over time and of the energy of the beam.
 
+The BLMS must respond to dangerous losses quickly, but determining whether losses are dangerous may require analysis of the data recorded over a long period. 
+Furthermore, the BLMS must continue recording large amounts of data in real time while processing.
+To achieve these goals, the BLMS maintains approximate cumulative sums of particle losses over a variety of sizes of moving windows.
+The component responsible for maintaining these sums is called the Successive Running Sums (SRS) structure.
+The SRS is implemented in hardware, in order to be fast enough to work in real time, and on FPGAs in particular so that they can be easily reprogrammed with future designs if necessary.
+Multiple implementations of the SRS structure are installed regularly around the LHC as part of the BLMS.
+<this paragraph probably needs some references>
+
+<SPECIFICATION/REQUIREMENTS OF SRS: it is approximate sums of loss counts, but they need to be relatively close to the exact sums>
+<SIMULATION APPROACH TO ANALYSIS OF SRS>
+<FORMAL APPROACH>
 {\bf myNote: Add 1) the approach currently taken for the analysis of SRS, 2) the approach we took in this paper for the analysis, 3) the reasons behind that  the goals for this case study 4) the main contributions}

### 2011-03-19 fmics/paper.tex
paragraph in intro about why srs sum needs to be close to exact sum

@@
 
-<SPECIFICATION/REQUIREMENTS OF SRS: it is approximate sums of loss counts, but they need to be relatively close to the exact sums>
+The BLMS detects beam loss in the form of counts proportional to the number of stray particles.
+Dangerous loss levels can be expressed as thresholds on the rate of counts over time.
+The SRS maintains approximate sums of counts over time windows of various sizes, from which loss rates can be calculated.
+It is preferable for the SRS sum to be above the exact sum over a particular time window, so that dangerous losses are always detected.
+But stopping the beam when the loss level is not dangerous is time consuming and costly.
+Therefore, the discrepancy between the SRS sum and the exact sum needs to be small.
+
 <SIMULATION APPROACH TO ANALYSIS OF SRS>

### 2011-03-19 fmics/paper.tex
Add HOL4 and Isabelle bibliography entries

@@
  here could have been done using other systems such as HOL Light
- [ref], Isabelle/HOL [ref], ProofPower [ref], PVS [ref] or Coq
+ [ref], Isabelle/HOL~\cite{Isabelle}, ProofPower [ref], PVS [ref] or Coq
  [ref]. The first three of these support essentially the same version

### 2011-03-19 fmics/paper.tex
Start drafting another couple of paragraphs in intro

@@
 To achieve these goals, the BLMS maintains approximate cumulative sums of particle losses over a variety of sizes of moving windows.
-The component responsible for maintaining these sums is called the Successive Running Sums (SRS) structure.
+The component responsible for maintaining these sums is called Successive Running Sums (SRS).
 The SRS is implemented in hardware, in order to be fast enough to work in real time, and on FPGAs in particular so that they can be easily reprogrammed with future designs if necessary.
@@
 
-<SIMULATION APPROACH TO ANALYSIS OF SRS>
-<FORMAL APPROACH>
+The SRS component is a complex structure, but the correctness of its behaviour is important for safe and productive use of the LHC.
+Simulation of the SRS on sample streams of input counts is one approach, currently in use [ref?], to building an understanding of the SRS's behaviour and confidence that the behaviour is correct. 
+<details about simulation>
+
+Formal methods are another approach to building understanding and confidence.
+Compared to test-based methods, like simulation, formal methods offer a much higher confidence of correctness, but require precise specifications and proof effort.
+The work on constructing proofs may be compared to the work on analysing the results of simulations.
+The work on formulating precise specifications is extra, but is arguably beneficial for both methods.
+<need to be more descriptive here i.e. to do number 2) below rather than 3)>
+
 {\bf myNote: Add 1) the approach currently taken for the analysis of SRS, 2) the approach we took in this paper for the analysis, 3) the reasons behind that  the goals for this case study 4) the main contributions}

### 2011-03-22 fmics/paper.tex
merge Dropbox changes to paper.tex, with some modifications

--- 2011-03-19 fmics/paper.tex
@@
 We describe a collaborative effort in which the HOL4 theorem prover is being used to formally verify properties of a structure within the Large Hadron Collider (LHC) machine protection system at the European Organization for Nuclear Research (CERN).
-The structure, known as \emph{Successive Running Sums} (SRS), provides input to the decision logic for activating the LHC machine protection system in response to the detection of a dangerous level of beam particle loss.
+This structure, known as the \emph{Successive Running Sums} (SRS), generates the primary input to the decision logic that must initiate a critical action by the LHC machine protection system in response to the detection of a dangerous level of beam particle loss.
 The use of mechanized logical deduction complements an intensive study of the SRS structure using simulation.
@@
 The Large Hadron Collider (LHC) at the European Organization for Nuclear Research (CERN) is a high-energy particle accelerator.
-It collides beams of protons, where each particle has an energy of 14 TeV, and the results of these collisions are used for high-energy physics research.
+It provides a head-on collisions of protons at a center of a mass energy of 14 TeV for high-energy particle physics research.
 In order to reach the required magnetic field strengths, the LHC has superconducting magnets cooled with superfluid helium.
-Because the energy of the beams is so high (700 MJ), if even a small fraction of the beam particles leak into the equipment they can cause the superconductors to transition to their normal conducting state.
-Such a transition is called a quench.
-The consequences of a quench range from several hours of down time (to cool the magnets back down to their superconducting state), to months of repairs (in the case of equipment damage).
+Due to a very high energy stored in the circulating beams (700 MJ), if even a small fraction of the beam particles deposit their energy in the equipment, they can cause the superconductors to transition to normal conducting state.
+Such a transition is called a \emph{quench}.
+The consequences of a quench range from several hours of down time (to cool down the magnets to reach the superconducting state), to months of repairs (in the case of equipment damage).
 
-The main strategy for protecting the LHC from quenches is the Beam Loss Monitoring System (BLMS), which stops the beams safely if the particle loss exceeds thresholds that are likely to result in a quench.
-At each cycle of the beams around the LHC, the BLMS records and processes thousands of data points in order to decide whether to stop the beams.
+The strategy for machine protection of the LHC is mainly based on the Beam Loss Monitoring System (BLMS), which triggers the safe extraction of the beam if the particle loss exceeds thresholds that are likely to result in a quench.
+At each cycle of the beams around the LHC, the BLMS records and processes several thousands of data points in order to decide if the beams should be permitted to continue circulating or their safe extraction is necessary to be triggered.
 The processing includes analysis of the loss pattern over time and of the energy of the beam.
 
-The BLMS must respond to dangerous losses quickly, but determining whether losses are dangerous may require analysis of the data recorded over a long period. 
-Furthermore, the BLMS must continue recording large amounts of data in real time while processing.
+The BLMS must respond to dangerous losses quickly, but determining whether losses are dangerous may require analysis of the loss data recorded over long periods of time.
+Furthermore, the BLMS must continue recording large amounts of data in real-time while processing.
 To achieve these goals, the BLMS maintains approximate cumulative sums of particle losses over a variety of sizes of moving windows.
-The component responsible for maintaining these sums is called the Successive Running Sums (SRS) structure.
-The SRS is implemented in hardware, in order to be fast enough to work in real time, and on FPGAs in particular so that they can be easily reprogrammed with future designs if necessary.
-Multiple implementations of the SRS structure are installed regularly around the LHC as part of the BLMS.
-<this paragraph probably needs some references>
+The component responsible for maintaining these sums is called \emph{Successive Running Sums} (SRS).
+The SRS is implemented in hardware, in order to be fast enough to work in real-time, and on Field Programmable Gate Arrays (FPGAs) in particular so that they can be easily reprogrammed with future upgrades~\cite{Chris-FPGA}.
+%Multiple implementations of the SRS structure are installed regularly around the LHC as part of the BLMS.
 
-The BLMS detects beam loss in the form of counts proportional to the number of stray particles.
+The SRS component has a complex structure and the correctness of its behaviour is critical for safe and productive use of the LHC.
+Any error in SRS implementation would either compromise the availability of the LHC (unnecessary request for beam dump) or safety (not triggering a necessary beam dump).
+The current approach for analysis of SRS is mainly  based on simulation of its behavior on sample streams of input~\cite{}.
+(one more sentence needs to be added on the current verification process)
+
+In this paper, we describe a formal verification approach based on logical deduction for the analysis of SRS.
+(need to add few sentences explaining in a very high level the verification approach/strategy)
+
+Compared to test-based methods, like simulation, formal methods not only offer a much higher confidence in correctness of the systems' behavior, but also help in better understanding their specification.
+One of the main challenges in implementing a formal verification approach for SRS was to capture the intricate details of the system's specification via experiment and refinement with a team with different backgrounds and expertise.
+Our confidence in the SRS design as a result of this effort ultimately rests upon our deep understanding of why the design is correct rather than the fact that we obtained ``Theorem Proved" as the final output of a software tool at the end of a long and intense effort.
+In particular, our use of mechanized logical deduction was a highly iterative process that incrementally refined our understanding of (1) the implementation (2) the intended behavior and (3) the ``white board level" argument or explanation for why the implementation achieves the intended behaviors.
+The most important use of HOL4  was its role as an ``implacable skeptic" that compelled us to really understand the details~\cite{rushby}.
+
+The rest of the paper is organized as follows.
+An overview of BLMS and its main components is given in Section~\ref{sec-BLM} and is followed by a description of SRS implementation in Section~\ref{sec-SRS}.
+The verification approach of SRS is presented in Section~\ref{sec-verification}.
+We conclude in Section~\ref{sec-conclusions} with a summary of lessons learned and future directions.
+
+%The BLMS detects beam loss in the form of counts proportional to the number of stray particles.
 Dangerous loss levels can be expressed as thresholds on the rate of counts over time.
@@
 
-<SIMULATION APPROACH TO ANALYSIS OF SRS>
-<FORMAL APPROACH>
-{\bf myNote: Add 1) the approach currently taken for the analysis of SRS, 2) the approach we took in this paper for the analysis, 3) the reasons behind that  the goals for this case study 4) the main contributions}
+%Simulation of the SRS on sample streams of input is one approach, currently in use~\cite{}, to building an understanding of the SRS's behaviour and confidence that the behaviour is correct.
 
- Rushby's emphasis on the role of mechanized logical deduction as an "implacable skeptic that insists on its human user providing justification" thus rings true.
- %Rushby's description of mechanized logical deduction as an "implacable skeptic that insists on its human user providing justification" rings true for this effort.
+%Formal methods are another approach to building understanding and confidence.
+%The work on constructing proofs may be compared to the work on analysing the results of simulations.
+The work on formulating precise specifications is extra, but is arguably beneficial for both methods.
+<need to be more descriptive here i.e.
+to do number 2) below rather than 3)>
 
-The your process of trying to
-capture the specification via experiment and refinement by a team with
-different backgrounds and expertise is very interesting. the eventual goal is to use HOL4 theorem prover to show
-the correctness of the SRS technique in BLM
+%{\bf myNote: Add 1) the approach currently taken for the analysis of SRS, 2) the approach we took in this paper for the analysis, 3) the reasons behind that  the goals for this case study 4) the main contributions}
 
-%The loss of a very small fraction of the circulating beam may induce a quench of the super conducting magnets or even physical damage to machine components.
-The detection of the lost beam protons allows protection of the equipment against quenches and damage by generating a beam dump trigger when the losses exceed thresholds.
-At each turn of the beam, there will be several thousands of data to record and process in order to decide if the beams should be permitted to continue circulating or their safe extraction is necessary to be triggered.
+%************** Rushby's quote to be added later on *****
+%Rushby's emphasis on the role of mechanized logical deduction as an "implacable skeptic that insists on its human user providing justification" thus rings true.
 
-%In general, the BLM system converts the signal given by the recording of shower particles into electrical signals and decides if threshold limits are being exceeded.
-The beam-permit signal is de-asserted if one detector exceeds the limit and consecutively is transferred to the beam dump system  by the Beam Interlock system.
-
-\section{BLM System Overview}
+\section{BLMS Overview}
+\label{sec-BLM}
 
@@
 The surface card provides data to the Logging, the Post Mortem and the Collimation systems that will be used to drive on-line displays in the control room to allow offline analysis of the losses and to setup automatically the collimators.
-Due to demanding performance requirements, the BLM system is making use of modern field programmable gate arrays (FPGAs), which include the resources needed to design complex processing and can be reprogrammed making them ideal for future upgrades or system specification changes.
+Due to demanding performance requirements, the BLMS is implemented on modern FPGAs, which include the resources needed to design complex processing and can be reprogrammed making them ideal for future upgrades or system specification changes.
 Figure~\ref{fig:BLETC} shows a block diagram of the processes assigned in BLETC FPGA.
@@
 the domain of variation of the beam losses) requested by the specification, the detectors use both Current-to-Frequency Converters (CFCs) and ADC circuitries.
-The data processing block merges these two types of data coming from the same detector into one value and send it to Successive Running Sums (SRS) block.
-The verification of SRS block is the focus of this paper and its implementation is discussed in more detail in the next section.
+The data processing block merges these two types of data coming from the same detector into one value and send it to the SRS block.
+The implementation of the SRS is discussed in more detail in the next section.
 
@@
 \section{Successive Running Sums (SRS)}
+\label{sec-SRS}
 
@@
 
-Beam losses can happen either in a single turn of the beam in the tunnel, with a sudden beam loss, or in progressive losses during numerous turns.
-One-turn failures are called \emph{ultra-fast} losses.
-Multi-turn losses can be divided between \emph{very fast} losses those which happen in less than 10 ms, \emph{fast} losses which happen in more than 10 ms and \emph{steady} losses, where the beam is lost over one second or more~\cite{Schmidt-ICFA}.
+Beam losses can happen either in a single cycle of the beam in the tunnel, with a sudden beam loss, or in progressive losses during numerous cycles.
+One-cycle failures are called \emph{ultra-fast} losses.
+Multi-cycle losses can be divided between \emph{very fast} losses those which happen in less than 10 ms, \emph{fast} losses which happen in more than 10 ms and \emph{steady} losses, where the beam is lost over one second or more~\cite{Schmidt-ICFA}.
 
-The processing of the data collected by the detectors involves a proper analysis of the loss patterns in time and a proper account of the energy of the beam.
+The processing of the data collected by the detectors involves a proper analysis of the loss patterns over time and a proper account of the energy of the beam.
 The procedure for the data processing is based on the idea that a constantly updated moving window can be kept by adding to a register the incoming newest value and subtracting its oldest value (see Figure~\ref{fig:RS-basic}).
@@
 By cascading more of these elements, very long moving windows could be constructed with significantly small memory space.
-The above scheme is the basis for \emph{Successive Running Sums} (SRS) implementation in BLETC.
+The above scheme is the basis for the SRS implementation in BLETC.
 
@@
 
-%A similar configuration for the production of the running sums is used in the BLMTC where in order to increase the efficiency in resources, a set of parameterized multipoint shift registers are used.
-In this configuration, called \emph{Successive Running Sums}, each incoming new value is delayed with a fixed number of cycles by passing it through a shift register and adding the difference of the new and the outputted from the shift register to an accumulator (see Figure ).
-
 \section{Verification of SRS Implementation using HOL}
+\label{sec-verification}
 
@@
 \section{Conclusions and Lessons Learned}
+\label{sec-conclusions}
 

### 2011-03-22 fmics/paper.tex
Tweak some wordings

@@
 We describe a collaborative effort in which the HOL4 theorem prover is being used to formally verify properties of a structure within the Large Hadron Collider (LHC) machine protection system at the European Organization for Nuclear Research (CERN).
-This structure, known as the \emph{Successive Running Sums} (SRS), generates the primary input to the decision logic that must initiate a critical action by the LHC machine protection system in response to the detection of a dangerous level of beam particle loss.
+This structure, known as \emph{Successive Running Sums} (SRS), generates the primary input to the decision logic that must initiate a critical action by the LHC machine protection system in response to the detection of a dangerous level of beam particle loss.
 The use of mechanized logical deduction complements an intensive study of the SRS structure using simulation.
@@
 The Large Hadron Collider (LHC) at the European Organization for Nuclear Research (CERN) is a high-energy particle accelerator.
-It provides a head-on collisions of protons at a center of a mass energy of 14 TeV for high-energy particle physics research.
+It provides head-on collisions of protons at a center of a mass energy of 14 TeV for high-energy particle physics research.
 In order to reach the required magnetic field strengths, the LHC has superconducting magnets cooled with superfluid helium.
-Due to a very high energy stored in the circulating beams (700 MJ), if even a small fraction of the beam particles deposit their energy in the equipment, they can cause the superconductors to transition to normal conducting state.
+Due to the high energy stored in the circulating beams (700 MJ), if even a small fraction of the beam particles deposit their energy in the equipment, they can cause the superconductors to transition to their normal conducting state.
 Such a transition is called a \emph{quench}.
@@
 
-The strategy for machine protection of the LHC is mainly based on the Beam Loss Monitoring System (BLMS), which triggers the safe extraction of the beam if the particle loss exceeds thresholds that are likely to result in a quench.
-At each cycle of the beams around the LHC, the BLMS records and processes several thousands of data points in order to decide if the beams should be permitted to continue circulating or their safe extraction is necessary to be triggered.
+The main strategy for protecting the LHC is based on the Beam Loss Monitoring System (BLMS), which triggers the safe extraction of the beams if particle loss exceeds thresholds that are likely to result in a quench.
+At each cycle of the beams around the LHC, the BLMS records and processes several thousands of data points to decide whether the beams should be permitted to continue circulating or whether their safe extraction should be triggered.
 The processing includes analysis of the loss pattern over time and of the energy of the beam.
 
-The BLMS must respond to dangerous losses quickly, but determining whether losses are dangerous may require analysis of the loss data recorded over long periods of time.
+The BLMS must respond to dangerous losses quickly, but determining whether losses are dangerous may require analysis of loss data recorded over a long period of time.
 Furthermore, the BLMS must continue recording large amounts of data in real-time while processing.
@@
 The component responsible for maintaining these sums is called \emph{Successive Running Sums} (SRS).
-The SRS is implemented in hardware, in order to be fast enough to work in real-time, and on Field Programmable Gate Arrays (FPGAs) in particular so that they can be easily reprogrammed with future upgrades~\cite{Chris-FPGA}.
+The SRS component is implemented in hardware, in order to be fast enough to work in real-time, and on Field Programmable Gate Arrays (FPGAs) in particular so that they can be easily reprogrammed with future upgrades~\cite{Chris-FPGA}.
 %Multiple implementations of the SRS structure are installed regularly around the LHC as part of the BLMS.
@@
 The SRS component has a complex structure and the correctness of its behaviour is critical for safe and productive use of the LHC.
-Any error in SRS implementation would either compromise the availability of the LHC (unnecessary request for beam dump) or safety (not triggering a necessary beam dump).
-The current approach for analysis of SRS is mainly  based on simulation of its behavior on sample streams of input~\cite{}.
+Any error in the SRS implementation would either compromise the availability of the LHC (unnecessary request for beam dump) or safety (not triggering a necessary beam dump).
+The current approach for analysing the SRS implementation is simulation of its behavior on sample streams of input~\cite{}.
 (one more sentence needs to be added on the current verification process)
 
-In this paper, we describe a formal verification approach based on logical deduction for the analysis of SRS.
+In this paper, we describe a formal verification approach based on logical deduction for analysing the SRS implementation.
 (need to add few sentences explaining in a very high level the verification approach/strategy)
 
-Compared to test-based methods, like simulation, formal methods not only offer a much higher confidence in correctness of the systems' behavior, but also help in better understanding their specification.
-One of the main challenges in implementing a formal verification approach for SRS was to capture the intricate details of the system's specification via experiment and refinement with a team with different backgrounds and expertise.
-Our confidence in the SRS design as a result of this effort ultimately rests upon our deep understanding of why the design is correct rather than the fact that we obtained ``Theorem Proved" as the final output of a software tool at the end of a long and intense effort.
-In particular, our use of mechanized logical deduction was a highly iterative process that incrementally refined our understanding of (1) the implementation (2) the intended behavior and (3) the ``white board level" argument or explanation for why the implementation achieves the intended behaviors.
-The most important use of HOL4  was its role as an ``implacable skeptic" that compelled us to really understand the details~\cite{rushby}.
+Compared to test-based methods, like simulation, formal methods not only offer a much higher confidence in correctness of a system's behavior, but also help improve our understanding of its specification.
+One of the challenges in pursuing a formal verification approach for SRS was capturing the intricate details of the system's specification via experiment and refinement with a team with different backgrounds and expertise.
+Our confidence in the SRS design as a result of this effort ultimately rests upon our deep understanding of why the design is correct rather than the fact that we obtained ``Theorem Proved'' as the final output of a software tool at the end of a long and intense effort.
+In particular, our use of mechanized logical deduction was a highly iterative process that incrementally refined our understanding of (1) the implementation (2) the intended behavior and (3) the ``white board level'' argument or explanation for why the implementation achieves its intended behaviors.
+The most important use of HOL4 was its role as an ``implacable skeptic'' that compelled us to really understand the details~\cite{rushby}.
 

### 2011-03-22 fmics/paper.tex
Start drafting first paragraphs of Section 4

@@
 
-\section{Verification of SRS Implementation using HOL}
+\section{Verification of the SRS Implementation using HOL4}
 \label{sec-verification}
 
-Calculation of the running sums is a critical component of BLM.
-In case of no dangerous loss, BLM should not inhabit the beam.
-If it fails, it generates a false alarm and three hours will be lost to recover the previous situation.
-Such false alarms compromise the availability of the system.
-Moreover, in case of a dangerous loss, BLM it has to inhibit the beam permit.
-If it fails there will be at least 30 days of downtime.
-Any error in this computation would either compromise the availability of the system (unnecessary request for beam dump) or safety of the system (nor requesting a necessary beam dump).
-The current practice to check the correctness of the SRS is by simulation and testing in which (explanation of how such a test gets performed).
+%The calculation of running sums is a critical component of the BLMS.
+%If there is no dangerous loss, the BLMS should not inhibit the beam.
+%A false alarm costs at least three hours to resume operation, and compromises the availability of the system.
+%However, if there is a dangerous loss, the BLMS must inhibit the beam.
+%A quench due to dangerous loss costs at least 30 days of downtime.
+%Thus, any error in the calculation compromises either the availability or the safety of the system.
+%The current practice to check the correctness of the SRS is by simulation and testing.
+%(isn't all this in the introduction now? is this the right place to repeat it/give more details?)
+%In this paper we present a formal verification approach for the analysis of SRS implementation in BLM.
 
-In this paper we present a formal verification approach for the analysis of SRS implementation in BLM.
+Our formal verification effort uses mechanised logical deduction, or \emph{theorem proving}.
+In general, theorem proving is used to show that desired properties of a system are logically implied by a formal model of the system.
+We use the HOL4 open source software tool~\cite{HOL4,DBLP:conf/tphol/SlindN08}, which was developed initially at the University of Cambridge, but now by an international team.
+HOL4 enables the construction of theories in higher-order logic~\cite{DBLP:journals/jsyml/Church40}, a formal logic with a similar expressive power to set theory that is widely used for formalising hardware and software models and statements about them.
+The implementation of HOL4 uses Milner's LCF approach~\cite{Milner:1972:LCF:891954}: a small ``kernel'' implementing the primitive rules of the logic, and convenient derived rules and tactics implemented in terms of the kernel.
+Every theorem ultimately comes from the kernel, and this fact provides high assurance of the logical soundness of the verification results obtained using the system.
 
-The formal verification effort described in this paper is based on
- the use of mechanised logical deduction, or theorem proving, to
- demonstrate that desired properties of a system are logically implied
- by a formal model of the system.  Our approach to theorem-proving
- uses the HOL4 open source software tool, which was developed
- initially at the University of Cambridge, but now by an international
- team~\cite{HOL4}.  HOL4 partially automates the task of proving theorems of
- higher-order logic [ref], a formal logic with a similar expressive
- power to set theory that is widely used for formalising hardware and
- software models and statements about them.  Typically, a very large
- portion of the logical deduction is performed automatically by the
- HOL4 system leaving the user to provide the high level proof
- strategy.  The implementation of the HOL4 system is based on Milner's
- LCF approach [ref] and consists of a small “kernel” that provides a
- very high level of confidence in the logical soundness of the
- verification results obtained using the system. The work described
- here could have been done using other systems such as HOL Light
- [ref], Isabelle/HOL~\cite{Isabelle}, ProofPower [ref], PVS [ref] or Coq
- [ref]. The first three of these support essentially the same version
- of higher-order logic as HOL4, whilst PVS and Coq support more
- powerful logics. <What about ACL2?>
-
- While offering less “push button” automation than many other kinds of
- formal verification such as model-checking, machine-assisted theorem
- proving is the most appropriate approach to verifying the BLM SRS,
- since it is unlikely that the correctness theorems, which need lemmas
- proved by manually guided mathematical induction, could be generated
- automatically.
+HOL4 is an \emph{interactive} theorem prover: the user provides the high level proof strategy by composing functions that automate common chains of logical deduction.
+The work described here could have been done using other systems such as HOL Light [ref], Isabelle/HOL~\cite{Isabelle}, ProofPower [ref], PVS [ref] or Coq [ref].
+The first three of these support essentially the same version of higher-order logic as HOL4, whilst PVS and Coq support more powerful logics.
+<What about ACL2?>
+While offering less ``push-button'' automation than many other kinds of formal verification such as model-checking, machine-assisted theorem proving is the most appropriate approach to verifying the SRS, since it is unlikely that the correctness theorems, which need lemmas proved by manually guided mathematical induction, could be generated automatically.
 

### 2011-03-22 fmics/paper.tex
tweak a sentence in intro

@@
 Such a transition is called a \emph{quench}.
-The consequences of a quench range from several hours of down time (to cool down the magnets to reach the superconducting state), to months of repairs (in the case of equipment damage).
+The consequences of a quench range from several hours of downtime (for cooling the magnets down to their superconducting state), to months of repairs (in the case of equipment damage).
 

### 2011-03-22 fmics/paper.tex
Add citations for all those theorem provers
This is probably excessive...

@@
 HOL4 is an \emph{interactive} theorem prover: the user provides the high level proof strategy by composing functions that automate common chains of logical deduction.
-The work described here could have been done using other systems such as HOL Light [ref], Isabelle/HOL~\cite{Isabelle}, ProofPower [ref], PVS [ref] or Coq [ref].
-The first three of these support essentially the same version of higher-order logic as HOL4, whilst PVS and Coq support more powerful logics.
+The work described here could have been done using other systems such as HOL Light~\cite{HOLLight,DBLP:conf/tphol/Harrison09a}, Isabelle/HOL~\cite{Isabelle}, ProofPower~\cite{ProofPower}, PVS~\cite{PVS,DBLP:conf/tphol/OwreS08} or Coq~\cite{Coq}.
+The first three use essentially the same higher-order logic as HOL4, whilst PVS and Coq support more powerful logics.
 <What about ACL2?>
-While offering less ``push-button'' automation than many other kinds of formal verification such as model-checking, machine-assisted theorem proving is the most appropriate approach to verifying the SRS, since it is unlikely that the correctness theorems, which need lemmas proved by manually guided mathematical induction, could be generated automatically.
+While offering less ``push-button'' automation than other kinds of formal verification such as model-checking, machine-assisted theorem proving is appropriate for verifying the SRS, since it is unlikely that the correctness theorems, which need lemmas proved by manually guided mathematical induction, could be generated automatically.
 

### 2011-03-23 fmics/paper.tex
rough outline of another para in sec 4

@@
 
-\usepackage{times,graphicx,epsfig,boxedminipage,url,xspace,array}
+\usepackage{times,graphicx,epsfig,boxedminipage,url,xspace,array,amsmath}
 
@@
 The first three use essentially the same higher-order logic as HOL4, whilst PVS and Coq support more powerful logics.
-<What about ACL2?>
+%<What about ACL2?>
 While offering less ``push-button'' automation than other kinds of formal verification such as model-checking, machine-assisted theorem proving is appropriate for verifying the SRS, since it is unlikely that the correctness theorems, which need lemmas proved by manually guided mathematical induction, could be generated automatically.
 
-The main objective of formalizing the functionality of SRS was to keep the topology parameterized.
-It is very probable that in the future the configuration of the slices and shift registers changes to accommodate higher precision.
-If the model is parameterized then, the same model can be used in the future uses of the system.
-For this, we formalized the definitions of the model by six parameters:
+Our goal is to build a generic model of the SRS structure, and to prove that it calculates approximate running sums of counts within suitable error margins.
+Any statement in formal logic is necessarily precise, and one of our challenges has been characterizing our goal precisely.
+A sketch of the desired statement is \[\forall{n}.\,\text{\textsf{RS} }n=\text{true sum}\pm\text{acceptable error}\]
+<need a formal model of all components of this equation>
+<we only care about 0 <= n <= 11 for the current SRS, but by making the statement forall n, we obtain a generic (parameterized) model>
+<start with "RS" - need to model the SRS structure>
+<we decide to work with discrete time steps and encode the state of the whole SRS structure as a collection of values, one for each component, at each time step>
+
+<two phases: simplified model, more realistic model, plan to do even more realistic model later>
+<the overall sketch of the theory for the more realistic model followed that of the simplified model, but it's not completely clear that if we actually saved time/effort by doing it that way>
+
+%One of our objectives in formally modelling SRS was to keep the topology parameterized.
+%It is likely that the configuration of the slices and shift registers will in the future change to accommodate higher precision.
+%A parameterized model can be used both for current and future versions of the system.
+%For this, we formalized the definitions of the model by six parameters:
 

### 2011-03-24 fmics/paper.tex
some more words for section 4

@@
 Any statement in formal logic is necessarily precise, and one of our challenges has been characterizing our goal precisely.
-A sketch of the desired statement is \[\forall{n}.\,\text{\textsf{RS} }n=\text{true sum}\pm\text{acceptable error}\]
-<need a formal model of all components of this equation>
-<we only care about 0 <= n <= 11 for the current SRS, but by making the statement forall n, we obtain a generic (parameterized) model>
-<start with "RS" - need to model the SRS structure>
-<we decide to work with discrete time steps and encode the state of the whole SRS structure as a collection of values, one for each component, at each time step>
+A sketch of the desired statement is \[\forall{n}.\,\text{\textsf{RS} }n=\text{sum for output $n$}\pm\text{acceptable error}\]
+Although Figure~\ref{fig:srs} would suggest we only care about $0\leq{n}\leq11$, if we can define and prove something for all values of $n$, we obtain a generic model.
 
-<two phases: simplified model, more realistic model, plan to do even more realistic model later>
-<the overall sketch of the theory for the more realistic model followed that of the simplified model, but it's not completely clear that if we actually saved time/effort by doing it that way>
+To prove the equation above, we need to formalize all of its terms.
+First, we make two items explicit: that the equation is to be understood as holding at each time step, and that the terms depend on the stream of input counts over time.
+With these refinements, our statement becomes \[\forall{D\,n\,t}.\,\text{\textsf{RS} }D\,n\,t=\text{sum, given input stream $D$, for output $n$ at time $t$}\pm\text{acceptable error}\]
+
+To formalize the term on the left hand side, we need to build a logical model of the SRS structure.
+Our work proceeded in two phases, starting with a model of a simplified structure and proceeding to a more realistic model.
+In both cases, we represent each piece of the SRS structure that has a value---for example, each cell in each shift register---as a function in HOL that takes the global input stream and current time step as input and returns the value of the piece of the structure at that time.
+Our simplifications in the first phase include
+\begin{enumerate}
+\item Ignore the middle taps in each slice
+\item A more regular arrangement of slices than the actual structure
+\item Ignore subtractor and accumulator, by defining the output of a slice as the sum of its cells directly
+\item Simplify timing <what? how?>
+\item Use natural numbers, rather than finite words, for cell values
+\end{enumerate}
+In the second phase, we drop the first three simplifications by adding middle taps, a more flexible arrangement of slices, and the subtractor and accumulator to our model.
+In both cases, however, the model is simultaneously a simplification and a generalization of the actual structure in the following sense: we model an infinite number of shift registers, each with an infinite number of cells, simply by not specifying their widths or any limits.
+<how does that work>
+<it's not completely clear that if we actually saved time/effort by doing the simplified one first...>
 

### 2011-03-24 fmics/paper.tex
Add description+definitions of model to section 4

@@
 First, we make two items explicit: that the equation is to be understood as holding at each time step, and that the terms depend on the stream of input counts over time.
-With these refinements, our statement becomes \[\forall{D\,n\,t}.\,\text{\textsf{RS} }D\,n\,t=\text{sum, given input stream $D$, for output $n$ at time $t$}\pm\text{acceptable error}\]
+With these refinements, our statement becomes \[\forall{D\,n\,t}.\,\mathsf{RS}\;D\;n\;t=\text{sum, given input stream $D$, for output $n$ at time $t$}\pm\text{acceptable error}\]
+(The input stream is a function of time: $D\;t$ is the input value to the whole SRS structure at time $t$.)
 
@@
 In the second phase, we drop the first three simplifications by adding middle taps, a more flexible arrangement of slices, and the subtractor and accumulator to our model.
-In both cases, however, the model is simultaneously a simplification and a generalization of the actual structure in the following sense: we model an infinite number of shift registers, each with an infinite number of cells, simply by not specifying their widths or any limits.
-<how does that work>
-<it's not completely clear that if we actually saved time/effort by doing the simplified one first...>
+In both cases, however, the model is simultaneously a simplification and a generalization of the actual structure in the following sense: we model an infinite number of slices, each with an infinite number of shift register cells and taps, simply by letting indices range over the natural numbers without explicitly giving limits.
+
+We describe the more realistic (second phase) model here.
+Table~\ref{tab:descriptions} lists the names of the HOL functions we define and their intended meanings, and Figure~\ref{fig:definitions} gives their definitions. 
+For our correctness equation above, we only need the last entry in the table, $\mathsf{RS}$; the first two entries, $\mathsf{tap}$ and $\mathsf{input}$, enable us to specify the exact configuration of the structure, and the remaining entries are used as ``glue'' leading up to the definition of $\mathsf{RS}$.
+By using these intermediary functions, every definition in Figure~\ref{fig:definitions} is \emph{local} (only talks about a small part of the overall SRS structure) and therefore is easy to verify against its intended meaning.
+
+\begin{table}
+\caption{
+Descriptions of the HOL functions comprising our model of the SRS structure.
+\label{tab:descriptions}
+}
+\begin{tabular}{lp{0.8\textwidth}}
+Function&Intended meaning\\
+\hline\\
+\(\mathsf{tap}\;n\;x\)&The position of tap $x$ of slice $n$.\\
+\(\mathsf{input}\;n\)&A pair $(n',x)$ indicating that the input to slice $n$ is output $x$ of slice $n'$.\\
+\(\mathsf{delay}\;n\)&The number of time steps between updates of slice $n$.\\
+\(\mathsf{update\_time}\;n\;t\)&A boolean indicating whether $t$ is an update time for slice $n$.\\
+\(\mathsf{source}\;D\;n\;m\;t\)&The value of the cell that is the direct input to cell $m$ of slice $n$, at time $t$, given input stream $D$.\\
+\(\mathsf{SR}\;D\;n\;m\;t\)&The value of cell $m$ of shift register $n$, at time $t$, given input stream $D$.\\
+\(\mathsf{output}\;D\;n\;x\;t\)&The value of output $x$ of slice $n$, at time $t$, given input stream $D$.\\
+\(\mathsf{RS}\;D\;n\;t\)&The value of Running Sum register $n$ at time $t$, given input stream $D$.
+\end{tabular}
+\end{table}
+
+\begin{figure}
+\caption{
+Definitions of the HOL functions comprising our model of the SRS structure.
+\(\mathsf{tap}\) and \(\mathsf{input}\) are defined to match Figure~\ref{fig:srs}, counting positions within a shift register from $0$.
+Slice $0$ is a fake slice representing the global input; this enables a succinct definition of $\mathsf{source}$.
+The definition of $\mathsf{input}$ when $n=0$ or when $n>6$ does not change the structure represented, so we use definitions convenient for proving theorems.
+A similar comment applies to other definitions when $n$, representing a slice number, is $0$, or when $x$, representing a shift register position, is above $1$.
+\label{fig:definitions}}
+\begin{align*}
+\mathsf{tap}\;0\;0&=0&\mathsf{tap}\;0\;x&=0\\
+\mathsf{tap}\;1\;0&=1-1&\mathsf{tap}\;1\;x&=2-1\\
+\mathsf{tap}\;2\;0&=8-1&\mathsf{tap}\;2\;x&=16-1\\
+&\dots&\mathsf{tap}\;6\;x&=128-1\quad\text{(where $x>0$)}\\
+\mathsf{input}\;0&=(0,0)&\mathsf{input}\;1&=(0,0)\\
+\mathsf{input}\;2&=(0,0)&\mathsf{input}\;3&=(1,1)\\
+\mathsf{input}\;4&=(3,0)&\mathsf{input}\;5&=(4,0)\\
+\mathsf{input}\;6&=(4,1)&\mathsf{input}\;n&=(n-1,0)\quad\text{(where $n>6$)}
+\end{align*}
+\begin{align*}
+\mathsf{delay}\;0&=1\\
+\mathsf{delay}\;n&=\mathsf{delay}\;n'\times((\mathsf{tap}\;n'\;x)+1)\\
+&\qquad\text{where $(n',x)=\mathsf{input\;n}$ (and $n>0$)}
+\end{align*}
+\begin{align*}
+\mathsf{update\_time}\;n\;t&\iff(t\operatorname{mod}\mathsf{delay}\;n=0)
+\end{align*}
+\begin{align*}
+\mathsf{source}\;D\;n\;0\;t&=\mathsf{output}\;D\;n'\;x\;t\quad\text{where $(n',x)=\mathsf{input\;n}$}\\
+\mathsf{source}\;D\;n\;m\;t&=\;\mathsf{SR}\;D\;n\;(m-1)\;\quad\text{(where $m>0$)}
+\end{align*}
+\begin{align*}
+\mathsf{output}\;D\;0\;x\;t&=D\;t\\
+\mathsf{output}\;D\;n\;x\;0&=0\\
+\mathsf{output}\;D\;n\;x\;t&=
+\text{if $\mathsf{update\_time}\;n\;t$}\\
+&\qquad\text{then }((\mathsf{output}\;D\;n\;x\;(t-1))+(\mathsf{source}\;D\;n\;0\;(t-1)))-(\mathsf{SR}\;D\;n(\mathsf{tap}\;n\;x)\;(t-1))\\
+&\qquad\text{else }\mathsf{output}\;D\;n\;x\;(t-1)\qquad\text{(where $t>0$)}
+\end{align*}
+\begin{align*}
+\mathsf{SR}\;D\;n\;m\;0&=0\\
+\mathsf{SR}\;D\;n\;m\;t&=\text{if $\mathsf{update\_time}\;n\;t$}\\
+&\qquad\text{then }\mathsf{source}\;D\;n\;m\;(t-1)\\
+&\qquad\text{else }\mathsf{SR}\;D\;n\;m\;t\qquad\text{(where $t>0$)}
+\end{align*}
+\begin{align*}
+\mathsf{RS}\;D\;n\;t&=\mathsf{output}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)
+\end{align*}
+\end{figure}
 

### 2011-03-24 fmics/paper.tex
remove trailing whitespace
merge changes from dropbox

@@
 Any error in the SRS implementation would either compromise the availability of the LHC (unnecessary request for beam dump) or safety (not triggering a necessary beam dump).
-The current approach for analysing the SRS implementation is simulation of its behavior on sample streams of input~\cite{}.
+The current approach for analyzing the SRS implementation is simulation of its behavior on sample streams of input~\cite{}.
 (one more sentence needs to be added on the current verification process)
 
-In this paper, we describe a formal verification approach based on logical deduction for analysing the SRS implementation.
+In this paper, we describe a formal verification approach based on logical deduction for analyzing the SRS implementation.
 (need to add few sentences explaining in a very high level the verification approach/strategy)
@@
 One of the challenges in pursuing a formal verification approach for SRS was capturing the intricate details of the system's specification via experiment and refinement with a team with different backgrounds and expertise.
-Our confidence in the SRS design as a result of this effort ultimately rests upon our deep understanding of why the design is correct rather than the fact that we obtained ``Theorem Proved'' as the final output of a software tool at the end of a long and intense effort.
-In particular, our use of mechanized logical deduction was a highly iterative process that incrementally refined our understanding of (1) the implementation (2) the intended behavior and (3) the ``white board level'' argument or explanation for why the implementation achieves its intended behaviors.
-The most important use of HOL4 was its role as an ``implacable skeptic'' that compelled us to really understand the details~\cite{rushby}.
+Our confidence in the SRS design as a result of this effort ultimately rests upon our deep understanding of why the design is correct rather than the fact that we obtained ``Theorem Proved" as the final output of a software tool at the end of a long and intense effort.
+In particular, our use of mechanized logical deduction was a highly iterative process that incrementally refined our understanding of (1) the implementation (2) the intended behavior and (3) the ``white board level" argument or explanation for why the implementation achieves the intended behaviors.
+The most important use of HOL4  was its role as an ``implacable skeptic" that compelled us to really understand the details~\cite{rushby}.
 
@@
 
-In order to construct long moving windows, it is required to keep long histories of received data.
-The technique employed to reach long integration periods with relatively small in length shift registers uses consecutive storage of sums of the received values.
+In order to construct long moving windows, it is required to keep long histories of received data, referred to as \emph{count} values.
+The technique employed to reach long integration periods with relatively small in length shift registers uses consecutive storage of sums of the counts.
 Instead of storing all values needed for the sum, this technique stores successively parts of the total sum using only a fraction of the otherwise needed memory space.
@@
 
-Our goal is to build a generic model of the SRS structure, and to prove that it calculates approximate running sums of counts within suitable error margins.
-Any statement in formal logic is necessarily precise, and one of our challenges has been characterizing our goal precisely.
-A sketch of the desired statement is \[\forall{n}.\,\text{\textsf{RS} }n=\text{sum for output $n$}\pm\text{acceptable error}\]
-Although Figure~\ref{fig:srs} would suggest we only care about $0\leq{n}\leq11$, if we can define and prove something for all values of $n$, we obtain a generic model.
+Our goal is to build a generic model of the SRS structure, and to prove that it satisfies its specification, i.e.
+it calculates approximate running sums of received count values within acceptable error margins.
+%Any statement in formal logic is necessarily precise, and one of our challenges has been characterizing our goal precisely.
+%One of our challenges has been characterizing this goal precisely
+Let $\mathsf{output} \, n$ denote the sum of the last $m$ received count values and $\mathsf{RS} \, n$ denote one of the outputs computed by the slices in the structure of SRS which is supposed to compute the sum of the last $m$ received count values with some delay.
+A sketch of the desired correctness statement is \[\forall{n} \cdot \, \mathsf{RS} \,n= \mathsf{output} \, n \pm\text{acceptable error}\]
+Although Figure~\ref{fig:srs} suggests that in the SRS structure there are only twelve outputs  (i.e.
+$1\leq{n}\leq11$), we would obtain a more generic result (that is useful in future upgrades of the system) if we prove the correctness of the above statement for all values of $n$.
 
-To prove the equation above, we need to formalize all of its terms.
-First, we make two items explicit: that the equation is to be understood as holding at each time step, and that the terms depend on the stream of input counts over time.
-With these refinements, our statement becomes \[\forall{D\,n\,t}.\,\mathsf{RS}\;D\;n\;t=\text{sum, given input stream $D$, for output $n$ at time $t$}\pm\text{acceptable error}\]
-(The input stream is a function of time: $D\;t$ is the input value to the whole SRS structure at time $t$.)
+%To prove the equation above, we need to formalize all of its terms.
+In order to make the above correctness statement more precise, we need to include the notion of time.
+The count values arrive at the input of the SRS block every 40 micro-seconds that is abstracted by one time step in our logical model.
+We formalize the input stream to SRS as a function of time: $D\;t$ denotes the input value to the SRS structure at time $t$.
+In addition, the terms in the above statement depend on the stream of input counts over time.
+With these refinements, the correctness statement is defines as: \[\forall{D\,n\,t} \cdot \,\mathsf{RS} \,D\,n\,t = \mathsf{output} \,D\,\,n\, t \pm\text{acceptable error}\]
 
-To formalize the term on the left hand side, we need to build a logical model of the SRS structure.
-Our work proceeded in two phases, starting with a model of a simplified structure and proceeding to a more realistic model.
-In both cases, we represent each piece of the SRS structure that has a value---for example, each cell in each shift register---as a function in HOL that takes the global input stream and current time step as input and returns the value of the piece of the structure at that time.
-Our simplifications in the first phase include
+%First, we make two items explicit: that the equation is to be understood as holding at each time step, and that the terms depend on the stream of input counts over time.
+%With these refinements, our statement becomes \[\forall{D\,n\,t}.\,\mathsf{RS}\;D\;n\;t=\text{sum, given input stream $D$, for output $n$ at time $t$}\pm\text{acceptable error}\]
+%(The input stream is a function of time: $D\;t$ is the input value to the whole SRS structure at time $t$.)
+
+%To formalize the term on the left hand side, we need to build a logical model of the SRS structure.
+The first natural step to prove the correctness of the above statement is to build a logical model of the SRS structure.
+Our work proceeded in two phases, starting with a model of a simplified structure and proceeding to a model which is closer to the SRS implementation in BLMS.
+Our main approach in building the logical model of SRS (in both phases) was to model each building block of the SRS structure that holds a count value--for example, each cell in each shift register--as a function in HOL that takes the input stream and current time step as inputs and returns the value which the corresponding building block holds at that time.
+
+%In both cases, we represent each piece of the SRS structure that has a value---for example, each cell in each shift register---as a function in HOL that takes the global input stream and current time step as input and returns the value of the piece of the structure at that time.
+
+Our simplifications in the first phase include:
 \begin{enumerate}
-\item Ignore the middle taps in each slice
-\item A more regular arrangement of slices than the actual structure
-\item Ignore subtractor and accumulator, by defining the output of a slice as the sum of its cells directly
-\item Simplify timing <what? how?>
-\item Use natural numbers, rather than finite words, for cell values
+\item Masking the middle taps in each slice,
+\item Considering a more regular arrangement of slices than the SRS implementation in BLMS,
+\item Masking the subtractors and the accumulators in each slice, instead, defining the output of a slice as the sum of the values in its cells directly,
+\item Simplifying timing by ignoring some of the delays that each building block may introduce,
+\item Using natural numbers, rather than finite words, for count values
 \end{enumerate}
-In the second phase, we drop the first three simplifications by adding middle taps, a more flexible arrangement of slices, and the subtractor and accumulator to our model.
-In both cases, however, the model is simultaneously a simplification and a generalization of the actual structure in the following sense: we model an infinite number of slices, each with an infinite number of shift register cells and taps, simply by letting indices range over the natural numbers without explicitly giving limits.
 
+Building a logical model for simplified structure of the SRS helped us improve our understanding of the behavior of the SRS.
+{\bf what else did we learn through this process? If there was no particular advantage, I don't think we should mention the simplified version}.
+
+In the second phase, we eliminated the first four simplifications (listed above) by adding middle taps, a more flexible arrangement of slices, the subtractor and accumulator, and considering the delays that each block may introduce.
+In both cases, however, the model is simultaneously a simplification and a generalization of the actual structure of SRS in BLMS in the following aspects: we model an infinite number of slices (rather than six slices), each with an infinite number of shift register cells and taps (rather than fixed width shift registers and only two taps), simply by letting indices range over the natural numbers without explicitly giving limits.
+Such a parameterization would make the model and the analysis applicable to the future upgraded/modified versions of the system.
+
+%%%%% upto here %%%
 We describe the more realistic (second phase) model here.

### 2011-03-24 fmics/paper.tex
revise the last changes

@@
 
-Our goal is to build a generic model of the SRS structure, and to prove that it satisfies its specification, i.e.
-it calculates approximate running sums of received count values within acceptable error margins.
+Our goal is to build a generic model of the SRS structure, and to prove that it satisfies its specification, that it calculates approximate running sums of received count values within acceptable error margins.
 %Any statement in formal logic is necessarily precise, and one of our challenges has been characterizing our goal precisely.
 %One of our challenges has been characterizing this goal precisely
-Let $\mathsf{output} \, n$ denote the sum of the last $m$ received count values and $\mathsf{RS} \, n$ denote one of the outputs computed by the slices in the structure of SRS which is supposed to compute the sum of the last $m$ received count values with some delay.
-A sketch of the desired correctness statement is \[\forall{n} \cdot \, \mathsf{RS} \,n= \mathsf{output} \, n \pm\text{acceptable error}\]
-Although Figure~\ref{fig:srs} suggests that in the SRS structure there are only twelve outputs  (i.e.
-$1\leq{n}\leq11$), we would obtain a more generic result (that is useful in future upgrades of the system) if we prove the correctness of the above statement for all values of $n$.
+Let $\mathsf{RS}\;n$ denote, as in Figure~\ref{fig:srs}, the output of slice $n$ in the SRS structure, which is supposed to compute a sum of received count values with some delay, and let $\mathsf{exact}\;n$ denote this sum.
+A sketch of the desired correctness statement is \[\forall{n} \cdot \, \mathsf{RS}\;n= \mathsf{exact}\; n \pm\text{acceptable error}\]
+Although Figure~\ref{fig:srs} suggests that the SRS structure has only twelve outputs  (i.e.
+$0\leq{n}\leq11$), we obtain a more generic result (that is useful in future upgrades of the system) by proving the correctness of the above statement for all values of $n$.
 
 %To prove the equation above, we need to formalize all of its terms.
-In order to make the above correctness statement more precise, we need to include the notion of time.
-The count values arrive at the input of the SRS block every 40 micro-seconds that is abstracted by one time step in our logical model.
-We formalize the input stream to SRS as a function of time: $D\;t$ denotes the input value to the SRS structure at time $t$.
-In addition, the terms in the above statement depend on the stream of input counts over time.
-With these refinements, the correctness statement is defines as: \[\forall{D\,n\,t} \cdot \,\mathsf{RS} \,D\,n\,t = \mathsf{output} \,D\,\,n\, t \pm\text{acceptable error}\]
+To make the above correctness statement more precise, we need to include the notion of time.
+The count values arrive at the input of the SRS block every 40 micro-seconds, which we abstract as a single time step in our logical model.
+We formalize the input stream as a function of time: $D\;t$ denotes the input value to the SRS structure at time $t$.
+In addition, the terms in the above statement depend on this stream of input counts.
+With these refinements, the correctness statement is: \[\forall{D\,n\,t} \cdot \,\mathsf{RS}\;D\;n\;t = \mathsf{exact}\;D\;\;n\;t\pm\text{acceptable error}\]
 
@@
 Our work proceeded in two phases, starting with a model of a simplified structure and proceeding to a model which is closer to the SRS implementation in BLMS.
-Our main approach in building the logical model of SRS (in both phases) was to model each building block of the SRS structure that holds a count value--for example, each cell in each shift register--as a function in HOL that takes the input stream and current time step as inputs and returns the value which the corresponding building block holds at that time.
+In both phases, we model each building block of the SRS structure that holds a count value--for example, each cell in each shift register--as a function in HOL that takes the input stream and current time step as inputs and returns the value the corresponding building block holds at that time.
 
@@
 \item Masking the subtractors and the accumulators in each slice, instead, defining the output of a slice as the sum of the values in its cells directly,
-\item Simplifying timing by ignoring some of the delays that each building block may introduce,
+\item Simplifying timing by ignoring some of the delays that each building block may introduce, {\bf this is not true - there were delays in the simplified model}
 \item Using natural numbers, rather than finite words, for count values
@@
 Building a logical model for simplified structure of the SRS helped us improve our understanding of the behavior of the SRS.
-{\bf what else did we learn through this process? If there was no particular advantage, I don't think we should mention the simplified version}.
+{\bf what else did we learn through this process? If there was no particular advantage, I don't think we should mention the simplified version.
+(I don't think there was any particular advantage -- Ramana)}.
 
 In the second phase, we eliminated the first four simplifications (listed above) by adding middle taps, a more flexible arrangement of slices, the subtractor and accumulator, and considering the delays that each block may introduce.
-In both cases, however, the model is simultaneously a simplification and a generalization of the actual structure of SRS in BLMS in the following aspects: we model an infinite number of slices (rather than six slices), each with an infinite number of shift register cells and taps (rather than fixed width shift registers and only two taps), simply by letting indices range over the natural numbers without explicitly giving limits.
+In both cases, however, the model is simultaneously a simplification and a generalization of the actual structure of SRS in BLMS in the following way: we model an infinite number of slices (rather than six slices), each with an infinite number of shift register cells and taps (rather than fixed width shift registers and only two taps), simply by letting indices range over the natural numbers without explicitly giving limits.
 Such a parameterization would make the model and the analysis applicable to the future upgraded/modified versions of the system.

### 2011-03-24 fmics/paper.tex
tweaks

@@
 In addition, the terms in the above statement depend on this stream of input counts.
-With these refinements, the correctness statement is: \[\forall{D\,n\,t} \cdot \,\mathsf{RS}\;D\;n\;t = \mathsf{exact}\;D\;\;n\;t\pm\text{acceptable error}\]
+With these refinements, the correctness statement becomes: \[\forall{D\,n\,t} \cdot \,\mathsf{RS}\;D\;n\;t = \mathsf{exact}\;D\;\;n\;t\pm\text{acceptable error}\]
 
@@
 %To formalize the term on the left hand side, we need to build a logical model of the SRS structure.
-The first natural step to prove the correctness of the above statement is to build a logical model of the SRS structure.
+The first natural step to prove the correctness statement is to build a logical model of the SRS structure.
 Our work proceeded in two phases, starting with a model of a simplified structure and proceeding to a model which is closer to the SRS implementation in BLMS.
-In both phases, we model each building block of the SRS structure that holds a count value--for example, each cell in each shift register--as a function in HOL that takes the input stream and current time step as inputs and returns the value the corresponding building block holds at that time.
+In both phases, we model each building block of the SRS structure that holds a count value -- for example, each cell in each shift register -- as a function in HOL that takes the input stream and current time step as inputs and returns the value the corresponding building block holds at that time.
 
@@
 In both cases, however, the model is simultaneously a simplification and a generalization of the actual structure of SRS in BLMS in the following way: we model an infinite number of slices (rather than six slices), each with an infinite number of shift register cells and taps (rather than fixed width shift registers and only two taps), simply by letting indices range over the natural numbers without explicitly giving limits.
-Such a parameterization would make the model and the analysis applicable to the future upgraded/modified versions of the system.
+This parameterization makes the model and the analysis more likely to be applicable to future upgraded/modified versions of the system.
 

### 2011-03-25 fmics/paper.tex
merge in changes from dropbox

@@
 Our goal is to build a generic model of the SRS structure, and to prove that it satisfies its specification, that it calculates approximate running sums of received count values within acceptable error margins.
-%Any statement in formal logic is necessarily precise, and one of our challenges has been characterizing our goal precisely.
-%One of our challenges has been characterizing this goal precisely
-Let $\mathsf{RS}\;n$ denote, as in Figure~\ref{fig:srs}, the output of slice $n$ in the SRS structure, which is supposed to compute a sum of received count values with some delay, and let $\mathsf{exact}\;n$ denote this sum.
-A sketch of the desired correctness statement is \[\forall{n} \cdot \, \mathsf{RS}\;n= \mathsf{exact}\; n \pm\text{acceptable error}\]
+Let $\mathsf{RS}\;n$ denote, as in Figure~\ref{fig:srs}, the $n$-th output of the SRS structure, which is supposed to compute a sum of received count values with some delay, and let $\mathsf{true\_sum}\;n$ denote this sum.
+A sketch of the desired correctness statement is \[\forall{n} \cdot \, \mathsf{RS}\;n= \mathsf{true\_sum}\; n \pm\text{acceptable error}\]
 Although Figure~\ref{fig:srs} suggests that the SRS structure has only twelve outputs  (i.e.
@@
 
-%To prove the equation above, we need to formalize all of its terms.
 To make the above correctness statement more precise, we need to include the notion of time.
@@
 In addition, the terms in the above statement depend on this stream of input counts.
-With these refinements, the correctness statement becomes: \[\forall{D\,n\,t} \cdot \,\mathsf{RS}\;D\;n\;t = \mathsf{exact}\;D\;\;n\;t\pm\text{acceptable error}\]
+With these refinements, the correctness statement becomes: \[\forall{D\,n\,t} \cdot \,\mathsf{RS}\;D\;n\;t = \mathsf{true\_sum}\;D\;n\;t\pm\text{acceptable error}\]
 
-%First, we make two items explicit: that the equation is to be understood as holding at each time step, and that the terms depend on the stream of input counts over time.
-%With these refinements, our statement becomes \[\forall{D\,n\,t}.\,\mathsf{RS}\;D\;n\;t=\text{sum, given input stream $D$, for output $n$ at time $t$}\pm\text{acceptable error}\]
-%(The input stream is a function of time: $D\;t$ is the input value to the whole SRS structure at time $t$.)
-
-%To formalize the term on the left hand side, we need to build a logical model of the SRS structure.
 The first natural step to prove the correctness statement is to build a logical model of the SRS structure.
@@
 In both phases, we model each building block of the SRS structure that holds a count value -- for example, each cell in each shift register -- as a function in HOL that takes the input stream and current time step as inputs and returns the value the corresponding building block holds at that time.
-
-%In both cases, we represent each piece of the SRS structure that has a value---for example, each cell in each shift register---as a function in HOL that takes the global input stream and current time step as input and returns the value of the piece of the structure at that time.
 
@@
 
-%%%%% upto here %%%
-We describe the more realistic (second phase) model here.
-Table~\ref{tab:descriptions} lists the names of the HOL functions we define and their intended meanings, and Figure~\ref{fig:definitions} gives their definitions.
-For our correctness equation above, we only need the last entry in the table, $\mathsf{RS}$; the first two entries, $\mathsf{tap}$ and $\mathsf{input}$, enable us to specify the exact configuration of the structure, and the remaining entries are used as ``glue'' leading up to the definition of $\mathsf{RS}$.
-By using these intermediary functions, every definition in Figure~\ref{fig:definitions} is \emph{local} (only talks about a small part of the overall SRS structure) and therefore is easy to verify against its intended meaning.
+%We describe the more realistic (second phase) model here.
+Table~\ref{tab:descriptions} lists the HOL functions we defined in the second phase along with their intended meanings.
+The arrangement of the slices is described by functions $\mathsf{input}\;n$ and $\mathsf{tap}\;n\;x$.
+The read delay of a slice is modeled by $\mathsf{delay}\;n$.
+Each slice is modeled by three functions: $\mathsf{SR}\;D\;n\;m\;t$ represents the value of each cell of the slice's shift register, $\mathsf{source}\;D\;n\;m\;t$ represents the direct input of each cell, and $\mathsf{output}\;D\;n\;x\;t$ represents the value of the slice's output at its taps.
+The $\mathsf{RS}\;D\;n\;t$ models the running sums and $\mathsf{update\_time}\;n\;t$ checks if it is the time to refresh the contents of slice $n$.
+{\bf should we explain why we have both output function and RS?}
 
-\begin{table}
+The formal definitions of the functions listed in Table~\ref{tab:descriptions} are given in Figure~\ref{fig:definitions}.
+Note that the definition of $\mathsf{input}$ when $n=0$ or when $n>6$ does not change the structure presented. %-- so we use definitions convenient for proving theorems.
+A similar comment applies to other definitions when $n$, representing a slice number, is $0$, or when $x$, representing a shift register position, is greater than $1$.
+Every definition in Figure~\ref{fig:definitions} is \emph{local} (only represents a small part of the SRS structure) and therefore the verification against its intended meaning is relatively straightforward.
+
+ %Table~\ref{tab:descriptions} lists the names of the HOL functions we define and their intended meanings, and Figure~\ref{fig:definitions} gives their definitions.
+
+%For our correctness equation above, we only need the last entry in the table, $\mathsf{RS}$; the first two entries, $\mathsf{tap}$ and $\mathsf{input}$, enable us to specify the exact configuration of the structure, and the remaining entries are used as ``glue'' leading up to the definition of $\mathsf{RS}$.
+
+%By using these intermediary functions, every definition in Figure~\ref{fig:definitions} is \emph{local} (only talks about a small part of the overall SRS structure) and therefore is easy to verify against its intended meaning.
+
+\begin{table}[t]
 \caption{
@@
 \(\mathsf{delay}\;n\)&The number of time steps between updates of slice $n$.\\
-\(\mathsf{update\_time}\;n\;t\)&A boolean indicating whether $t$ is an update time for slice $n$.\\
 \(\mathsf{source}\;D\;n\;m\;t\)&The value of the cell that is the direct input to cell $m$ of slice $n$, at time $t$, given input stream $D$.\\
@@
 \(\mathsf{output}\;D\;n\;x\;t\)&The value of output $x$ of slice $n$, at time $t$, given input stream $D$.\\
-\(\mathsf{RS}\;D\;n\;t\)&The value of Running Sum register $n$ at time $t$, given input stream $D$.
+\(\mathsf{RS}\;D\;n\;t\)&The value of Running Sum slice (?) $n$ at time $t$, given input stream $D$.\\
+\(\mathsf{update\_time}\;n\;t\)&A boolean indicating whether $t$ is an update time for slice $n$.
 \end{tabular}
@@
 Definitions of the HOL functions comprising our model of the SRS structure.
-\(\mathsf{tap}\) and \(\mathsf{input}\) are defined to match Figure~\ref{fig:srs}, counting positions within a shift register from $0$.
-Slice $0$ is a fake slice representing the global input; this enables a succinct definition of $\mathsf{source}$.
-The definition of $\mathsf{input}$ when $n=0$ or when $n>6$ does not change the structure represented, so we use definitions convenient for proving theorems.
-A similar comment applies to other definitions when $n$, representing a slice number, is $0$, or when $x$, representing a shift register position, is above $1$.
+\(\mathsf{tap}\) and \(\mathsf{input}\) are defined to match Figure~\ref{fig:srs}, %counting positions within a shift register from $0$.
+Slice $0$ is a fake slice representing the SRS input; this enables a succinct definition of $\mathsf{source}$.
+%The definition of $\mathsf{input}$ when $n=0$ or when $n>6$ does not change the structure represented, so we use definitions convenient for proving theorems.
+%A similar comment applies to other definitions when $n$, representing a slice number, is $0$, or when $x$, representing a shift register position, is above $1$.
 \label{fig:definitions}}
 \begin{align*}
-\mathsf{tap}\;0\;0&=0&\mathsf{tap}\;0\;x&=0\\
-\mathsf{tap}\;1\;0&=1-1&\mathsf{tap}\;1\;x&=2-1\\
-\mathsf{tap}\;2\;0&=8-1&\mathsf{tap}\;2\;x&=16-1\\
-&\dots&\mathsf{tap}\;6\;x&=128-1\quad\text{(where $x>0$)}\\
+\mathsf{tap}\;0\;0&=0&\mathsf{tap}\;0\;1&=0\\
+\mathsf{tap}\;1\;0&=1-1&\mathsf{tap}\;1\;1&=2-1\\
+\mathsf{tap}\;2\;0&=8-1&\mathsf{tap}\;2\;1&=16-1\\
+&\dots&\mathsf{tap}\;6\;x&=128-1\quad\text{(where $x>0$)}\\\\
 \mathsf{input}\;0&=(0,0)&\mathsf{input}\;1&=(0,0)\\

### 2011-03-25 fmics/paper.tex
Correct the definition of tap
Explain why we have both RS and output

@@
 Each slice is modeled by three functions: $\mathsf{SR}\;D\;n\;m\;t$ represents the value of each cell of the slice's shift register, $\mathsf{source}\;D\;n\;m\;t$ represents the direct input of each cell, and $\mathsf{output}\;D\;n\;x\;t$ represents the value of the slice's output at its taps.
-The $\mathsf{RS}\;D\;n\;t$ models the running sums and $\mathsf{update\_time}\;n\;t$ checks if it is the time to refresh the contents of slice $n$.
-{\bf should we explain why we have both output function and RS?}
+The function $\mathsf{RS}\;D\;n\;t$ models the running sums and $\mathsf{update\_time}\;n\;t$ checks if it is time to refresh the contents of a slice.
+We have both $\mathsf{RS}$ and $\mathsf{output}$ functions, though $\mathsf{RS}$ is easily defined in terms of $\mathsf{output}$, because $\mathsf{RS}$ represents a whole SRS output (indexed by a single number), whereas $\mathsf{output}$ represents an individual slice output (indexed by slice and tap).
+By separating $\mathsf{RS}$ and $\mathsf{output}$, we allow for designs where some slice outputs are not SRS outputs, but may still be used internally between slices.
 
@@
 \begin{align*}
-\mathsf{tap}\;0\;0&=0&\mathsf{tap}\;0\;1&=0\\
-\mathsf{tap}\;1\;0&=1-1&\mathsf{tap}\;1\;1&=2-1\\
-\mathsf{tap}\;2\;0&=8-1&\mathsf{tap}\;2\;1&=16-1\\
-&\dots&\mathsf{tap}\;6\;x&=128-1\quad\text{(where $x>0$)}\\\\
+\mathsf{tap}\;0\;0&=0&\mathsf{tap}\;0\;x&=0\\
+\mathsf{tap}\;1\;0&=1-1&\mathsf{tap}\;1\;x&=2-1\\
+\mathsf{tap}\;2\;0&=8-1&\mathsf{tap}\;2\;x&=16-1\\
+&\dots&\mathsf{tap}\;6\;x&=128-1\quad\text{(where each $x>0$)}\\\\
 \mathsf{input}\;0&=(0,0)&\mathsf{input}\;1&=(0,0)\\

### 2011-03-25 fmics/paper.tex
Clarify definition of RS in Table 1

@@
 \(\mathsf{SR}\;D\;n\;m\;t\)&The value of cell $m$ of shift register $n$, at time $t$, given input stream $D$.\\
-\(\mathsf{output}\;D\;n\;x\;t\)&The value of output $x$ of slice $n$, at time $t$, given input stream $D$.\\
-\(\mathsf{RS}\;D\;n\;t\)&The value of Running Sum slice (?) $n$ at time $t$, given input stream $D$.\\
+\(\mathsf{output}\;D\;n\;x\;t\)&The output at tap $x$ of slice $n$, at time $t$, given input stream $D$.\\
+\(\mathsf{RS}\;D\;n\;t\)&The $n$-th running sum at time $t$, given input stream $D$.\\
 \(\mathsf{update\_time}\;n\;t\)&A boolean indicating whether $t$ is an update time for slice $n$.

### 2011-03-25 fmics/paper.tex
turn a comma into a full-stop in fig 7 caption
merge with dropbox

@@
 This collaborative effort has individuals with diverse backgrounds ranging from theoretical physics to system safety.
-One interesting result is the extent to which the use of a formal method has compelled the stakeholders to clarify intricate details of the SRS structure and behaviour.
+%One interesting result is the extent to which
+The use of a formal method has compelled the stakeholders to clarify intricate details of the SRS structure and behaviour.
 
@@
 \item Simplifying timing by ignoring some of the delays that each building block may introduce, {\bf this is not true - there were delays in the simplified model}
-\item Using natural numbers, rather than finite words, for count values
+\item Using natural numbers, rather than finite words, for count values.
 \end{enumerate}
@@
 
-%We describe the more realistic (second phase) model here.
 Table~\ref{tab:descriptions} lists the HOL functions we defined in the second phase along with their intended meanings.
@@
 The function $\mathsf{RS}\;D\;n\;t$ models the running sums and $\mathsf{update\_time}\;n\;t$ checks if it is time to refresh the contents of a slice.
-We have both $\mathsf{RS}$ and $\mathsf{output}$ functions, though $\mathsf{RS}$ is easily defined in terms of $\mathsf{output}$, because $\mathsf{RS}$ represents a whole SRS output (indexed by a single number), whereas $\mathsf{output}$ represents an individual slice output (indexed by slice and tap).
-By separating $\mathsf{RS}$ and $\mathsf{output}$, we allow for designs where some slice outputs are not SRS outputs, but may still be used internally between slices.
+We have both $\mathsf{RS}$ and $\mathsf{output}$ functions, though $\mathsf{RS}$ is easily defined in terms of $\mathsf{output}$, because $\mathsf{RS}$ represents an SRS output (indexed by a single number), whereas $\mathsf{output}$ represents an individual slice output (indexed by slice and tap numbers).
+By separating $\mathsf{RS}$ and $\mathsf{output}$, we allow for designs where some slice outputs are not SRS outputs, but may still be used internally as inputs to other slices.
 
 The formal definitions of the functions listed in Table~\ref{tab:descriptions} are given in Figure~\ref{fig:definitions}.
-Note that the definition of $\mathsf{input}$ when $n=0$ or when $n>6$ does not change the structure presented. %-- so we use definitions convenient for proving theorems.
+Note that the definition of $\mathsf{input}$ when $n=0$ or when $n>6$ does not change the structure presented.
 A similar comment applies to other definitions when $n$, representing a slice number, is $0$, or when $x$, representing a shift register position, is greater than $1$.
 Every definition in Figure~\ref{fig:definitions} is \emph{local} (only represents a small part of the SRS structure) and therefore the verification against its intended meaning is relatively straightforward.
-
- %Table~\ref{tab:descriptions} lists the names of the HOL functions we define and their intended meanings, and Figure~\ref{fig:definitions} gives their definitions.
-
-%For our correctness equation above, we only need the last entry in the table, $\mathsf{RS}$; the first two entries, $\mathsf{tap}$ and $\mathsf{input}$, enable us to specify the exact configuration of the structure, and the remaining entries are used as ``glue'' leading up to the definition of $\mathsf{RS}$.
-
-%By using these intermediary functions, every definition in Figure~\ref{fig:definitions} is \emph{local} (only talks about a small part of the overall SRS structure) and therefore is easy to verify against its intended meaning.
 
@@
 Definitions of the HOL functions comprising our model of the SRS structure.
-\(\mathsf{tap}\) and \(\mathsf{input}\) are defined to match Figure~\ref{fig:srs}, %counting positions within a shift register from $0$.
+\(\mathsf{tap}\) and \(\mathsf{input}\) are defined to match Figure~\ref{fig:srs}.
 Slice $0$ is a fake slice representing the SRS input; this enables a succinct definition of $\mathsf{source}$.
-%The definition of $\mathsf{input}$ when $n=0$ or when $n>6$ does not change the structure represented, so we use definitions convenient for proving theorems.
-%A similar comment applies to other definitions when $n$, representing a slice number, is $0$, or when $x$, representing a shift register position, is above $1$.
 \label{fig:definitions}}

### 2011-03-25 fmics/paper.tex
start formatting statement of one of the main theorems

@@
 A similar comment applies to other definitions when $n$, representing a slice number, is $0$, or when $x$, representing a shift register position, is greater than $1$.
-Every definition in Figure~\ref{fig:definitions} is \emph{local} (only represents a small part of the SRS structure) and therefore the verification against its intended meaning is relatively straightforward.
+Every definition in Figure~\ref{fig:definitions} is \emph{local} (only represents a small part of the SRS structure) and therefore verification against its intended meaning is relatively straightforward.
 
@@
 
+\begin{figure}
+\caption{
+Definitions of auxiliary HOL functions used while proving theorems about the model of the SRS structure.
+\label{fig:auxdefs}
+}
+\begin{align*}
+\mathsf{delay\_sum}\;0&=0\\
+\mathsf{delay\_sum}\;n&=\mathsf{delay}\;n'+\mathsf{delay\_sum}\;n'\\
+&\qquad\text{where $(n',\_)=\mathsf{input}\;n$}
+\end{align*}
+\begin{align*}
+\mathsf{last\_update}\;n\;0&=0\\
+\mathsf{last\_update}\;n\;t&=\text{if $\mathsf{update\_time}\;n\;t$ then $t$}\\
+&\qquad\text{else }\mathsf{last\_update}\;n\;(t-1)\qquad\text{(where $t>0$)}
+\end{align*}
+\end{figure}
+
+\begin{multline*}
+\forall{D\,n\,x\,t}.\,{0 < n}\land{\mathsf{update\_time\;n\;t}}\implies\\\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)-1}\begin{cases}0&t<m+\mathsf{delay\_sum}\;n\\D\;(t-m-\mathsf{delay\_sum}\;n)&\text{otherwise}\end{cases}
+\end{multline*}
+
 %One of our objectives in formally modelling SRS was to keep the topology parameterized.

### 2011-03-26 fmics/paper.tex
merge with dropbox

@@
 The use of a formal method has compelled the stakeholders to clarify intricate details of the SRS structure and behaviour.
-
 \end{abstract}
@@
 The SRS component is implemented in hardware, in order to be fast enough to work in real-time, and on Field Programmable Gate Arrays (FPGAs) in particular so that they can be easily reprogrammed with future upgrades~\cite{Chris-FPGA}.
-%Multiple implementations of the SRS structure are installed regularly around the LHC as part of the BLMS.
 
@@
 The current approach for analyzing the SRS implementation is simulation of its behavior on sample streams of input~\cite{}.
-(one more sentence needs to be added on the current verification process)
+{\bf one more sentence needs to be added on the current verification process}
 
 In this paper, we describe a formal verification approach based on logical deduction for analyzing the SRS implementation.
-(need to add few sentences explaining in a very high level the verification approach/strategy)
+Our high level proof strategy takes advantage of the regular structure of the SRS.
+This structure consists of multiple layers of shift registers and some simple arithmetic hardware.
+There is a degree of regularity in how the output of each layer is used as input to the next layer of this structure.
+There is also a degree of regularity in the timing of each layer with respect to its position in the stack of layers.
+This regularity serves as the basis of using inductive reasoning, which allows us to keep the amount of verification effort linear with respect to the number of layers in the SRS structure even though the ``state space" of the possible values of the structure increases exponentially with the number of layers.
+{\bf Ramana, please verify if this is actually true, that is, we use induction}
 
@@
 Therefore, the discrepancy between the SRS sum and the exact sum needs to be small.
-
-%Simulation of the SRS on sample streams of input is one approach, currently in use~\cite{}, to building an understanding of the SRS's behaviour and confidence that the behaviour is correct.
-
-%Formal methods are another approach to building understanding and confidence.
-%The work on constructing proofs may be compared to the work on analysing the results of simulations.
-The work on formulating precise specifications is extra, but is arguably beneficial for both methods.
-<need to be more descriptive here i.e.
-to do number 2) below rather than 3)>
-
-%{\bf myNote: Add 1) the approach currently taken for the analysis of SRS, 2) the approach we took in this paper for the analysis, 3) the reasons behind that  the goals for this case study 4) the main contributions}
 
@@
 
-Table~\ref{tab:descriptions} lists the HOL functions we defined in the second phase along with their intended meanings.
-The arrangement of the slices is described by functions $\mathsf{input}\;n$ and $\mathsf{tap}\;n\;x$.
-The read delay of a slice is modeled by $\mathsf{delay}\;n$.
-Each slice is modeled by three functions: $\mathsf{SR}\;D\;n\;m\;t$ represents the value of each cell of the slice's shift register, $\mathsf{source}\;D\;n\;m\;t$ represents the direct input of each cell, and $\mathsf{output}\;D\;n\;x\;t$ represents the value of the slice's output at its taps.
-The function $\mathsf{RS}\;D\;n\;t$ models the running sums and $\mathsf{update\_time}\;n\;t$ checks if it is time to refresh the contents of a slice.
-We have both $\mathsf{RS}$ and $\mathsf{output}$ functions, though $\mathsf{RS}$ is easily defined in terms of $\mathsf{output}$, because $\mathsf{RS}$ represents an SRS output (indexed by a single number), whereas $\mathsf{output}$ represents an individual slice output (indexed by slice and tap numbers).
-By separating $\mathsf{RS}$ and $\mathsf{output}$, we allow for designs where some slice outputs are not SRS outputs, but may still be used internally as inputs to other slices.
-
-The formal definitions of the functions listed in Table~\ref{tab:descriptions} are given in Figure~\ref{fig:definitions}.
-Note that the definition of $\mathsf{input}$ when $n=0$ or when $n>6$ does not change the structure presented.
-A similar comment applies to other definitions when $n$, representing a slice number, is $0$, or when $x$, representing a shift register position, is greater than $1$.
-Every definition in Figure~\ref{fig:definitions} is \emph{local} (only represents a small part of the SRS structure) and therefore verification against its intended meaning is relatively straightforward.
-
 \begin{table}[t]
@@
 \end{table}
+
+Table~\ref{tab:descriptions} lists the HOL functions we defined in the second phase along with their intended meanings.
+The arrangement of the slices is described by functions $\mathsf{input}\;n$ and $\mathsf{tap}\;n\;x$.
+The read delay of a slice is modeled by $\mathsf{delay}\;n$.
+Each slice is modeled by three functions: $\mathsf{SR}\;D\;n\;m\;t$ represents the value of each cell of the slice's shift register, $\mathsf{source}\;D\;n\;m\;t$ represents the direct input of each cell, and $\mathsf{output}\;D\;n\;x\;t$ represents the value of the slice's output at its taps.
+The function $\mathsf{RS}\;D\;n\;t$ models the running sums and $\mathsf{update\_time}\;n\;t$ checks if it is time to refresh the contents of a slice.
+We have both $\mathsf{RS}$ and $\mathsf{output}$ functions, though $\mathsf{RS}$ is easily defined in terms of $\mathsf{output}$, because $\mathsf{RS}$ represents an SRS output (indexed by a single number), whereas $\mathsf{output}$ represents an individual slice output (indexed by slice and tap numbers).
+By separating $\mathsf{RS}$ and $\mathsf{output}$, we allow for designs where some slice outputs are not SRS outputs, but may still be used internally as inputs to other slices.
+
+The formal definitions of the functions listed in Table~\ref{tab:descriptions} are given in Figure~\ref{fig:definitions}.
+Note that the definition of $\mathsf{input}$ when $n=0$ or when $n>6$ does not change the structure presented.
+A similar comment applies to other definitions when $n$, representing a slice number, is $0$, or when $x$, representing a shift register position, is greater than $1$.
+We define all excess taps (where $x > 1$) to be in the same position as the last tap.
+
+%Some text description for Figure 7, we can remove it later on
+As explained in Section~\ref{sec-SRS}, the delay of each slice is equal to the delay of the slice it receives its input from multiplied by the number of the elements in the input slice used for the sum.
+For example in Figure~\ref{fig:srs}, $\mathsf{delay}\;4 = \mathsf{delay}\;3 \times \mathsf{tap}\;3\;0 = 2 \times 32$.
+The definition of the $\mathsf{source}$ function states that the source of each cell $m$ in a shift register is cell $m-1$, except for the first cell whose source is the output of the input slice.
+For example, $\mathsf{source}\;D\;4\;7\;t = \mathsf{SR}\;D\;4\;6\;t$ and $\mathsf{source}\;D\;4\;0\;t = \mathsf{output}\;D\;3\;0\;t$.
+The output of each slice is computed every time the contents of its shift register are updated by adding the incoming newest value (specified by $\mathsf{source}$) and subtracting its oldest value, that is the value in the cell at the tap position.
+The content of each cell of a shift register, $\mathsf{SR}\;D\;n\;m\;t$  is also computed at every update time based on the value of its source.
+Every definition in Figure~\ref{fig:definitions} is \emph{local} (only represents a small part of the SRS structure) and therefore verification against its intended meaning is relatively straightforward.
 

### 2011-03-27 fmics/paper.tex
Revise intro paragraph about inductive reasoning

@@
 In this paper, we describe a formal verification approach based on logical deduction for analyzing the SRS implementation.
-Our high level proof strategy takes advantage of the regular structure of the SRS.
-This structure consists of multiple layers of shift registers and some simple arithmetic hardware.
-There is a degree of regularity in how the output of each layer is used as input to the next layer of this structure.
+Our high level proof strategy takes advantage of the regular structure of the SRS, which consists of multiple layers of shift registers and some simple arithmetic hardware.
+There is a degree of regularity in how the output of each layer is used as input to the next layer.
 There is also a degree of regularity in the timing of each layer with respect to its position in the stack of layers.
-This regularity serves as the basis of using inductive reasoning, which allows us to keep the amount of verification effort linear with respect to the number of layers in the SRS structure even though the ``state space" of the possible values of the structure increases exponentially with the number of layers.
-{\bf Ramana, please verify if this is actually true, that is, we use induction}
+This regularity serves as a basis for inductive reasoning, which makes the amount of verification effort impervious to the number of layers in the structure.
 

### 2011-03-27 fmics/paper.tex
questions about specification

@@
 Our goal is to build a generic model of the SRS structure, and to prove that it satisfies its specification, that it calculates approximate running sums of received count values within acceptable error margins.
-Let $\mathsf{RS}\;n$ denote, as in Figure~\ref{fig:srs}, the $n$-th output of the SRS structure, which is supposed to compute a sum of received count values with some delay, and let $\mathsf{true\_sum}\;n$ denote this sum.
+Let $\mathsf{RS}\;n$ denote, as in Figure~\ref{fig:srs}, the $n$-th output of the SRS structure, which is supposed {\bf really?
+supposed to have a delay?} to compute a sum of received count values with some delay, and let $\mathsf{true\_sum}\;n$ denote this sum {\bf ``this'' being the delayed one, or the non-delayed one?}.
 A sketch of the desired correctness statement is \[\forall{n} \cdot \, \mathsf{RS}\;n= \mathsf{true\_sum}\; n \pm\text{acceptable error}\]

### 2011-03-27 fmics/paper.tex
remove 2 phases/simplified model from section 4

@@
 The first natural step to prove the correctness statement is to build a logical model of the SRS structure.
-Our work proceeded in two phases, starting with a model of a simplified structure and proceeding to a model which is closer to the SRS implementation in BLMS.
-In both phases, we model each building block of the SRS structure that holds a count value -- for example, each cell in each shift register -- as a function in HOL that takes the input stream and current time step as inputs and returns the value the corresponding building block holds at that time.
+We model each building block of the SRS structure that holds a count value -- for example, each cell in each shift register -- as a function in HOL that takes the input stream and current time step as inputs and returns the value the corresponding building block holds at that time.
 
-Our simplifications in the first phase include:
-\begin{enumerate}
-\item Masking the middle taps in each slice,
-\item Considering a more regular arrangement of slices than the SRS implementation in BLMS,
-\item Masking the subtractors and the accumulators in each slice, instead, defining the output of a slice as the sum of the values in its cells directly,
-\item Simplifying timing by ignoring some of the delays that each building block may introduce, {\bf this is not true - there were delays in the simplified model}
-\item Using natural numbers, rather than finite words, for count values.
-\end{enumerate}
-
-Building a logical model for simplified structure of the SRS helped us improve our understanding of the behavior of the SRS.
-{\bf what else did we learn through this process? If there was no particular advantage, I don't think we should mention the simplified version.
-(I don't think there was any particular advantage -- Ramana)}.
-
-In the second phase, we eliminated the first four simplifications (listed above) by adding middle taps, a more flexible arrangement of slices, the subtractor and accumulator, and considering the delays that each block may introduce.
-In both cases, however, the model is simultaneously a simplification and a generalization of the actual structure of SRS in BLMS in the following way: we model an infinite number of slices (rather than six slices), each with an infinite number of shift register cells and taps (rather than fixed width shift registers and only two taps), simply by letting indices range over the natural numbers without explicitly giving limits.
-This parameterization makes the model and the analysis more likely to be applicable to future upgraded/modified versions of the system.
+Our model is both a simplification and a generalization of the actual structure of SRS in BLMS in the following sense.
+We model an infinite number of slices (rather than six slices), each with an infinite number of shift register cells and taps (rather than fixed width shift registers and only two taps), simply by letting indices range over the natural numbers without explicitly giving limits.
+This parameterization makes the model more likely to be applicable to future versions of the system, but also fits more naturally into HOL than would a finite model.
+We use natural numbers throughout, rather than, for example, finite words that would more accurately model count values.
 
@@
 
-Table~\ref{tab:descriptions} lists the HOL functions we defined in the second phase along with their intended meanings.
+Table~\ref{tab:descriptions} lists the HOL functions comprising our model along with their intended meanings.
 The arrangement of the slices is described by functions $\mathsf{input}\;n$ and $\mathsf{tap}\;n\;x$.

### 2011-03-27 fmics/paper.tex
add statements of 6 theorems
and names of 6 more that would probably be needed if I want to go
through any proofs...

@@
 \end{align*}
+\begin{align*}
+\mathsf{exact}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)}\begin{cases}0&t<m+1\\D\;(t-m-1)&\text{otherwise}\end{cases}
+\end{align*}
 \end{figure}
 
-\begin{multline*}
+\begin{figure}
+\caption{
+A selection of theorems about our model proved using HOL4.
+\label{fig:theorems}
+}
+\begin{equation}\label{update_time_last_update}
+\forall{n\,t}.\,\mathsf{update\_time}\;n\;(\mathsf{last\_update}\;n\;t)
+\end{equation}
+\begin{equation}\label{last_update_thm}
+\forall{n\,t}.\,\mathsf{last\_update}\;n\;t=t-t\operatorname{mod}\mathsf{delay}\;n
+\end{equation}
+%SR_prev
+%SR_first
+\begin{equation}\label{output_last_update}
+\forall{D\,n\,x\,t}.\,\mathsf{output}\;D\;n\;x\;t=\mathsf{output}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t)
+\end{equation}
+\begin{equation}\label{output_sum}
+\forall{D\,n\,x\,t}.\,{0<n}\implies\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\mathsf{SR}\;D\;n\;m\;t
+\end{equation}
+%output_first
+%output_source_at_update_times
+%output_prev_at_update_times
+\begin{multline}\label{output_input_at_update_times}
 \forall{D\,n\,x\,t}.\,{0 < n}\land{\mathsf{update\_time\;n\;t}}\implies\\\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)-1}\begin{cases}0&t<m+\mathsf{delay\_sum}\;n\\D\;(t-m-\mathsf{delay\_sum}\;n)&\text{otherwise}\end{cases}
-\end{multline*}
-
-%One of our objectives in formally modelling SRS was to keep the topology parameterized.
-%It is likely that the configuration of the slices and shift registers will in the future change to accommodate higher precision.
-%A parameterized model can be used both for current and future versions of the system.
-%For this, we formalized the definitions of the model by six parameters:
+\end{multline}
+\begin{multline}\label{output_eq_exact}
+\forall{D\,n\,x\,t}.\,{0 < n}\implies\\\mathsf{output}\;D\;n\;x\;t=\begin{cases}0&\mathsf{last\_update}\;n\;t<\mathsf{delay}\;n\\\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)\end{cases}
+\end{multline}
+%max_error
+\end{figure}
 

### 2011-03-27 fmics/paper.tex
Paragraph describing one of the main theorems
Not describing any proofs yet...

@@
 
+Figure~\ref{fig:theorems} gives a selection of theorems we have proved in HOL, some of which refer to additional constants defined in Figure~\ref{fig:auxdefs}.
+Our central result, Theorem~(\ref{output_input_at_update_times}), equates the output of a slice to a sum of consecutive input counts.
+More precisely, it says that if slice $n$ was just updated, then the output at tap $x$ is equal to the sum of $((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)$ input values, starting from the input $\mathsf{delay\_sum}\;n$ time steps ago.
+In light of Theorem~(\ref{output_last_update}), which says a slice's output stays constant between update times for the slice, Theorem~(\ref{output_input_at_update_times}) characterises all outputs at all times.
+Thus, the SRS structure's outputs are exact running sums of the input counts over windows whose sizes depend on the the width of the slice and position of taps.
+However, the sums are delayed by the total delay across all previous slices (represented by $\mathsf{delay\_sum}$).
+
 \begin{figure}

### 2011-03-28 fmics/paper.tex
merge changes from dropbox

@@
 \usepackage{times,graphicx,epsfig,boxedminipage,url,xspace,array,amsmath}
+\newcommand\T{\rule{0pt}{2.6ex}}
+\newcommand\B{\rule[-1.2ex]{0pt}{0pt}}
 
@@
 Our goal is to build a generic model of the SRS structure, and to prove that it satisfies its specification, that it calculates approximate running sums of received count values within acceptable error margins.
-Let $\mathsf{RS}\;n$ denote, as in Figure~\ref{fig:srs}, the $n$-th output of the SRS structure, which is supposed {\bf really?
-supposed to have a delay?} to compute a sum of received count values with some delay, and let $\mathsf{true\_sum}\;n$ denote this sum {\bf ``this'' being the delayed one, or the non-delayed one?}.
+Let $\mathsf{RS}\;n$ denote, as in Figure~\ref{fig:srs}, the $n$-th output of the SRS structure, which is supposed to compute a sum of received count values, and let $\mathsf{true\_sum}\;n$ denote this sum.
+The multi-layerd structure of the SRS and the read delay of each shift register result in the outputs of the SRS to be delayed from the $\mathsf{true\_sum}$ values.
 A sketch of the desired correctness statement is \[\forall{n} \cdot \, \mathsf{RS}\;n= \mathsf{true\_sum}\; n \pm\text{acceptable error}\]
@@
 Slice $0$ is a fake slice representing the SRS input; this enables a succinct definition of $\mathsf{source}$.
+\vspace{0.2in}
 \label{fig:definitions}}
-\begin{align*}
-\mathsf{tap}\;0\;0&=0&\mathsf{tap}\;0\;x&=0\\
-\mathsf{tap}\;1\;0&=1-1&\mathsf{tap}\;1\;x&=2-1\\
-\mathsf{tap}\;2\;0&=8-1&\mathsf{tap}\;2\;x&=16-1\\
-&\dots&\mathsf{tap}\;6\;x&=128-1\quad\text{(where each $x>0$)}\\\\
-\mathsf{input}\;0&=(0,0)&\mathsf{input}\;1&=(0,0)\\
-\mathsf{input}\;2&=(0,0)&\mathsf{input}\;3&=(1,1)\\
-\mathsf{input}\;4&=(3,0)&\mathsf{input}\;5&=(4,0)\\
-\mathsf{input}\;6&=(4,1)&\mathsf{input}\;n&=(n-1,0)\quad\text{(where $n>6$)}
-\end{align*}
-\begin{align*}
-\mathsf{delay}\;0&=1\\
-\mathsf{delay}\;n&=\mathsf{delay}\;n'\times((\mathsf{tap}\;n'\;x)+1)\\
-&\qquad\text{where $(n',x)=\mathsf{input\;n}$ (and $n>0$)}
-\end{align*}
-\begin{align*}
-\mathsf{update\_time}\;n\;t&\iff(t\operatorname{mod}\mathsf{delay}\;n=0)
-\end{align*}
-\begin{align*}
-\mathsf{source}\;D\;n\;0\;t&=\mathsf{output}\;D\;n'\;x\;t\quad\text{where $(n',x)=\mathsf{input\;n}$}\\
-\mathsf{source}\;D\;n\;m\;t&=\;\mathsf{SR}\;D\;n\;(m-1)\;\quad\text{(where $m>0$)}
-\end{align*}
-\begin{align*}
-\mathsf{output}\;D\;0\;x\;t&=D\;t\\
-\mathsf{output}\;D\;n\;x\;0&=0\\
-\mathsf{output}\;D\;n\;x\;t&=
-\text{if $\mathsf{update\_time}\;n\;t$}\\
-&\qquad\text{then }((\mathsf{output}\;D\;n\;x\;(t-1))+(\mathsf{source}\;D\;n\;0\;(t-1)))-(\mathsf{SR}\;D\;n(\mathsf{tap}\;n\;x)\;(t-1))\\
-&\qquad\text{else }\mathsf{output}\;D\;n\;x\;(t-1)\qquad\text{(where $t>0$)}
-\end{align*}
-\begin{align*}
-\mathsf{SR}\;D\;n\;m\;0&=0\\
-\mathsf{SR}\;D\;n\;m\;t&=\text{if $\mathsf{update\_time}\;n\;t$}\\
-&\qquad\text{then }\mathsf{source}\;D\;n\;m\;(t-1)\\
-&\qquad\text{else }\mathsf{SR}\;D\;n\;m\;t\qquad\text{(where $t>0$)}
-\end{align*}
-\begin{align*}
-\mathsf{RS}\;D\;n\;t&=\mathsf{output}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)
-\end{align*}
+\begin{tabular}{l|ll}
+Function \T \B & Definition \\\hline
+$\mathsf{tap}\;n\;x$ \T & $\;\mathsf{tap}\;0\;0=0\quad\quad$ & $\mathsf{tap}\;0\;x=0$\\
+&$\;\mathsf{tap}\;1\;0=1-1\quad$ &  $\mathsf{tap}\;1\;x=2-1$\\
+&$\;\mathsf{tap}\;2\;0=8-1\quad$ &  $\mathsf{tap}\;2\;x=16-1$\\
+&$\;\dots\quad$ \B & $\mathsf{tap}\;6\;x=128-1\quad\text{(where each $x>0$)}$\\\hline
+
+$\mathsf{input}\;n$ \T & $\;\mathsf{input}\;0=(0,0)\quad$ & $\mathsf{input}\;1=(0,0)$\\
+&$\;\mathsf{input}\;2=(0,0)\quad$ & $\mathsf{input}\;3=(1,1)$\\
+&$\;\mathsf{input}\;4=(3,0)\quad$ & $\mathsf{input}\;5=(4,0)$\\
+&$\;\mathsf{input}\;6=(4,1)\quad$ \B & $\mathsf{input}\;n=(n-1,0)\quad\text{(where $n>6$)}$\\\hline
+
+$\mathsf{delay}\;n$ \T & $\;\mathsf{delay}\;0=1$\\
+&\multicolumn{2}{|l}{$\;\mathsf{delay}\;n=\mathsf{delay}\;n'\times((\mathsf{tap}\;n'\;x)+1)$}\\
+&\multicolumn{2}{|l}{$\qquad\qquad\qquad\text{where $(n',x)=\mathsf{input\;n}$ (and $n>0$)}$} \B \\\hline
+
+$\mathsf{source}\;D\;n\;m\;t$ \T & \multicolumn{2}{|l}{\;$\mathsf{source}\;D\;n\;0\;t=\mathsf{output}\;D\;n'\;x\;t\quad\text{where $(n',x)=\mathsf{input\;n}$}$}\\
+&\multicolumn{2}{|l}{$\;\mathsf{source}\;D\;n\;m\;t=\;\mathsf{SR}\;D\;n\;(m-1)\;\quad\text{(where $m>0$)}$} \B \\\hline
+
+$\mathsf{SR}\;D\;n\;m\;t$ \T &  \multicolumn{2}{|l}{\;$\mathsf{SR}\;D\;n\;m\;0=0$}\\
+&\multicolumn{2}{|l}{$\;\mathsf{SR}\;D\;n\;m\;t=\text{if $\mathsf{update\_time}\;n\;t$}\;\;\;\text{then }$}\\
+&&$\mathsf{source}\;D\;n\;m\;(t-1)$\\
+&\multicolumn{2}{|l}{$\qquad\qquad\qquad\quad\text{else }\mathsf{SR}\;D\;n\;m\;t\qquad\text{(where $t>0$)}$}\B \\\hline
+
+$\mathsf{output}\;D\;n\;x\;t$ \T &\multicolumn{2}{|l}{$\;\mathsf{output}\;D\;0\;x\;t=D\;t$}\\
+&\multicolumn{2}{|l}{$\;\mathsf{output}\;D\;n\;x\;0=0$}\\
+&\multicolumn{2}{|l}{$\;\mathsf{output}\;D\;n\;x\;t = \text{if $\mathsf{update\_time}\;n\;t$}\;\;\;\text{then }$}\\
+&\multicolumn{2}{|l}{$\qquad\quad((\mathsf{output}\;D\;n\;x\;(t-1))+(\mathsf{source}\;D\;n\;0\;(t-1)))-$}\\
+&&$\qquad\qquad\qquad(\mathsf{SR}\;D\;n(\mathsf{tap}\;n\;x)\;(t-1))$\\
+&\multicolumn{2}{|l}{$\qquad\qquad\qquad\qquad\text{else }\mathsf{output}\;D\;n\;x\;(t-1)\qquad\text{(where $t>0$)}$}\B\\\hline
+
+$\mathsf{RS}\;D\;n\;t$ \T \B & \multicolumn{2}{|l}{$\;\mathsf{RS}\;D\;n\;t=\mathsf{output}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)$}\\\hline
+
+$\mathsf{update\_time}\;n\;t$ \T & \multicolumn{2}{|l}{$\;\mathsf{update\_time}\;n\;t\iff(t\operatorname{mod}\mathsf{delay}\;n=0)$}
+\end{tabular}
 \end{figure}
@@
 
+%One of our objectives in formally modelling SRS was to keep the topology parameterized.
+%It is likely that the configuration of the slices and shift registers will in the future change to accommodate higher precision.
+%A parameterized model can be used both for current and future versions of the system.
+%For this, we formalized the definitions of the model by six parameters:
+
+%Our work proceeded in two phases, starting with a model of a simplified structure and proceeding to a model which is closer to the SRS implementation in BLMS.
+%In both phases, we model each building block of the SRS structure that holds a count value -- for example, each cell in each shift register -- as a function in HOL that takes the input stream and current time step as inputs and returns the value the corresponding building block holds at that time.
+
+%Our simplifications in the first phase include:
+%\begin{enumerate}
+%\item Masking the middle taps in each slice,
+%\item Considering a more regular arrangement of slices than the SRS implementation in BLMS,
+%\item Masking the subtractors and the accumulators in each slice, instead, defining the output of a slice as the sum of the values in its cells directly,
+%\item Simplifying timing by ignoring some of the delays that each building block may introduce, {\bf this is not true - there were delays in the simplified model}
+%\item Using natural numbers, rather than finite words, for count values.
+%\end{enumerate}
+
+%Building a logical model for simplified structure of the SRS helped us improve our understanding of the behavior of the SRS.
+
+%In the second phase, we eliminated the first four simplifications (listed above) by adding middle taps, a more flexible arrangement of slices, the subtractor and accumulator, and considering the delays that each block may introduce.
+%In both cases, however, the model is simultaneously a simplification and a generalization of the actual structure of SRS in BLMS in the following way: we model an infinite number of slices (rather than six slices), each with an infinite number of shift register cells and taps (rather than fixed width shift registers and only two taps), simply by letting indices range over the natural numbers without explicitly giving limits.
+%This parameterization makes the model and the analysis more likely to be applicable to future upgraded/modified versions of the system.
+
 \section{Conclusions and Lessons Learned}

### 2011-03-28 fmics/paper.tex
remove lines from table
minor correction to definition of output
make fig 8 tabular too

@@
 \usepackage{times,graphicx,epsfig,boxedminipage,url,xspace,array,amsmath}
-\newcommand\T{\rule{0pt}{2.6ex}}
-\newcommand\B{\rule[-1.2ex]{0pt}{0pt}}
 
@@
 Function&Intended meaning\\
-\hline\\
+\hline
 \(\mathsf{tap}\;n\;x\)&The position of tap $x$ of slice $n$.\\
@@
 \label{fig:definitions}}
-\begin{tabular}{l|ll}
-Function \T \B & Definition \\\hline
-$\mathsf{tap}\;n\;x$ \T & $\;\mathsf{tap}\;0\;0=0\quad\quad$ & $\mathsf{tap}\;0\;x=0$\\
+\begin{tabular}{lll}
+Function & Definition\\\hline
+$\mathsf{tap}\;n\;x$ & $\;\mathsf{tap}\;0\;0=0\quad\quad$ & $\mathsf{tap}\;0\;x=0$\\
 &$\;\mathsf{tap}\;1\;0=1-1\quad$ &  $\mathsf{tap}\;1\;x=2-1$\\
 &$\;\mathsf{tap}\;2\;0=8-1\quad$ &  $\mathsf{tap}\;2\;x=16-1$\\
-&$\;\dots\quad$ \B & $\mathsf{tap}\;6\;x=128-1\quad\text{(where each $x>0$)}$\\\hline
+&$\;\dots\quad$ & $\mathsf{tap}\;6\;x=128-1\quad\text{(where each $x>0$)}$\\
 
-$\mathsf{input}\;n$ \T & $\;\mathsf{input}\;0=(0,0)\quad$ & $\mathsf{input}\;1=(0,0)$\\
+$\mathsf{input}\;n$ & $\;\mathsf{input}\;0=(0,0)\quad$ & $\mathsf{input}\;1=(0,0)$\\
 &$\;\mathsf{input}\;2=(0,0)\quad$ & $\mathsf{input}\;3=(1,1)$\\
 &$\;\mathsf{input}\;4=(3,0)\quad$ & $\mathsf{input}\;5=(4,0)$\\
-&$\;\mathsf{input}\;6=(4,1)\quad$ \B & $\mathsf{input}\;n=(n-1,0)\quad\text{(where $n>6$)}$\\\hline
+&$\;\mathsf{input}\;6=(4,1)\quad$ & $\mathsf{input}\;n=(n-1,0)\quad\text{(where $n>6$)}$\\
 
-$\mathsf{delay}\;n$ \T & $\;\mathsf{delay}\;0=1$\\
-&\multicolumn{2}{|l}{$\;\mathsf{delay}\;n=\mathsf{delay}\;n'\times((\mathsf{tap}\;n'\;x)+1)$}\\
-&\multicolumn{2}{|l}{$\qquad\qquad\qquad\text{where $(n',x)=\mathsf{input\;n}$ (and $n>0$)}$} \B \\\hline
+$\mathsf{delay}\;n$ & $\;\mathsf{delay}\;0=1$\\
+&\multicolumn{2}{l}{$\;\mathsf{delay}\;n=\mathsf{delay}\;n'\times((\mathsf{tap}\;n'\;x)+1)$}\\
+&\multicolumn{2}{l}{$\qquad\qquad\qquad\text{where $(n',x)=\mathsf{input\;n}$ (and $n>0$)}$}\\
 
-$\mathsf{source}\;D\;n\;m\;t$ \T & \multicolumn{2}{|l}{\;$\mathsf{source}\;D\;n\;0\;t=\mathsf{output}\;D\;n'\;x\;t\quad\text{where $(n',x)=\mathsf{input\;n}$}$}\\
-&\multicolumn{2}{|l}{$\;\mathsf{source}\;D\;n\;m\;t=\;\mathsf{SR}\;D\;n\;(m-1)\;\quad\text{(where $m>0$)}$} \B \\\hline
+$\mathsf{source}\;D\;n\;m\;t$ & \multicolumn{2}{l}{\;$\mathsf{source}\;D\;n\;0\;t=\mathsf{output}\;D\;n'\;x\;t\quad\text{where $(n',x)=\mathsf{input\;n}$}$}\\
+&\multicolumn{2}{l}{$\;\mathsf{source}\;D\;n\;m\;t=\;\mathsf{SR}\;D\;n\;(m-1)\;\quad\text{(where $m>0$)}$}\\
 
-$\mathsf{SR}\;D\;n\;m\;t$ \T &  \multicolumn{2}{|l}{\;$\mathsf{SR}\;D\;n\;m\;0=0$}\\
-&\multicolumn{2}{|l}{$\;\mathsf{SR}\;D\;n\;m\;t=\text{if $\mathsf{update\_time}\;n\;t$}\;\;\;\text{then }$}\\
+$\mathsf{SR}\;D\;n\;m\;t$ &  \multicolumn{2}{l}{\;$\mathsf{SR}\;D\;n\;m\;0=0$}\\
+&\multicolumn{2}{l}{$\;\mathsf{SR}\;D\;n\;m\;t=\text{if $\mathsf{update\_time}\;n\;t$}\;\;\;\text{then }$}\\
 &&$\mathsf{source}\;D\;n\;m\;(t-1)$\\
-&\multicolumn{2}{|l}{$\qquad\qquad\qquad\quad\text{else }\mathsf{SR}\;D\;n\;m\;t\qquad\text{(where $t>0$)}$}\B \\\hline
+&\multicolumn{2}{l}{$\qquad\qquad\qquad\quad\text{else }\mathsf{SR}\;D\;n\;m\;t\qquad\text{(where $t>0$)}$}\\
 
-$\mathsf{output}\;D\;n\;x\;t$ \T &\multicolumn{2}{|l}{$\;\mathsf{output}\;D\;0\;x\;t=D\;t$}\\
-&\multicolumn{2}{|l}{$\;\mathsf{output}\;D\;n\;x\;0=0$}\\
-&\multicolumn{2}{|l}{$\;\mathsf{output}\;D\;n\;x\;t = \text{if $\mathsf{update\_time}\;n\;t$}\;\;\;\text{then }$}\\
-&\multicolumn{2}{|l}{$\qquad\quad((\mathsf{output}\;D\;n\;x\;(t-1))+(\mathsf{source}\;D\;n\;0\;(t-1)))-$}\\
+$\mathsf{output}\;D\;n\;x\;t$ &\multicolumn{2}{l}{$\;\mathsf{output}\;D\;0\;x\;t=D\;t$}\\
+&\multicolumn{2}{l}{$\;\mathsf{output}\;D\;n\;x\;0=0$}\\
+&\multicolumn{2}{l}{$\;\mathsf{output}\;D\;n\;x\;t = \text{if $\mathsf{update\_time}\;n\;t$}\;\;\;\text{then }$}\\
+&\multicolumn{2}{l}{$\qquad\quad((\mathsf{output}\;D\;n\;x\;(t-1))+(\mathsf{source}\;D\;n\;0\;(t-1)))-$}\\
 &&$\qquad\qquad\qquad(\mathsf{SR}\;D\;n(\mathsf{tap}\;n\;x)\;(t-1))$\\
-&\multicolumn{2}{|l}{$\qquad\qquad\qquad\qquad\text{else }\mathsf{output}\;D\;n\;x\;(t-1)\qquad\text{(where $t>0$)}$}\B\\\hline
+&\multicolumn{2}{l}{$\qquad\qquad\qquad\qquad\text{else }\mathsf{output}\;D\;n\;x\;(t-1)\qquad\text{(where $n,t>0$)}$}\\
 
-$\mathsf{RS}\;D\;n\;t$ \T \B & \multicolumn{2}{|l}{$\;\mathsf{RS}\;D\;n\;t=\mathsf{output}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)$}\\\hline
+$\mathsf{RS}\;D\;n\;t$ & \multicolumn{2}{l}{$\;\mathsf{RS}\;D\;n\;t=\mathsf{output}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)$}\\
 
-$\mathsf{update\_time}\;n\;t$ \T & \multicolumn{2}{|l}{$\;\mathsf{update\_time}\;n\;t\iff(t\operatorname{mod}\mathsf{delay}\;n=0)$}
+$\mathsf{update\_time}\;n\;t$ & \multicolumn{2}{l}{$\;\mathsf{update\_time}\;n\;t\iff(t\operatorname{mod}\mathsf{delay}\;n=0)$}
 \end{tabular}
@@
 }
-\begin{align*}
-\mathsf{delay\_sum}\;0&=0\\
-\mathsf{delay\_sum}\;n&=\mathsf{delay}\;n'+\mathsf{delay\_sum}\;n'\\
-&\qquad\text{where $(n',\_)=\mathsf{input}\;n$}
-\end{align*}
-\begin{align*}
-\mathsf{last\_update}\;n\;0&=0\\
-\mathsf{last\_update}\;n\;t&=\text{if $\mathsf{update\_time}\;n\;t$ then $t$}\\
-&\qquad\text{else }\mathsf{last\_update}\;n\;(t-1)\qquad\text{(where $t>0$)}
-\end{align*}
-\begin{align*}
-\mathsf{exact}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)}\begin{cases}0&t<m+1\\D\;(t-m-1)&\text{otherwise}\end{cases}
-\end{align*}
+\begin{tabular}{ll}
+Function & Definition\\\hline
+$\mathsf{delay\_sum}\;n$
+&$\mathsf{delay\_sum}\;0=0$\\
+&$\mathsf{delay\_sum}\;n=\mathsf{delay}\;n'+\mathsf{delay\_sum}\;n'$\\
+&$\qquad\text{where $(n',\_)=\mathsf{input}\;n$}$\\
+
+$\mathsf{last\_update}\;n$
+&$\mathsf{last\_update}\;n\;0=0$\\
+&$\mathsf{last\_update}\;n\;t=\text{if $\mathsf{update\_time}\;n\;t$ then $t$}$\\
+&$\qquad\text{else }\mathsf{last\_update}\;n\;(t-1)\qquad\text{(where $t>0$)}$\\
+
+$\mathsf{exact}\;D\;n\;x\;t$
+&$\mathsf{exact}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)}\begin{cases}0&t<m+1\\D\;(t-m-1)&\text{otherwise}\end{cases}$
+\end{tabular}
 \end{figure}

### 2011-03-28 fmics/paper.tex
remove left column from figs 7 and 8
tweak some spacings in the tables
A (relatively unreadable) proof sketch for the paper

@@
 Slice $0$ is a fake slice representing the SRS input; this enables a succinct definition of $\mathsf{source}$.
-\vspace{0.2in}
+\vspace{2ex}
 \label{fig:definitions}}
-\begin{tabular}{lll}
-Function & Definition\\\hline
-$\mathsf{tap}\;n\;x$ & $\;\mathsf{tap}\;0\;0=0\quad\quad$ & $\mathsf{tap}\;0\;x=0$\\
-&$\;\mathsf{tap}\;1\;0=1-1\quad$ &  $\mathsf{tap}\;1\;x=2-1$\\
-&$\;\mathsf{tap}\;2\;0=8-1\quad$ &  $\mathsf{tap}\;2\;x=16-1$\\
-&$\;\dots\quad$ & $\mathsf{tap}\;6\;x=128-1\quad\text{(where each $x>0$)}$\\
+\begin{tabular}{ll}
+$\;\mathsf{tap}\;0\;0=0\quad\quad$ & $\mathsf{tap}\;0\;x=0$\\
+$\;\mathsf{tap}\;1\;0=1-1\quad$ &  $\mathsf{tap}\;1\;x=2-1$\\
+$\;\mathsf{tap}\;2\;0=8-1\quad$ &  $\mathsf{tap}\;2\;x=16-1$\\
+$\;\dots\quad$ & $\mathsf{tap}\;6\;x=128-1\quad\text{(where each $x>0$)}$\\[1ex]
 
-$\mathsf{input}\;n$ & $\;\mathsf{input}\;0=(0,0)\quad$ & $\mathsf{input}\;1=(0,0)$\\
-&$\;\mathsf{input}\;2=(0,0)\quad$ & $\mathsf{input}\;3=(1,1)$\\
-&$\;\mathsf{input}\;4=(3,0)\quad$ & $\mathsf{input}\;5=(4,0)$\\
-&$\;\mathsf{input}\;6=(4,1)\quad$ & $\mathsf{input}\;n=(n-1,0)\quad\text{(where $n>6$)}$\\
+$\;\mathsf{input}\;0=(0,0)\quad$ & $\mathsf{input}\;1=(0,0)$\\
+$\;\mathsf{input}\;2=(0,0)\quad$ & $\mathsf{input}\;3=(1,1)$\\
+$\;\mathsf{input}\;4=(3,0)\quad$ & $\mathsf{input}\;5=(4,0)$\\
+$\;\mathsf{input}\;6=(4,1)\quad$ & $\mathsf{input}\;n=(n-1,0)\quad\text{(where $n>6$)}$\\[1ex]
 
-$\mathsf{delay}\;n$ & $\;\mathsf{delay}\;0=1$\\
-&\multicolumn{2}{l}{$\;\mathsf{delay}\;n=\mathsf{delay}\;n'\times((\mathsf{tap}\;n'\;x)+1)$}\\
-&\multicolumn{2}{l}{$\qquad\qquad\qquad\text{where $(n',x)=\mathsf{input\;n}$ (and $n>0$)}$}\\
+$\;\mathsf{delay}\;0=1$\\
+\multicolumn{2}{l}{$\;\mathsf{delay}\;n=\mathsf{delay}\;n'\times((\mathsf{tap}\;n'\;x)+1)$}\\
+\multicolumn{2}{l}{$\qquad\qquad\qquad\text{where $(n',x)=\mathsf{input\;n}$ (and $n>0$)}$}\\[1ex]
 
-$\mathsf{source}\;D\;n\;m\;t$ & \multicolumn{2}{l}{\;$\mathsf{source}\;D\;n\;0\;t=\mathsf{output}\;D\;n'\;x\;t\quad\text{where $(n',x)=\mathsf{input\;n}$}$}\\
-&\multicolumn{2}{l}{$\;\mathsf{source}\;D\;n\;m\;t=\;\mathsf{SR}\;D\;n\;(m-1)\;\quad\text{(where $m>0$)}$}\\
+\multicolumn{2}{l}{\;$\mathsf{source}\;D\;n\;0\;t=\mathsf{output}\;D\;n'\;x\;t\quad\text{where $(n',x)=\mathsf{input\;n}$}$}\\
+\multicolumn{2}{l}{$\;\mathsf{source}\;D\;n\;m\;t=\;\mathsf{SR}\;D\;n\;(m-1)\;\quad\text{(where $m>0$)}$}\\[1ex]
 
-$\mathsf{SR}\;D\;n\;m\;t$ &  \multicolumn{2}{l}{\;$\mathsf{SR}\;D\;n\;m\;0=0$}\\
-&\multicolumn{2}{l}{$\;\mathsf{SR}\;D\;n\;m\;t=\text{if $\mathsf{update\_time}\;n\;t$}\;\;\;\text{then }$}\\
-&&$\mathsf{source}\;D\;n\;m\;(t-1)$\\
-&\multicolumn{2}{l}{$\qquad\qquad\qquad\quad\text{else }\mathsf{SR}\;D\;n\;m\;t\qquad\text{(where $t>0$)}$}\\
+\multicolumn{2}{l}{\;$\mathsf{SR}\;D\;n\;m\;0=0$}\\
+\multicolumn{2}{l}{$\;\mathsf{SR}\;D\;n\;m\;t=\text{if $\mathsf{update\_time}\;n\;t$ then }$}\\
+&$\mathsf{source}\;D\;n\;m\;(t-1)$\\
+\multicolumn{2}{l}{$\qquad\qquad\qquad\quad\text{else }\mathsf{SR}\;D\;n\;m\;t\qquad\text{(where $t>0$)}$}\\[1ex]
 
-$\mathsf{output}\;D\;n\;x\;t$ &\multicolumn{2}{l}{$\;\mathsf{output}\;D\;0\;x\;t=D\;t$}\\
-&\multicolumn{2}{l}{$\;\mathsf{output}\;D\;n\;x\;0=0$}\\
-&\multicolumn{2}{l}{$\;\mathsf{output}\;D\;n\;x\;t = \text{if $\mathsf{update\_time}\;n\;t$}\;\;\;\text{then }$}\\
-&\multicolumn{2}{l}{$\qquad\quad((\mathsf{output}\;D\;n\;x\;(t-1))+(\mathsf{source}\;D\;n\;0\;(t-1)))-$}\\
-&&$\qquad\qquad\qquad(\mathsf{SR}\;D\;n(\mathsf{tap}\;n\;x)\;(t-1))$\\
-&\multicolumn{2}{l}{$\qquad\qquad\qquad\qquad\text{else }\mathsf{output}\;D\;n\;x\;(t-1)\qquad\text{(where $n,t>0$)}$}\\
+\multicolumn{2}{l}{$\;\mathsf{output}\;D\;0\;x\;t=D\;t$}\\
+\multicolumn{2}{l}{$\;\mathsf{output}\;D\;n\;x\;0=0$}\\
+\multicolumn{2}{l}{$\;\mathsf{output}\;D\;n\;x\;t = \text{if $\mathsf{update\_time}\;n\;t$ then }$}\\
+\multicolumn{2}{l}{$\qquad\quad((\mathsf{output}\;D\;n\;x\;(t-1))+(\mathsf{source}\;D\;n\;0\;(t-1)))-{}$}\\
+&$\qquad\qquad\qquad(\mathsf{SR}\;D\;n\;(\mathsf{tap}\;n\;x)\;(t-1))$\\
+\multicolumn{2}{l}{$\qquad\qquad\qquad\qquad\text{else }\mathsf{output}\;D\;n\;x\;(t-1)\qquad\text{(where $n,t>0$)}$}\\[1ex]
 
-$\mathsf{RS}\;D\;n\;t$ & \multicolumn{2}{l}{$\;\mathsf{RS}\;D\;n\;t=\mathsf{output}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)$}\\
+\multicolumn{2}{l}{$\;\mathsf{RS}\;D\;n\;t=\mathsf{output}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)$}\\[1ex]
 
-$\mathsf{update\_time}\;n\;t$ & \multicolumn{2}{l}{$\;\mathsf{update\_time}\;n\;t\iff(t\operatorname{mod}\mathsf{delay}\;n=0)$}
+\multicolumn{2}{l}{$\;\mathsf{update\_time}\;n\;t\iff(t\operatorname{mod}\mathsf{delay}\;n=0)$}
 \end{tabular}
@@
 }
-\begin{tabular}{ll}
-Function & Definition\\\hline
-$\mathsf{delay\_sum}\;n$
-&$\mathsf{delay\_sum}\;0=0$\\
-&$\mathsf{delay\_sum}\;n=\mathsf{delay}\;n'+\mathsf{delay\_sum}\;n'$\\
-&$\qquad\text{where $(n',\_)=\mathsf{input}\;n$}$\\
+\vspace{2ex}
+\begin{tabular}{l}
+$\mathsf{delay\_sum}\;0=0$\\
+$\mathsf{delay\_sum}\;n=\mathsf{delay}\;n'+\mathsf{delay\_sum}\;n'\quad\text{where $(n',\_)=\mathsf{input}\;n$}$\\[1ex]
 
-$\mathsf{last\_update}\;n$
-&$\mathsf{last\_update}\;n\;0=0$\\
-&$\mathsf{last\_update}\;n\;t=\text{if $\mathsf{update\_time}\;n\;t$ then $t$}$\\
-&$\qquad\text{else }\mathsf{last\_update}\;n\;(t-1)\qquad\text{(where $t>0$)}$\\
+$\mathsf{last\_update}\;n\;0=0$\\
+$\mathsf{last\_update}\;n\;t=\text{if $\mathsf{update\_time}\;n\;t$ then $t$}$\\
+$\quad\;\;\qquad\qquad\qquad\text{else }\mathsf{last\_update}\;n\;(t-1)\quad\text{(where $t>0$)}$\\[1ex]
 
-$\mathsf{exact}\;D\;n\;x\;t$
-&$\mathsf{exact}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)}\begin{cases}0&t<m+1\\D\;(t-m-1)&\text{otherwise}\end{cases}$
+$\mathsf{exact}\;D\;n\;x\;t=\displaystyle\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)}\begin{cases}0&t<m+1\\D\;(t-m-1)&\text{otherwise}\end{cases}$
 \end{tabular}
@@
 
+To prove Theorem~(\ref{output_input_at_update_times}), we first make the following observations, for sufficiently large $t$:
+\begin{enumerate}
+\item
+$\mathsf{SR}\;D\;n\;(m+1)\;t=\mathsf{SR}\;D\;n\;m\;(t-\mathsf{delay}\;n)$, since shift register cells get their values from the previous cell, but only update every $\mathsf{delay}\;n$ time steps;
+\item
+$\mathsf{SR}\;D\;n\;m\;t=\mathsf{SR}\;D\;n\;0\;(t-(m\times\mathsf{delay}\;n))$, by induction on $m$ using the result above;
+\item
+$\mathsf{output}\;D\;n\;x\;t$ is the sum of values of consecutive shift register cells (Theorem~(\ref{output_sum})), by induction on $t$;
+\item
+Combining the last two results, $\mathsf{output\;D\;n\;x\;t}$ is the sum of consecutive values of the first shift register cell, $\mathsf{SR\;D\;n\;0}$;
+\item
+<Thus Theorem~(\ref{output_prev_at_update_times}), because the first shift register cell gets its value from the output of a previous slice>
+\item
+<Now that we have expressed one output in terms of a previous output, Theorem~(\ref{output_input_at_update_times}) can be proved by induction on $n$ (but this is not trivial - need to check what else apart from $\ref{output_prev_at_update_times}$ is really needed for the inductive step)>
+\end{enumerate}
+
 \begin{figure}
@@
 }
-\begin{equation}\label{update_time_last_update}
-\forall{n\,t}.\,\mathsf{update\_time}\;n\;(\mathsf{last\_update}\;n\;t)
-\end{equation}
-\begin{equation}\label{last_update_thm}
-\forall{n\,t}.\,\mathsf{last\_update}\;n\;t=t-t\operatorname{mod}\mathsf{delay}\;n
-\end{equation}
+\begin{align}
+\label{update_time_last_update}
+&\forall{n\,t}.\,\mathsf{update\_time}\;n\;(\mathsf{last\_update}\;n\;t)\\
+%\label{last_update_thm}
+%&\forall{n\,t}.\,\mathsf{last\_update}\;n\;t=t-t\operatorname{mod}\mathsf{delay}\;n\\
 %SR_prev
 %SR_first
-\begin{equation}\label{output_last_update}
-\forall{D\,n\,x\,t}.\,\mathsf{output}\;D\;n\;x\;t=\mathsf{output}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t)
-\end{equation}
-\begin{equation}\label{output_sum}
-\forall{D\,n\,x\,t}.\,{0<n}\implies\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\mathsf{SR}\;D\;n\;m\;t
-\end{equation}
-%output_first
-%output_source_at_update_times
-%output_prev_at_update_times
+\label{output_last_update}
+&\forall{D\,n\,x\,t}.\,\mathsf{output}\;D\;n\;x\;t=\mathsf{output}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t)\\
+\label{output_sum}
+&\forall{D\,n\,x\,t}.\,{0<n}\implies\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\mathsf{SR}\;D\;n\;m\;t
+\end{align}
+%\begin{multline}
+%\label{output_source_at_update_times}
+%\forall{D\,n\,x\,t}.\,{0<n}\land{\mathsf{update\_time\;n\;t}}\implies\\
+%\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\begin{cases}0&t\leq m\times(\mathsf{delay}\;n)\\\mathsf{source}\;D\;n\;0\;(t-m\times(\mathsf{delay\;n})-1)&\text{otherwise}\end{cases}
+%\end{multline}
+\begin{multline}
+\label{output_prev_at_update_times}
+\forall{D\,n\,x\,t}.\,{0<n}\land{\mathsf{update\_time}\;n\;t}\implies\\
+\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\begin{cases}0&t < m\times\mathsf{delay}\;n+\mathsf{delay}\;n'\\\mathsf{output}\;D\;n'\;x'\;(t-m\times\mathsf{delay}\;n-\mathsf{delay}\;n')&\text{otherwise}\end{cases}\\
+\qquad\text{where $(n',x')=\mathsf{input}\;n$}
+\end{multline}
 \begin{multline}\label{output_input_at_update_times}

### 2011-03-28 fmics/paper.tex
improve/fill in proof paragraph

@@
 
-To prove Theorem~(\ref{output_input_at_update_times}), we first make the following observations, for sufficiently large $t$:
+To prove Theorem~(\ref{output_input_at_update_times}), we first make the following observations\footnote{The precise statements of these observations are complicated by details of what happens when $t$ is low, or equivalently, when we are looking at shift register cells that have not yet received any input.}
 \begin{enumerate}
@@
 \item
-$\mathsf{output}\;D\;n\;x\;t$ is the sum of values of consecutive shift register cells (Theorem~(\ref{output_sum})), by induction on $t$;
+$\mathsf{output}\;D\;n\;x\;t$ is a sum of values of consecutive shift register cells (Theorem~(\ref{output_sum})), which can be proved by induction on $t$;
 \item
-Combining the last two results, $\mathsf{output\;D\;n\;x\;t}$ is the sum of consecutive values of the first shift register cell, $\mathsf{SR\;D\;n\;0}$;
+Combining the last two results, $\mathsf{output}\;D\;n\;x\;t$ is a sum of consecutive values of the first shift register cell, $\mathsf{SR}\;D\;n\;0$;
 \item
-<Thus Theorem~(\ref{output_prev_at_update_times}), because the first shift register cell gets its value from the output of a previous slice>
+Thus, we can express $\mathsf{output}\;D\;n\;x\;t$ in terms of $\mathsf{output}\;D\;n'\;x'\;t$, where $(n',x')=\mathsf{input}\;n$, since the source of $\mathsf{SR}\;D\;n\;0$ is the previous (input) slice's output;
 \item
-<Now that we have expressed one output in terms of a previous output, Theorem~(\ref{output_input_at_update_times}) can be proved by induction on $n$ (but this is not trivial - need to check what else apart from $\ref{output_prev_at_update_times}$ is really needed for the inductive step)>
+Finally, by induction on $n$, we can express $\mathsf{output}\;D\;n\;x\;t$ as a function of $D$ alone, as in Theorem~(\ref{output_input_at_update_times}), using the result above for the inductive case.
 \end{enumerate}

### 2011-03-28 fmics/paper.tex
merge some dropbox changes

@@
 
-Figure~\ref{fig:theorems} gives a selection of theorems we have proved in HOL, some of which refer to additional constants defined in Figure~\ref{fig:auxdefs}.
-Our central result, Theorem~(\ref{output_input_at_update_times}), equates the output of a slice to a sum of consecutive input counts.
+To prove Theorem~(\ref{output_input_at_update_times}), we first make the following observations\footnote{The precise statements of these observations are complicated by details of what happens when $t$ is low, or equivalently, when we are looking at shift register cells that have not yet received any input.}
+%Figure~\ref{fig:theorems} gives a selection of theorems we have proved in HOL, some of which refer to additional constants defined in Figure~\ref{fig:auxdefs}.
+%Our central result, Theorem~(\ref{output_input_at_update_times}), equates the output of a slice to a sum of consecutive input counts.
+Our central result equates the output of a slice to a sum of consecutive input counts.
 More precisely, it says that if slice $n$ was just updated, then the output at tap $x$ is equal to the sum of $((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)$ input values, starting from the input $\mathsf{delay\_sum}\;n$ time steps ago.
-In light of Theorem~(\ref{output_last_update}), which says a slice's output stays constant between update times for the slice, Theorem~(\ref{output_input_at_update_times}) characterises all outputs at all times.
-Thus, the SRS structure's outputs are exact running sums of the input counts over windows whose sizes depend on the the width of the slice and position of taps.
-However, the sums are delayed by the total delay across all previous slices (represented by $\mathsf{delay\_sum}$).
+\begin{theorem}
 
-To prove Theorem~(\ref{output_input_at_update_times}), we first make the following observations\footnote{The precise statements of these observations are complicated by details of what happens when $t$ is low, or equivalently, when we are looking at shift register cells that have not yet received any input.}
+The output of slice n at update time t ... (statement of the theorem)
+\begin{multline}\label{output_input_at_update_times}
+\forall{D\,n\,x\,t}.\,{0 < n}\land{\mathsf{update\_time\;n\;t}}\implies\\\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)-1}\begin{cases}0&t<m+\mathsf{delay\_sum}\;n\\D\;(t-m-\mathsf{delay\_sum}\;n)&\text{otherwise}\end{cases}
+\end{multline}
+\end{theorem}
+
+\begin{proof}
 \begin{enumerate}
@@
 \end{enumerate}
+
+\end{proof}
+
+In light of Theorem~(\ref{output_last_update}), which says a slice's output stays constant between update times for the slice, Theorem~(\ref{output_input_at_update_times}) characterises all outputs at all times.
+Thus, the SRS structure's outputs are exact running sums of the input counts over windows whose sizes depend on the the width of the slice and position of taps.
+However, the sums are delayed by the total delay across all previous slices (represented by $\mathsf{delay\_sum}$).
 

### 2011-03-28 fmics/paper.tex
incorporate Naghmeh's updates

@@
 
-To prove Theorem~(\ref{output_input_at_update_times}), we first make the following observations\footnote{The precise statements of these observations are complicated by details of what happens when $t$ is low, or equivalently, when we are looking at shift register cells that have not yet received any input.}
 %Figure~\ref{fig:theorems} gives a selection of theorems we have proved in HOL, some of which refer to additional constants defined in Figure~\ref{fig:auxdefs}.
+Figure~\ref{fig:auxdefs} defines a set of additional functions required to prove our main results.
+The function $\mathsf{delay\_sum}\;n$ represents the accumulative delay of the preceding shift registers of a slice.
+For example, $\mathsf{delay\_sum}\;4 = \mathsf{delay}\;3 + \mathsf{delay}\;1 + \mathsf{delay}\;0 = 2 +1 +1 = 4$.
+The function $\mathsf{last\_update}\;n\;t$ specifies the time that the $n$-th slice last updated and the $\mathsf{exact}\;D\;n\;x\;t$ {\bf Ramana, can say in simple words what exact represents?}
+
 %Our central result, Theorem~(\ref{output_input_at_update_times}), equates the output of a slice to a sum of consecutive input counts.
@@
 More precisely, it says that if slice $n$ was just updated, then the output at tap $x$ is equal to the sum of $((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)$ input values, starting from the input $\mathsf{delay\_sum}\;n$ time steps ago.
+Note that the precise statement of such a theorem is complicated by details of the value of the $\mathsf{output}$ at times when the shift registers have not yet received any values.
+
+%To prove Theorem~(\ref{output_input_at_update_times}), we first make the following observations\footnote{The precise statements of these observations are complicated by details of what happens when $t$ is low, or equivalently, when we are looking at shift register cells that have not yet received any input.}
+
 \begin{theorem}
-
-The output of slice n at update time t ... (statement of the theorem)
-\begin{multline}\label{output_input_at_update_times}
-\forall{D\,n\,x\,t}.\,{0 < n}\land{\mathsf{update\_time\;n\;t}}\implies\\\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)-1}\begin{cases}0&t<m+\mathsf{delay\_sum}\;n\\D\;(t-m-\mathsf{delay\_sum}\;n)&\text{otherwise}\end{cases}
-\end{multline}
+For all values of $D$, $n$, and $x$, the $\mathsf{output}$ of slice $n$ at update time $t$ is equal to
+\begin{equation*}%\label{output_input_at_update_times}
+%\forall{D\,n\,x\,t}.\,{0 < n}\land{\mathsf{update\_time\;n\;t}}\implies\\
+\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)-1}\begin{cases}0 \qquad\qquad\qquad  t<m+\mathsf{delay\_sum}\;n\\D\;(t-m-\mathsf{delay\_sum}\;n)\quad\text{otherwise}\end{cases}
+\end{equation*}
 \end{theorem}
@@
 \begin{proof}
-\begin{enumerate}
-\item
-$\mathsf{SR}\;D\;n\;(m+1)\;t=\mathsf{SR}\;D\;n\;m\;(t-\mathsf{delay}\;n)$, since shift register cells get their values from the previous cell, but only update every $\mathsf{delay}\;n$ time steps;
-\item
-$\mathsf{SR}\;D\;n\;m\;t=\mathsf{SR}\;D\;n\;0\;(t-(m\times\mathsf{delay}\;n))$, by induction on $m$ using the result above;
-\item
-$\mathsf{output}\;D\;n\;x\;t$ is a sum of values of consecutive shift register cells (Theorem~(\ref{output_sum})), which can be proved by induction on $t$;
-\item
-Combining the last two results, $\mathsf{output}\;D\;n\;x\;t$ is a sum of consecutive values of the first shift register cell, $\mathsf{SR}\;D\;n\;0$;
-\item
-Thus, we can express $\mathsf{output}\;D\;n\;x\;t$ in terms of $\mathsf{output}\;D\;n'\;x'\;t$, where $(n',x')=\mathsf{input}\;n$, since the source of $\mathsf{SR}\;D\;n\;0$ is the previous (input) slice's output;
-\item
+The shift register cells get their values from the previous cell but they only update every $\mathsf{delay}\;n$ time steps. Therefore,
+\begin{equation*}
+\mathsf{SR}\;D\;n\;(m+1)\;t=\mathsf{SR}\;D\;n\;m\;(t-\mathsf{delay}\;n)
+\end{equation*}
+By induction on $m$ on the above,
+\begin{equation*}
+\mathsf{SR}\;D\;n\;m\;t=\mathsf{SR}\;D\;n\;0\;(t-(m\times\mathsf{delay}\;n))
+\end{equation*}
+By induction on $t$, we can show that the $\mathsf{output}\;D\;n\;x\;t$ is a sum of values of consecutive shift register cells
+\begin{equation*}
+\forall{D\,n\,x\,t}.\,{0<n}\implies\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\mathsf{SR}\;D\;n\;m\;t
+\end{equation*}
+Combining the last two results, $\mathsf{output}\;D\;n\;x\;t$ is a sum of consecutive values of the first shift register cell, $\mathsf{SR}\;D\;n\;0$.
+Thus, we can express $\mathsf{output}\;D\;n\;x\;t$ in terms of $\mathsf{output}\;D\;n'\;x'\;t$, where $(n',x')=\mathsf{input}\;n$, since the source of $\mathsf{SR}\;D\;n\;0$ is the previous (input) slice's output.
 Finally, by induction on $n$, we can express $\mathsf{output}\;D\;n\;x\;t$ as a function of $D$ alone, as in Theorem~(\ref{output_input_at_update_times}), using the result above for the inductive case.
-\end{enumerate}
-
+\qed
 \end{proof}

### 2011-03-28 fmics/paper.tex
various minor corrections

@@
 Let $\mathsf{RS}\;n$ denote, as in Figure~\ref{fig:srs}, the $n$-th output of the SRS structure, which is supposed to compute a sum of received count values, and let $\mathsf{true\_sum}\;n$ denote this sum.
-The multi-layerd structure of the SRS and the read delay of each shift register result in the outputs of the SRS to be delayed from the $\mathsf{true\_sum}$ values.
+The multi-layered structure of the SRS and the read delay of each shift register result in the outputs of SRS being delayed from the $\mathsf{true\_sum}$ values.
 A sketch of the desired correctness statement is \[\forall{n} \cdot \, \mathsf{RS}\;n= \mathsf{true\_sum}\; n \pm\text{acceptable error}\]
@@
 Figure~\ref{fig:auxdefs} defines a set of additional functions required to prove our main results.
-The function $\mathsf{delay\_sum}\;n$ represents the accumulative delay of the preceding shift registers of a slice.
+The function $\mathsf{delay\_sum}\;n$ represents the cumulative delay of the preceding shift registers of a slice.
 For example, $\mathsf{delay\_sum}\;4 = \mathsf{delay}\;3 + \mathsf{delay}\;1 + \mathsf{delay}\;0 = 2 +1 +1 = 4$.
-The function $\mathsf{last\_update}\;n\;t$ specifies the time that the $n$-th slice last updated and the $\mathsf{exact}\;D\;n\;x\;t$ {\bf Ramana, can say in simple words what exact represents?}
+The function $\mathsf{last\_update}\;n\;t$ specifies the time that the $n$-th slice last updated, and $\mathsf{exact}\;D\;n\;x\;t$ computes the exact sum of consecutive input counts, without delay, that $\mathsf{output}\;D\;n\;x\;t$ is supposed to approximate (so $\mathsf{exact}$ is to $\mathsf{true\_sum}$ as $\mathsf{output}$ is to $\mathsf{RS}$).
 
@@
 More precisely, it says that if slice $n$ was just updated, then the output at tap $x$ is equal to the sum of $((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)$ input values, starting from the input $\mathsf{delay\_sum}\;n$ time steps ago.
-Note that the precise statement of such a theorem is complicated by details of the value of the $\mathsf{output}$ at times when the shift registers have not yet received any values.
+%Note that the precise statement of such a theorem is complicated by details of the value of the $\mathsf{output}$ at times when the shift registers have not yet received any values.
 
@@
 \begin{theorem}
-For all values of $D$, $n$, and $x$, the $\mathsf{output}$ of slice $n$ at update time $t$ is equal to
+For all values of $D$, $n$, and $x$, the $\mathsf{output}$ of slice $n$ at an update time $t$ satisfies
 \begin{equation*}%\label{output_input_at_update_times}
@@
 \begin{proof}
-The shift register cells get their values from the previous cell but they only update every $\mathsf{delay}\;n$ time steps. Therefore,
+The shift register cells get their values from the previous cell but they only update every $\mathsf{delay}\;n$ time steps.
+Therefore\footnote{We do not go through the details here of what happens when $t$ is small, for example $t<m\times\mathsf{delay}\;n$.},
 \begin{equation*}
@@
 \end{equation*}
-By induction on $m$ on the above,
+By induction on $m$, using the above,
 \begin{equation*}
@@
 \end{equation*}
-By induction on $t$, we can show that the $\mathsf{output}\;D\;n\;x\;t$ is a sum of values of consecutive shift register cells
+By induction on $t$, we can show that $\mathsf{output}\;D\;n\;x\;t$ is a sum of values of consecutive shift register cells
 \begin{equation*}
-\forall{D\,n\,x\,t}.\,{0<n}\implies\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\mathsf{SR}\;D\;n\;m\;t
+\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\mathsf{SR}\;D\;n\;m\;t
 \end{equation*}
@@
 Thus, we can express $\mathsf{output}\;D\;n\;x\;t$ in terms of $\mathsf{output}\;D\;n'\;x'\;t$, where $(n',x')=\mathsf{input}\;n$, since the source of $\mathsf{SR}\;D\;n\;0$ is the previous (input) slice's output.
-Finally, by induction on $n$, we can express $\mathsf{output}\;D\;n\;x\;t$ as a function of $D$ alone, as in Theorem~(\ref{output_input_at_update_times}), using the result above for the inductive case.
+Finally, by induction on $n$, we can express $\mathsf{output}\;D\;n\;x\;t$ as a function of $D$ alone, as required, using the result above for the inductive case.
 \qed

### 2011-03-28 fmics/paper.tex
Add last two theorems to main text, remove Figure 9
last theorem doesn't have proof yet

@@
 
-\begin{theorem}
-For all values of $D$, $n$, and $x$, the $\mathsf{output}$ of slice $n$ at an update time $t$ satisfies
+\begin{theorem}\label{output_input_at_update_times}
+For all values of $D$, $n$, and $x$, the $\mathsf{output}$ of slice $n>0$ at an update time $t$ satisfies
 \begin{equation*}%\label{output_input_at_update_times}
@@
 
-In light of Theorem~(\ref{output_last_update}), which says a slice's output stays constant between update times for the slice, Theorem~(\ref{output_input_at_update_times}) characterises all outputs at all times.
+Since a slice's output stays constant between update times for the slice, Theorem~\ref{output_input_at_update_times} characterises all outputs at all times.
 Thus, the SRS structure's outputs are exact running sums of the input counts over windows whose sizes depend on the the width of the slice and position of taps.
@@
 
-\begin{figure}
-\caption{
-A selection of theorems about our model proved using HOL4.
-\label{fig:theorems}
-}
-\begin{align}
-\label{update_time_last_update}
-&\forall{n\,t}.\,\mathsf{update\_time}\;n\;(\mathsf{last\_update}\;n\;t)\\
-%\label{last_update_thm}
-%&\forall{n\,t}.\,\mathsf{last\_update}\;n\;t=t-t\operatorname{mod}\mathsf{delay}\;n\\
-%SR_prev
-%SR_first
-\label{output_last_update}
-&\forall{D\,n\,x\,t}.\,\mathsf{output}\;D\;n\;x\;t=\mathsf{output}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t)\\
-\label{output_sum}
-&\forall{D\,n\,x\,t}.\,{0<n}\implies\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\mathsf{SR}\;D\;n\;m\;t
-\end{align}
+The fact that the SRS calculates exact, but delayed, sums, is captured in the next theorem, which relates $\mathsf{output}\;D\;n\;x\;t$ to $\mathsf{exact}\;D\;n\;x$ at an earlier time.
+
+\begin{theorem}\label{output_eq_exact}
+For all values of $D$, $n$, and $x$, the $\mathsf{output}$ of slice $n>0$ at all times $t$ satisfies
+\begin{equation*}
+\mathsf{output}\;D\;n\;x\;t=\begin{cases}0\qquad\qquad\qquad\qquad\qquad\quad\mathsf{last\_update}\;n\;t+1<\mathsf{delay\_sum}\;n\\\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)\quad\text{otherwise}\end{cases}
+\end{equation*}
+\end{theorem}
+\begin{proof}
+Follows from Theorem~\ref{output_input_at_update_times}, using the definition of $\mathsf{exact}$, and the fact that $\mathsf{output}\;D\;n\;x\;t$ has the same value as at the last update time.\qed
+\end{proof}
+
+We have also proved a preliminary result bounding the error of $\mathsf{output}\;D\;n\;x\;t$ compared to $\mathsf{exact}\;D\;n\;x\;t$.
+\begin{theorem}\label{max_error}
+For all $k$ and $D$ satisfying $|D\;(t'+1)-D\;t'|\leq k$ at all times, the following holds for all slices $n>0$, tap positions $x$, and times $t$:
+\begin{multline*}
+t>(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay}\;n)+\mathsf{delay}\;n+\mathsf{delay\_sum}\;n\implies\\
+|\mathsf{output}\;D\;n\;x\;t-\mathsf{exact}\;D\;n\;x\;t|\leq(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay\;n})\times k\times(\mathsf{delay}\;n+\mathsf{delay\_sum}\;n)
+\end{multline*}
+\end{theorem}
+\begin{proof}
+\end{proof}
+
+%\begin{figure}
+%\caption{
+%A selection of theorems about our model proved using HOL4.
+%\label{fig:theorems}
+%}
+%\begin{align}
+%\label{update_time_last_update}
+%&\forall{n\,t}.\,\mathsf{update\_time}\;n\;(\mathsf{last\_update}\;n\;t)\\
+%%\label{last_update_thm}
+%%&\forall{n\,t}.\,\mathsf{last\_update}\;n\;t=t-t\operatorname{mod}\mathsf{delay}\;n\\
+%%SR_prev
+%%SR_first
+%\label{output_last_update}
+%&\forall{D\,n\,x\,t}.\,\mathsf{output}\;D\;n\;x\;t=\mathsf{output}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t)\\
+%\label{output_sum}
+%&\forall{D\,n\,x\,t}.\,{0<n}\implies\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\mathsf{SR}\;D\;n\;m\;t
+%\end{align}
+%%\begin{multline}
+%%\label{output_source_at_update_times}
+%%\forall{D\,n\,x\,t}.\,{0<n}\land{\mathsf{update\_time\;n\;t}}\implies\\
+%%\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\begin{cases}0&t\leq m\times(\mathsf{delay}\;n)\\\mathsf{source}\;D\;n\;0\;(t-m\times(\mathsf{delay\;n})-1)&\text{otherwise}\end{cases}
+%%\end{multline}
 %\begin{multline}
-%\label{output_source_at_update_times}
-%\forall{D\,n\,x\,t}.\,{0<n}\land{\mathsf{update\_time\;n\;t}}\implies\\
-%\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\begin{cases}0&t\leq m\times(\mathsf{delay}\;n)\\\mathsf{source}\;D\;n\;0\;(t-m\times(\mathsf{delay\;n})-1)&\text{otherwise}\end{cases}
+%\label{output_prev_at_update_times}
+%\forall{D\,n\,x\,t}.\,{0<n}\land{\mathsf{update\_time}\;n\;t}\implies\\
+%\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\begin{cases}0&t < m\times\mathsf{delay}\;n+\mathsf{delay}\;n'\\\mathsf{output}\;D\;n'\;x'\;(t-m\times\mathsf{delay}\;n-\mathsf{delay}\;n')&\text{otherwise}\end{cases}\\
+%\qquad\text{where $(n',x')=\mathsf{input}\;n$}
 %\end{multline}
-\begin{multline}
-\label{output_prev_at_update_times}
-\forall{D\,n\,x\,t}.\,{0<n}\land{\mathsf{update\_time}\;n\;t}\implies\\
-\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\begin{cases}0&t < m\times\mathsf{delay}\;n+\mathsf{delay}\;n'\\\mathsf{output}\;D\;n'\;x'\;(t-m\times\mathsf{delay}\;n-\mathsf{delay}\;n')&\text{otherwise}\end{cases}\\
-\qquad\text{where $(n',x')=\mathsf{input}\;n$}
-\end{multline}
-\begin{multline}\label{output_input_at_update_times}
-\forall{D\,n\,x\,t}.\,{0 < n}\land{\mathsf{update\_time\;n\;t}}\implies\\\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)-1}\begin{cases}0&t<m+\mathsf{delay\_sum}\;n\\D\;(t-m-\mathsf{delay\_sum}\;n)&\text{otherwise}\end{cases}
-\end{multline}
-\begin{multline}\label{output_eq_exact}
-\forall{D\,n\,x\,t}.\,{0 < n}\implies\\\mathsf{output}\;D\;n\;x\;t=\begin{cases}0&\mathsf{last\_update}\;n\;t<\mathsf{delay}\;n\\\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)\end{cases}
-\end{multline}
-%max_error
-\end{figure}
+%\begin{multline}\label{output_eq_exact}
+%\forall{D\,n\,x\,t}.\,{0 < n}\implies\\\mathsf{output}\;D\;n\;x\;t=\begin{cases}0&\mathsf{last\_update}\;n\;t<\mathsf{delay}\;n\\\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)\end{cases}
+%\end{multline}
+%%max_error
+%\end{figure}
 

### 2011-03-28 fmics/paper.tex
some tweaks to a table in section 3

@@
 
-\usepackage{times,graphicx,epsfig,boxedminipage,url,xspace,array,amsmath}
+\usepackage{times,graphicx,epsfig,boxedminipage,url,xspace,array,amsmath,textcomp}
 
@@
 
-As shown in Table~\ref{}, by cascading 6 slices, it is enough to reach the 100 seconds upper integration limit requested by the specifications.
+As shown in Table~\ref{fig:example-bakery}, cascading 6 slices is enough to reach the 100 second upper integration limit requested by the specifications.
 
@@
 
-\begin{figure}[t]
+\begin{table}[t]
 \centering
-\begin{tabular}{|c|c|c|c|c|c|}
+\begin{tabular}{|r|r|r|r|c|l|}
 \hline
 \multicolumn{2}{|c}{Range} &\multicolumn{2}{|c|}{Refreshing} & & \\\cline{1-4}
-40 us steps & ms & 40 us steps  & ms & slice No & running sum \\\hline\hline
+40 {\textmu}s steps & ms & 40 {\textmu}s steps  & ms & slice & running sum \\\hline\hline
 1 & 0.04 & 1 & 0.04 & & RS0 \\\hline
@@
 2097152 & 83886.08 & 16384 & 655.36 & slice 5 & RS11 \\\hline
-
 \end{tabular}
 \vspace{0.04in}
-\label{fig:example-bakery}
-\caption{}
-\end{figure}
+\caption{{\bf<missing caption>}
+\label{fig:example-bakery}}
+\end{table}
 

### 2011-03-29 fmics/paper.tex
merge changes from dropbox

@@
 
-%Figure~\ref{fig:theorems} gives a selection of theorems we have proved in HOL, some of which refer to additional constants defined in Figure~\ref{fig:auxdefs}.
 Figure~\ref{fig:auxdefs} defines a set of additional functions required to prove our main results.
@@
 For example, $\mathsf{delay\_sum}\;4 = \mathsf{delay}\;3 + \mathsf{delay}\;1 + \mathsf{delay}\;0 = 2 +1 +1 = 4$.
-The function $\mathsf{last\_update}\;n\;t$ specifies the time that the $n$-th slice last updated, and $\mathsf{exact}\;D\;n\;x\;t$ computes the exact sum of consecutive input counts, without delay, that $\mathsf{output}\;D\;n\;x\;t$ is supposed to approximate (so $\mathsf{exact}$ is to $\mathsf{true\_sum}$ as $\mathsf{output}$ is to $\mathsf{RS}$).
+The function $\mathsf{last\_update}\;n\;t$ specifies the time that the $n$-th slice last updated, and $\mathsf{exact}\;D\;n\;x\;t$ computes the exact sum of consecutive input counts, without delay, that $\mathsf{output}\;D\;n\;x\;t$ is supposed to approximate.
+% (so $\mathsf{exact}$ is to $\mathsf{true\_sum}$ as $\mathsf{output}$ is to $\mathsf{RS}$).
 
-%Our central result, Theorem~(\ref{output_input_at_update_times}), equates the output of a slice to a sum of consecutive input counts.
 Our central result equates the output of a slice to a sum of consecutive input counts.
 More precisely, it says that if slice $n$ was just updated, then the output at tap $x$ is equal to the sum of $((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)$ input values, starting from the input $\mathsf{delay\_sum}\;n$ time steps ago.
-%Note that the precise statement of such a theorem is complicated by details of the value of the $\mathsf{output}$ at times when the shift registers have not yet received any values.
-
-%To prove Theorem~(\ref{output_input_at_update_times}), we first make the following observations\footnote{The precise statements of these observations are complicated by details of what happens when $t$ is low, or equivalently, when we are looking at shift register cells that have not yet received any input.}
 
@@
 For all values of $D$, $n$, and $x$, the $\mathsf{output}$ of slice $n>0$ at an update time $t$ satisfies
-\begin{equation*}%\label{output_input_at_update_times}
-%\forall{D\,n\,x\,t}.\,{0 < n}\land{\mathsf{update\_time\;n\;t}}\implies\\
+\begin{equation*}
 \mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)-1}\begin{cases}0 \qquad\qquad\qquad  t<m+\mathsf{delay\_sum}\;n\\D\;(t-m-\mathsf{delay\_sum}\;n)\quad\text{otherwise}\end{cases}
@@
 \begin{proof}
-The shift register cells get their values from the previous cell but they only update every $\mathsf{delay}\;n$ time steps.
-Therefore\footnote{We do not go through the details here of what happens when $t$ is small, for example $t<m\times\mathsf{delay}\;n$.},
+The shift register of slice $n$ is updated every $\mathsf{delay}\;n$ time steps.
+When updated, the values in its cells are shifted one cell.
+%The shift register cells get their values from the previous cell but they only update every $\mathsf{delay}\;n$ time steps. Therefore
+Therefore\footnote{For presentation convenience, here, we do not provide the details regarding the case when $t$ is small, for example $t<m\times\mathsf{delay}\;n$.
+For the full description of the theorems and proof script, please refer to {\bf Can we put a link pointing to the proof script here?}},
 \begin{equation*}
@@
 
-Since a slice's output stays constant between update times for the slice, Theorem~\ref{output_input_at_update_times} characterises all outputs at all times.
-Thus, the SRS structure's outputs are exact running sums of the input counts over windows whose sizes depend on the the width of the slice and position of taps.
+Since the outputs of a slice stay constant between update times for the slice, Theorem~\ref{output_input_at_update_times} suffices to characterize all outputs at all times.
+Thus, the SRS structure's outputs are equal to the exact running sums of the input counts over windows whose sizes depend on the the width of the slice's shift register and position of the taps.
 However, the sums are delayed by the total delay across all previous slices (represented by $\mathsf{delay\_sum}$).
@@
 \begin{proof}
-Follows from Theorem~\ref{output_input_at_update_times}, using the definition of $\mathsf{exact}$, and the fact that $\mathsf{output}\;D\;n\;x\;t$ has the same value as at the last update time.\qed
+The proof follows from Theorem~\ref{output_input_at_update_times}, using the definition of $\mathsf{exact}$, and the fact that $\mathsf{output}\;D\;n\;x\;t$ has the same value as at the last update time.\qed
 \end{proof}
+
+Note that the definition of the $\mathsf{true\_sum}$ is easily defined in terms of $\mathsf{exact}$, that is $\mathsf{true\_sum}\;D\;n\;t = \mathsf{exact}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)\;t$.
+Sine the $\mathsf{RS}$ is also similarly defined in terms of the $\mathsf{output}$, Theorem~\ref{output_eq_exact} establishes a relationship between the values of the $\mathsf{RS}$ and $\mathsf{true\_sum}$.
+
+We formulate the difference between the $\mathsf{RS}$ values and their corresponding $\mathsf{true\_sum}$s, caused by the delay, as $\mathsf{error}$. 
+In order to define a maximum bound on the $\mathsf{error}$, we had to formulate some characterization of the input stream to the SRS structure as constraints on $D$.
+One of these constraints is the bound on the difference between the two consecutive count values.
 
@@
 %\end{multline}
+%\begin{multline}\label{output_input_at_update_times}
+%\forall{D\,n\,x\,t}.\,{0 < n}\land{\mathsf{update\_time\;n\;t}}\implies\\\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{((\mathsf{tap}\;n\;x)+1)\times(\mathsf{delay}\;n)-1}\begin{cases}0&t<m+\mathsf{delay\_sum}\;n\\D\;(t-m-\mathsf{delay\_sum}\;n)&\text{otherwise}\end{cases}
+%\end{multline}
 %\begin{multline}\label{output_eq_exact}

### 2011-03-29 fmics/paper.tex
some minor wording tweaks

@@
 
-Note that the definition of the $\mathsf{true\_sum}$ is easily defined in terms of $\mathsf{exact}$, that is $\mathsf{true\_sum}\;D\;n\;t = \mathsf{exact}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)\;t$.
-Sine the $\mathsf{RS}$ is also similarly defined in terms of the $\mathsf{output}$, Theorem~\ref{output_eq_exact} establishes a relationship between the values of the $\mathsf{RS}$ and $\mathsf{true\_sum}$.
+Note that $\mathsf{true\_sum}$ is easily defined in terms of $\mathsf{exact}$, that is $\mathsf{true\_sum}\;D\;n\;t = \mathsf{exact}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)\;t$.
+Since $\mathsf{RS}$ is also similarly defined in terms of the $\mathsf{output}$, Theorem~\ref{output_eq_exact} establishes a relationship between the values of the $\mathsf{RS}$ and $\mathsf{true\_sum}$.
 
@@
 In order to define a maximum bound on the $\mathsf{error}$, we had to formulate some characterization of the input stream to the SRS structure as constraints on $D$.
-One of these constraints is the bound on the difference between the two consecutive count values.
+One of these constraints is a bound on the difference between two consecutive count values.
 
-We have also proved a preliminary result bounding the error of $\mathsf{output}\;D\;n\;x\;t$ compared to $\mathsf{exact}\;D\;n\;x\;t$.
+Under such a constraint, we have proved a preliminary result bounding the error of $\mathsf{output}\;D\;n\;x\;t$ compared to $\mathsf{exact}\;D\;n\;x\;t$.
 \begin{theorem}\label{max_error}

### 2011-03-29 fmics/paper.tex
add link to proof script online
proof sketch for theorem 3

@@
 Therefore\footnote{For presentation convenience, here, we do not provide the details regarding the case when $t$ is small, for example $t<m\times\mathsf{delay}\;n$.
-For the full description of the theorems and proof script, please refer to {\bf Can we put a link pointing to the proof script here?}},
+For complete theorem statements and the proof script, please refer to \url{https://github.com/xrchz/CERN-LHC-BLMTC-SRS/blob/master/hol/srsScript.sml}.},
 \begin{equation*}
@@
 \begin{proof}
+By applying Theorem~\ref{output_eq_exact}, we are left with an equation involving $\mathsf{exact}$ but not $\mathsf{output}$, specifically between $\mathsf{exact}\;D\;n\;x\;t$ and $\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)$.
+Our assumption on $t$ ensures that the case where the function is equal to $0$ in Theorem~\ref{output_eq_exact} or in the definition of $\mathsf{exact}$ never applies.
+Now, $\mathsf{exact}$ is defined as a sum of consecutive values of $D$, so we simply need to bound the difference between one such sum and another earlier one of the same size.
+If at each step the maximum difference is $k$, then the total difference is at most $k$ times the distance between the ends of the two sums, which is the bound stated in the theorem <actually the theorem may have an additional factor increasing the size of the bound, that made it easier to prove - not sure>.
 \end{proof}

### 2011-03-30 fmics/paper.tex
merge changes to paper.tex from dropbox

@@
 This collaborative effort has individuals with diverse backgrounds ranging from theoretical physics to system safety.
-%One interesting result is the extent to which
 The use of a formal method has compelled the stakeholders to clarify intricate details of the SRS structure and behaviour.
@@
 
-%The BLMS detects beam loss in the form of counts proportional to the number of stray particles.
-Dangerous loss levels can be expressed as thresholds on the rate of counts over time.
-The SRS maintains approximate sums of counts over time windows of various sizes, from which loss rates can be calculated.
-It is preferable for the SRS sum to be above the exact sum over a particular time window, so that dangerous losses are always detected.
-But stopping the beam when the loss level is not dangerous is time consuming and costly.
-Therefore, the discrepancy between the SRS sum and the exact sum needs to be small.
-
-%************** Rushby's quote to be added later on *****
-%Rushby's emphasis on the role of mechanized logical deduction as an "implacable skeptic that insists on its human user providing justification" thus rings true.
-
 \section{BLMS Overview}
@@
 The loss measurement principle \cite{Dehning-IPAC,Chris-FPGA} is based on the energy deposition detection of secondary shower particles using a specially-designed detector called \emph{ionization chamber}.
-There are around 4000 ionization chambers located under the magnets all around the tunnel.
+There are around 4000 ionization chambers strategically placed on the sides of the magnets all around the tunnel.
 These detectors convert the signal given by the recording of shower particles into electrical signals and send them to tunnel cards.
@@
 Each tunnel card receives data from eight detectors and each surface card receives data from two tunnel cards.
-The surface card provides data to the Logging, the Post Mortem and the Collimation systems that will be used to drive on-line displays in the control room to allow offline analysis of the losses and to setup automatically the collimators.
+The surface card provides data to the Logging, the Post Mortem and the Collimation systems that will be used to drive on-line displays in the control room, perform long-term storage for offline analysis, and to setup the collimators automatically.
 Due to demanding performance requirements, the BLMS is implemented on modern FPGAs, which include the resources needed to design complex processing and can be reprogrammed making them ideal for future upgrades or system specification changes.
@@
 40 {\textmu}s steps & ms & 40 {\textmu}s steps  & ms & slice & running sum \\\hline\hline
-1 & 0.04 & 1 & 0.04 & & RS0 \\\hline
-2 & 0.08 & 1 & 0.04 & & RS1 \\\hline
-8 & 0.32 & 1 & 0.04 & slice 1 & RS2 \\\hline
-16 & 0.64 & 1 & 0.04 & & RS3 \\\hline
-64 & 2.56 & 2 & 0.08 & slice 2 & RS4 \\\hline
-256 & 10.24 & 2 & 0.08 & & RS5 \\\hline
-2048 & 81.92 & 64 & 2.56 &slice  3 & RS6 \\\hline
-16384 & 655.36 & 64 & 2.56 &slice 3 & RS7 \\\hline
-32768 &1310.72 & 2048 & 81.92 & slice 4 & RS8 \\\hline
-131072 & 5242.88 & 2048 & 81.92 & slice 4 & RS9 \\\hline
-524288 & 2097.52 & 16384 & 655.36 & slice 5 &RS10 \\\hline
-2097152 & 83886.08 & 16384 & 655.36 & slice 5 & RS11 \\\hline
+1 & 0.04 & 1 & 0.04 & slice 1 & RS0 \\\hline
+2 & 0.08 & 1 & 0.04 & slice 1 & RS1 \\\hline
+8 & 0.32 & 1 & 0.04 & slice 2 & RS2 \\\hline
+16 & 0.64 & 1 & 0.04 & slice 2 & RS3 \\\hline
+64 & 2.56 & 2 & 0.08 & slice 3 & RS4 \\\hline
+256 & 10.24 & 2 & 0.08 & slice 3 & RS5 \\\hline
+2048 & 81.92 & 64 & 2.56 &slice 4 & RS6 \\\hline
+16384 & 655.36 & 64 & 2.56 &slice 4 & RS7 \\\hline
+32768 &1310.72 & 2048 & 81.92 & slice 5 & RS8 \\\hline
+131072 & 5242.88 & 2048 & 81.92 & slice 5 & RS9 \\\hline
+524288 & 2097.52 & 16384 & 655.36 & slice 6 &RS10 \\\hline
+2097152 & 83886.08 & 16384 & 655.36 & slice 6 & RS11 \\\hline
 \end{tabular}
@@
 \label{sec-verification}
-
-%The calculation of running sums is a critical component of the BLMS.
-%If there is no dangerous loss, the BLMS should not inhibit the beam.
-%A false alarm costs at least three hours to resume operation, and compromises the availability of the system.
-%However, if there is a dangerous loss, the BLMS must inhibit the beam.
-%A quench due to dangerous loss costs at least 30 days of downtime.
-%Thus, any error in the calculation compromises either the availability or the safety of the system.
-%The current practice to check the correctness of the SRS is by simulation and testing.
-%(isn't all this in the introduction now? is this the right place to repeat it/give more details?)
-%In this paper we present a formal verification approach for the analysis of SRS implementation in BLM.
 
@@
 The function $\mathsf{last\_update}\;n\;t$ specifies the time that the $n$-th slice last updated, and $\mathsf{exact}\;D\;n\;x\;t$ computes the exact sum of consecutive input counts, without delay, that $\mathsf{output}\;D\;n\;x\;t$ is supposed to approximate.
-% (so $\mathsf{exact}$ is to $\mathsf{true\_sum}$ as $\mathsf{output}$ is to $\mathsf{RS}$).
 
@@
 When updated, the values in its cells are shifted one cell.
-%The shift register cells get their values from the previous cell but they only update every $\mathsf{delay}\;n$ time steps. Therefore
 Therefore\footnote{For presentation convenience, here, we do not provide the details regarding the case when $t$ is small, for example $t<m\times\mathsf{delay}\;n$.
@@
 One of these constraints is a bound on the difference between two consecutive count values.
+Such a bound can be derived from the highest quench level thresholds.
 
@@
 \end{proof}
+
+\section{Conclusions and Lessons Learned}
+\label{sec-conclusions}
+
+We described a case study in which HOL4 theorem prover is used to formally verify properties of the SRS structure within the LHC machine protection system.
+In this case study, we built a parameterized model of the SRS structure and showed that its behavior is correct with respect to its specification.
+It is likely that the configuration of the slices and shift registers will change in the future to accommodate higher precision.
+Thus, the parameterization in our model was crucial to make it applicable to the future upgrades.
+
+One of our main challenges in this effort was building the formal model and formulating the correctness statements.
+To understand the structure of the SRS, we started with a model of a simplified structure, which had a more regular arrangements of the slices and ignored the middle taps, and proceeded to a model which is closer to the SRS implementation in BLMS.
+In addition, we used a basic spreadsheet simulation model for a few slices for sanity testing the correctness formulae.
+However, some sanity tests related to Theorem 1 led us astray when we did not know the correct formula for that theorem.
+Because the formulae we conjectured worked for small values of $n$, and simulating large values of $n$ was expensive, but the counterexamples were only on large values of $n$.
+The most important use of mechanized logical deduction in this effort was its role as an ``implacable skeptic" that insisted on us providing the justification and compelled us to clarify the intricate details of the SRS structure and behavior.
+
+The use of mechanized theorem proving to verify the correctness of the SRS behavior complements an intensive verification based on simulation performed at CERN.
+The simulation was mainly based on two scenarios one for recording fast losses and  another for the situation when a loss gradually builds up in the accelerator.
+The results of the simulation showed that the current implementation of SRS, given in Figure~\ref{fig:srs}, satisfies the specification, that is the running sums have  the maximum 20\% relative error.
+On the contrary, the results in this paper apply to all possible input steams, not just a limited number of input streams simulating fast or slow loss scenarios.
+But, the error bound given in Theorem~\ref{max_error} does not satisfy the maximum 20\% relative error.
+In order to make this bound tighter, we need to provide more characterization of the input as constraints on $D$.
+Due to the nature of the problem, such a characterization of input is not trivial.
+Therefore, the SRS structure in LHC is an interesting example in which the formal verification does not completely replace conventional verification approaches such as simulation.
+
+%Future direction, go with lower level implementation rather than finite words (closer  to vhdl coed) and improve the bound
+
+%The BLMS detects beam loss in the form of counts proportional to the number of stray particles.
+Dangerous loss levels can be expressed as thresholds on the rate of counts over time.
+The SRS maintains approximate sums of counts over time windows of various sizes, from which loss rates can be calculated.
+It is preferable for the SRS sum to be above the exact sum over a particular time window, so that dangerous losses are always detected.
+But stopping the beam when the loss level is not dangerous is time consuming and costly.
+Therefore, the discrepancy between the SRS sum and the exact sum needs to be small.
+
+%The calculation of running sums is a critical component of the BLMS.
+%If there is no dangerous loss, the BLMS should not inhibit the beam.
+%A false alarm costs at least three hours to resume operation, and compromises the availability of the system.
+%However, if there is a dangerous loss, the BLMS must inhibit the beam.
+%A quench due to dangerous loss costs at least 30 days of downtime.
+%Thus, any error in the calculation compromises either the availability or the safety of the system.
+%The current practice to check the correctness of the SRS is by simulation and testing.
+%(isn't all this in the introduction now? is this the right place to repeat it/give more details?)
+%In this paper we present a formal verification approach for the analysis of SRS implementation in BLM.
+
+\bibliographystyle{plain}
+\bibliography{refs}
+
+\end{document}
 
@@
 %\end{figure}
-
-%One of our objectives in formally modelling SRS was to keep the topology parameterized.
-%It is likely that the configuration of the slices and shift registers will in the future change to accommodate higher precision.
-%A parameterized model can be used both for current and future versions of the system.
-%For this, we formalized the definitions of the model by six parameters:
-
-%Our work proceeded in two phases, starting with a model of a simplified structure and proceeding to a model which is closer to the SRS implementation in BLMS.
-%In both phases, we model each building block of the SRS structure that holds a count value -- for example, each cell in each shift register -- as a function in HOL that takes the input stream and current time step as inputs and returns the value the corresponding building block holds at that time.
-
-%Our simplifications in the first phase include:
-%\begin{enumerate}
-%\item Masking the middle taps in each slice,
-%\item Considering a more regular arrangement of slices than the SRS implementation in BLMS,
-%\item Masking the subtractors and the accumulators in each slice, instead, defining the output of a slice as the sum of the values in its cells directly,
-%\item Simplifying timing by ignoring some of the delays that each building block may introduce, {\bf this is not true - there were delays in the simplified model}
-%\item Using natural numbers, rather than finite words, for count values.
-%\end{enumerate}
-
-%Building a logical model for simplified structure of the SRS helped us improve our understanding of the behavior of the SRS.
-
-%In the second phase, we eliminated the first four simplifications (listed above) by adding middle taps, a more flexible arrangement of slices, the subtractor and accumulator, and considering the delays that each block may introduce.
-%In both cases, however, the model is simultaneously a simplification and a generalization of the actual structure of SRS in BLMS in the following way: we model an infinite number of slices (rather than six slices), each with an infinite number of shift register cells and taps (rather than fixed width shift registers and only two taps), simply by letting indices range over the natural numbers without explicitly giving limits.
-%This parameterization makes the model and the analysis more likely to be applicable to future upgraded/modified versions of the system.
-
-\section{Conclusions and Lessons Learned}
-\label{sec-conclusions}
-
-\bibliographystyle{plain}
-\bibliography{refs}
-
-\end{document}

### 2011-03-30 fmics/paper.tex
tweak a paragraph in the introduction

@@
 
-Compared to test-based methods, like simulation, formal methods not only offer a much higher confidence in correctness of a system's behavior, but also help improve our understanding of its specification.
-One of the challenges in pursuing a formal verification approach for SRS was capturing the intricate details of the system's specification via experiment and refinement with a team with different backgrounds and expertise.
-Our confidence in the SRS design as a result of this effort ultimately rests upon our deep understanding of why the design is correct rather than the fact that we obtained ``Theorem Proved" as the final output of a software tool at the end of a long and intense effort.
+Compared to test-based methods, like simulation, formal methods not only offer much higher confidence in the correctness of a system's behavior, but also help improve our understanding of its specification.
+One of the challenges in pursuing a formal verification approach for SRS was capturing the intricate details of the system's specification via experiment and refinement with a team of different backgrounds and expertise.
+Our confidence in the SRS design ultimately rests upon our deep understanding of why the design is correct rather than the fact that we obtained ``Theorem Proved" as the final output of a software tool.
 In particular, our use of mechanized logical deduction was a highly iterative process that incrementally refined our understanding of (1) the implementation (2) the intended behavior and (3) the ``white board level" argument or explanation for why the implementation achieves the intended behaviors.

### 2011-03-30 fmics/paper.tex
reword last paragraph of introduction

@@
 
-The rest of the paper is organized as follows.
-An overview of BLMS and its main components is given in Section~\ref{sec-BLM} and is followed by a description of SRS implementation in Section~\ref{sec-SRS}.
-The verification approach of SRS is presented in Section~\ref{sec-verification}.
-We conclude in Section~\ref{sec-conclusions} with a summary of lessons learned and future directions.
+Our contributions in this paper are: a formal model of the SRS component of the BLMS, and commentary on the process and outcomes of taking a formal approach.
+We give an overview of the BLMS in Section~\ref{sec-BLM}, and describe the SRS component in particular in Section~\ref{sec-SRS}.
+In Section~\ref{sec-verification} we describe our approach to formal verification, and present the formal model and results.
+Finally, we reflect on the process, summarising lessons learned and future directions, in Section~\ref{sec-conclusions}.
 

### 2011-03-30 fmics/paper.tex
Reword first paragraph of Section 2

@@
   \centering  \scalebox{0.7}{ \includegraphics{BLETC.eps}}
-   \caption{Block diagram of the processes assigned to BLETC card.}
+   \caption{Block diagram of a BLETC card.}
   \label{fig:BLETC}
@@
 
-The loss measurement principle \cite{Dehning-IPAC,Chris-FPGA} is based on the energy deposition detection of secondary shower particles using a specially-designed detector called \emph{ionization chamber}.
-There are around 4000 ionization chambers strategically placed on the sides of the magnets all around the tunnel.
-These detectors convert the signal given by the recording of shower particles into electrical signals and send them to tunnel cards.
-Tunnel cards, which are located in the tunnel and implemented by radiation-tolerant electronics, acquire and digitize the data from the detectors and transmit those to the surface using optical links.
-There, the data processing cards, named BLETCs, receive those data and decide whether or not the beam should be permitted to be injected or continue circulating.
-Each tunnel card receives data from eight detectors and each surface card receives data from two tunnel cards.
-The surface card provides data to the Logging, the Post Mortem and the Collimation systems that will be used to drive on-line displays in the control room, perform long-term storage for offline analysis, and to setup the collimators automatically.
-Due to demanding performance requirements, the BLMS is implemented on modern FPGAs, which include the resources needed to design complex processing and can be reprogrammed making them ideal for future upgrades or system specification changes.
-Figure~\ref{fig:BLETC} shows a block diagram of the processes assigned in BLETC FPGA.
-In the following, we give a brief overview of each of the four main processing blocks in BLETC card.
+The purpose of the BLMS is to measure particle loss, and to request beam extraction if the loss level indicates that a quench is likely to occur.
+The physical principle underlying particle loss measurement \cite{Dehning-IPAC,Chris-FPGA} is the detection of energy deposited by \emph{shower particles} in \emph{ionization chambers}.
+There are around 4000 ionization chambers strategically placed on the sides of the magnets all around the LHC tunnel underground.
+The ionization chambers produce an electrical signal, based on the recording of shower particles, and transmit the signal to \emph{tunnel cards}.
+Tunnel cards, which are implemented by radiation-tolerant electronics, digitize the data and transmit the digitized data to the surface above the tunnel using optical links.
+At the surface, data processing cards named \emph{BLETCs} receive the data and decide whether or not the beam should be permitted to continue circulating.
+Each tunnel card receives data from eight ionization chambers, and each BLETC receives data from two tunnel cards.
+A BLETC provides data to the Logging, Post Mortem, and Collimation systems that drive on-line displays in the control room, perform long-term storage for offline analysis, and setup the collimators automatically {\bf What is a collimator?
+Why is it necessary to mention all this?}.
+Due to demanding performance requirements, BLETCs are implemented on FPGAs, which include the resources needed to design complex processing and can be reprogrammed making them ideal for future upgrades or system specification changes.
+Figure~\ref{fig:BLETC} shows a block diagram of the processes on a BLETC FPGA.
+In the following, we briefly describe each of the four processing blocks on a BLETC card.
 

### 2011-03-30 fmics/paper.tex
Revise rest of Section 2

@@
 
-\emph{(a) Receive, Check, and Compare (RCC)}: RCC is responsible for facilitating the reception process at the entry stage of the surface FPGA.
-It ensures the correct reception and detection of erroneous transmissions by redundancy in transmissions and using Cyclic Redundancy Check (CRC) and the 8B/10B algorithms.
+\emph{(a) Receive, Check, and Compare (RCC)}: The RCC block receives the data direct from the tunnel, and attempts to detect and correct erroneous transmissions by using Cyclic Redundancy Check and 8B/10B algorithms.{\bf references for these algorithms?}
 
-\emph{(b) Data Processing}: Quench of magnets that result from proton loss depends on the loss duration and the beam energy.
-Given the tolerance acceptable for quench prevention, the quench threshold versus loss time is approximated with a minimum number of sliding integration windows (called \emph{Running Sums}) fulfilling the tolerance.
-In order to achieve the \emph{dynamic range} (i.e.
-the domain of variation of the beam losses) requested by the specification, the detectors use both Current-to-Frequency Converters (CFCs) and ADC circuitries.
-The data processing block merges these two types of data coming from the same detector into one value and send it to the SRS block.
-The implementation of the SRS is discussed in more detail in the next section.
+\emph{(b) Data Processing}: Whether or not a quench results from particle loss depends on the loss duration and the beam energy.
+Given the tolerance acceptable for quench prevention, the loss level versus loss duration is approximated by sliding integration windows (called \emph{Running Sums}) fulfilling the tolerance.{\bf not entirely sure what this sentence means}
+In order to achieve the \emph{dynamic range} (domain of variation of losses) requested by the specification, the detectors use both Current-to-Frequency converters and ADC circuitries.{\bf ADC?}
+The data processing block merges these two types of data coming from the detector so as to send a single count value to the SRS block.
+The implementation of the SRS block is described in Section~\ref{sec-SRS}.
 
-\emph{(c) Threshold Comparator}: Every new calculated running sum need to be compared with the corresponding threshold that was chosen by the beam energy reading given that moment.
-The comparator will initiate a beam dump request if any of the running sums is higher than its corresponding threshold.
-The beam dump requests are forwarded to the Beam Interlock System which will initiate the beam dump.
-There are 12 running sums calculated for each 16 detector channels allocated to a BLETC card.
-The beam energy information is scaled into 32 levels (0.45 to 7 TeV) and each processing module will hold data only for those 16 detectors connected.
-That would give a total of 6,144 threshold values needed to be held on each card.
+\emph{(c) Threshold Comparator}: Every running sum needs to be compared to the threshold for the beam energy reading at that moment.
+The comparator initiates a beam dump request if any of the running sums is higher than its corresponding threshold.
+Beam dump requests are forwarded to the Beam Interlock System which initiates the beam dump.
+There are 12 running sums calculated for each 16-detector channel allocated to a BLETC card.
+The beam energy information is scaled to 32 levels (0.45 to 7 TeV) and each processing module holds data only for those 16 connected detectors.
+Thus a total of 6,144 threshold values need to be held on each card.
 
-\emph{(d) Logging, Post Mortem and Collimation}:  To be able to trace back the loss signal development, BLM should store the loss measurement data.
-This data will be sent over the VME-bus for online viewing and storage by the Logging and Post-Mortem systems.
-For the purpose of supervision, the BLM system will drive an online event display to show error and status information recorded by tunnel electronics and the RCC process as well as the maximum loss rates seen by the running sums.
-BLETC card also provides data to Collimation system for the correct alignment and setup of the collimators.
+\emph{(d) Logging, Post Mortem and Collimation}:  To be able to trace back the loss signal development, the BLMS stores the loss measurement data.
+This data is sent over the VME-bus {\bf what's the VME-bus?} for online viewing and storage by the Logging and Post-Mortem systems.
+For the purpose of supervision, the BLMS drives an online event display to show error and status information recorded by the tunnel electronics and the RCC process as well as the maximum loss rates seen by the running sums.
+Each BLETC card also provides data to the Collimation system for the correct alignment and setup of the collimators.
 

### 2011-03-30 fmics/paper.tex
Revise Section 3

@@
 
-Beam losses can happen either in a single cycle of the beam in the tunnel, with a sudden beam loss, or in progressive losses during numerous cycles.
+Beam losses can happen at different rates, compared to the number of cycles of the beam around the tunnel.
 One-cycle failures are called \emph{ultra-fast} losses.
-Multi-cycle losses can be divided between \emph{very fast} losses those which happen in less than 10 ms, \emph{fast} losses which happen in more than 10 ms and \emph{steady} losses, where the beam is lost over one second or more~\cite{Schmidt-ICFA}.
+Multi-cycle losses can be classified as: \emph{very fast} losses, which happen in less than 10 ms; \emph{fast} losses, which happen between 10 ms and 1 s; and, \emph{steady} losses, where the beam is lost over one second or more~\cite{Schmidt-ICFA}.
 
-The processing of the data collected by the detectors involves a proper analysis of the loss patterns over time and a proper account of the energy of the beam.
-The procedure for the data processing is based on the idea that a constantly updated moving window can be kept by adding to a register the incoming newest value and subtracting its oldest value (see Figure~\ref{fig:RS-basic}).
-The number of values that are kept under the window defines the integration time it represents.
-The ideal strategy is to have an infinite number of constantly updated moving windows with various lengths to cover the whole time region from 40 micro-seconds (the data arrival rate to BLETC card)  to 100 seconds for detecting ultra-fast losses to steady losses, respectively.
-Such an implementation is not feasible due to the limited amount of resources.
-Therefore, given the tolerance acceptable for quench prevention, the quench threshold versus loss time curve is approximated by a minimum number of steps that fulfill this tolerance.
+Processing the data collected by the detectors, referred to as \emph{count} values, involves an analysis of the loss pattern over time, accounting for the energy of the beam.
+The processing procedure is based on the idea that a constantly updated moving window can be maintained in an accumulator by adding the incoming (newest) value and subtracting the oldest value (see Figure~\ref{fig:RS-basic}).
+The number of values in the window is its \emph{integration time}.
+Ideally, we would have an infinite number of windows with lengths covering the whole spectrum of times from 40 micro-seconds (the rate at which count values enter a BLETC card) to 100 seconds, for detecting all losses from ultra-fast up to steady.
+To approximate this ideal with finite resources, the BLMS is given the tolerance acceptable for quench prevention, and the quench threshold versus loss time curve is approximated by the minimum number of windows that meet the tolerance.
 
@@
 
-In order to construct long moving windows, it is required to keep long histories of received data, referred to as \emph{count} values.
-The technique employed to reach long integration periods with relatively small in length shift registers uses consecutive storage of sums of the counts.
-Instead of storing all values needed for the sum, this technique stores successively parts of the total sum using only a fraction of the otherwise needed memory space.
+Long moving windows, that is, windows with large integration times, are required, which means keeping long histories of received count values.
+To accomplish this goal with relatively narrow shift registers, the SRS uses
+consecutive storage of sums of counts.
+Instead of storing all the values needed for a sum, the SRS accumulates many values as a partial sum, thereby using only a fraction of the memory space.
+The partial sums for a window with a large integration time are chosen so that they also serve as the sums calculated by a window with a smaller intergration time.
 This technique works by feeding the sum of one shift register's contents, every time its contents become completely updated, to the input of another shift register (see Figure~\ref{fig:SRS-basic}).
-By cascading more of these elements, very long moving windows could be constructed with significantly small memory space.
-The above scheme is the basis for the SRS implementation in BLETC.
+By cascading shift registers like this, very long moving windows can be constructed using a significantly small amount of memory.
+This scheme is the basis for the SRS implementation in each BLETC.
 
-SRS implementation minimizes the resource usage by using the already calculated running sums in order to calculate bigger in length running sums without the need of extra summation points.
+The SRS implementation minimizes resource usage by using smaller, previously calculated, running sums in the calculation of larger, later running sums, which therefore do not need extra summation values to be stored.
 In addition, it makes use of multipoint shift registers that are configured to give intermediate outputs, referred to as \emph{taps}.
-The taps provide data outputs at certain points in the shift register chain thus contribute towards the efficient use of resources.
+The taps provide data outputs at certain points in the shift register chain, thus contributing to the efficient use of resources.
 
-As in the SRS implementation the sum of one shift register's content is fed to the input of another shift register, the optimal achievable latency in the response of each stage is equal to the refreshing time of the preceding shift registers, i.e.
-the time needed to completely update its contents.
-The \emph{read delay} signal (see Figure~\ref{fig:SRS-basic}) of each shift register holds a delay equal to this latency to guarantee the correct operation.
-Thus, the delay is every time equal to the preceding shift register's delay multiplied by the elements planned to be used in the sum.
+In the SRS implementation, one shift register's sum is fed as input to another shift register.
+Therefore, the best achievable latency of each shift register is equal to the refreshing time of all preceding shift registers, i.e. the time needed to completely update their contents.
+The \emph{read delay} signal (see Figure~\ref{fig:SRS-basic}) of each shift register holds a delay equal to this latency to ensure correct operation.
+The delay is equal to the preceding shift register's delay multiplied by the number of cells to be used in the sum.
 
@@
 
-Figure~\ref{fig:srs} shows the implementation of SRS in BLETC card.
-It consists of 6 \emph{slices}, where each slice computes two running sums (noted by RS) with the use of a multipoint shift register, two subtractors and two accumulators (see Figure~\ref{fig:slice}).
+Figure~\ref{fig:srs} shows the implementation of SRS in a BLETC.
+It consists of 6 \emph{slices}, where each slice computes two running sums (denoted by RS) with the use of a multipoint shift register, two subtractors and two accumulators (see Figure~\ref{fig:slice}).
 
-As shown in Table~\ref{fig:example-bakery}, cascading 6 slices is enough to reach the 100 second upper integration limit requested by the specifications.
+As shown in Table~\ref{fig:example-bakery}, cascading 6 slices is enough to reach the 100 second integration limit requested by the specifications.
 

### 2011-03-30 fmics/paper.tex
Add subsection headings to Section 4

@@
 
+\subsection{Introduction}
 Our formal verification effort uses mechanised logical deduction, or \emph{theorem proving}.
@@
 
+\subsection{Formal Model of SRS}
 The first natural step to prove the correctness statement is to build a logical model of the SRS structure.
@@
 
+\subsection{Theorems about the Model}
 Our central result equates the output of a slice to a sum of consecutive input counts.

### 2011-03-30 fmics/paper.tex
Fix up proof of Theorem 3

@@
 By applying Theorem~\ref{output_eq_exact}, we are left with an equation involving $\mathsf{exact}$ but not $\mathsf{output}$, specifically between $\mathsf{exact}\;D\;n\;x\;t$ and $\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)$.
-Our assumption on $t$ ensures that the case where the function is equal to $0$ in Theorem~\ref{output_eq_exact} or in the definition of $\mathsf{exact}$ never applies.
+Our assumption on $t$ ensures that we are never in the $0$ case of either Theorem~\ref{output_eq_exact} or of the definition of $\mathsf{exact}$.
 Now, $\mathsf{exact}$ is defined as a sum of consecutive values of $D$, so we simply need to bound the difference between one such sum and another earlier one of the same size.
-If at each step the maximum difference is $k$, then the total difference is at most $k$ times the distance between the ends of the two sums, which is the bound stated in the theorem <actually the theorem may have an additional factor increasing the size of the bound, that made it easier to prove - not sure>.
+But if at each step the maximum difference is $k$, then the total difference is at most $k$ times the distance between the ends of the two sums, as required.
+(We use the fact that $\mathsf{last\_update}\;n\;t=t-t\operatorname{mod}\mathsf{delay}\;n$ and, in general, that $a\operatorname{mod}b<b$.)
 \end{proof}

### 2011-03-31 fmics/paper.tex
merge changes from dropbox

@@
 The Large Hadron Collider (LHC) at the European Organization for Nuclear Research (CERN) is a high-energy particle accelerator.
-It provides head-on collisions of protons at a center of a mass energy of 14 TeV for high-energy particle physics research.
+It is designed to provide head-on collisions of protons at a center of a mass energy of 14 TeV for high-energy particle physics research.
 In order to reach the required magnetic field strengths, the LHC has superconducting magnets cooled with superfluid helium.
@@
 The main strategy for protecting the LHC is based on the Beam Loss Monitoring System (BLMS), which triggers the safe extraction of the beams if particle loss exceeds thresholds that are likely to result in a quench.
-At each cycle of the beams around the LHC, the BLMS records and processes several thousands of data points to decide whether the beams should be permitted to continue circulating or whether their safe extraction should be triggered.
+At each cycle of the two counter-rotating beams around the 27 km tunnel of LHC, the BLMS records and processes several thousands of data points to decide whether the beams should be permitted to continue circulating or whether their safe extraction should be triggered.
 The processing includes analysis of the loss pattern over time and of the energy of the beam.
@@
 Any error in the SRS implementation would either compromise the availability of the LHC (unnecessary request for beam dump) or safety (not triggering a necessary beam dump).
-The current approach for analyzing the SRS implementation is simulation of its behavior on sample streams of input~\cite{}.
-{\bf one more sentence needs to be added on the current verification process}
+The current approach for analyzing the SRS implementation relies on the simulation of its behavior on sample streams of input for different loss scenarios~\cite{Chris-thesis}.
 
@@
 One of the challenges in pursuing a formal verification approach for SRS was capturing the intricate details of the system's specification via experiment and refinement with a team of different backgrounds and expertise.
-Our confidence in the SRS design ultimately rests upon our deep understanding of why the design is correct rather than the fact that we obtained ``Theorem Proved" as the final output of a software tool.
+Our confidence in the SRS design as a result of this effort ultimately rests upon our deep understanding of why the design is correct rather than the fact that we obtained ``Theorem Proved" as the final output of a software tool.
 In particular, our use of mechanized logical deduction was a highly iterative process that incrementally refined our understanding of (1) the implementation (2) the intended behavior and (3) the ``white board level" argument or explanation for why the implementation achieves the intended behaviors.
@@
 
-Our contributions in this paper are: a formal model of the SRS component of the BLMS, and commentary on the process and outcomes of taking a formal approach.
+Our contributions in this paper are: a formal model of the SRS component of the BLMS, a formal analysis of its behavior, and commentary on the process and outcomes of taking a formal approach.
 We give an overview of the BLMS in Section~\ref{sec-BLM}, and describe the SRS component in particular in Section~\ref{sec-SRS}.
-In Section~\ref{sec-verification} we describe our approach to formal verification, and present the formal model and results.
+In Section~\ref{sec-verification}, we describe our approach to formal verification, and present the formal model and results.
 Finally, we reflect on the process, summarising lessons learned and future directions, in Section~\ref{sec-conclusions}.
@@
 
-The purpose of the BLMS is to measure particle loss, and to request beam extraction if the loss level indicates that a quench is likely to occur.
-The physical principle underlying particle loss measurement \cite{Dehning-IPAC,Chris-FPGA} is the detection of energy deposited by \emph{shower particles} in \emph{ionization chambers}.
-There are around 4000 ionization chambers strategically placed on the sides of the magnets all around the LHC tunnel underground.
-The ionization chambers produce an electrical signal, based on the recording of shower particles, and transmit the signal to \emph{tunnel cards}.
-Tunnel cards, which are implemented by radiation-tolerant electronics, digitize the data and transmit the digitized data to the surface above the tunnel using optical links.
-At the surface, data processing cards named \emph{BLETCs} receive the data and decide whether or not the beam should be permitted to continue circulating.
-Each tunnel card receives data from eight ionization chambers, and each BLETC receives data from two tunnel cards.
-A BLETC provides data to the Logging, Post Mortem, and Collimation systems that drive on-line displays in the control room, perform long-term storage for offline analysis, and setup the collimators automatically {\bf What is a collimator?
-Why is it necessary to mention all this?}.
-Due to demanding performance requirements, BLETCs are implemented on FPGAs, which include the resources needed to design complex processing and can be reprogrammed making them ideal for future upgrades or system specification changes.
+The main purpose of the BLMS is to measure particle loss, and to request beam extraction if the loss level indicates that a quench is likely to occur.
+The physical principle underlying particle loss measurement \cite{Dehning-IPAC,Chris-FPGA} is the detection of energy deposited by \emph{secondary shower particles} using specially-designed detectors called  \emph{ionization chambers}.
+There are approximately 4000 ionization chambers strategically placed on the sides of the magnets all around the LHC tunnel underground.
+The ionization chambers produce an electrical signal, based on the recording of shower particles, which are readout by \emph{acquisition cards}.
+Acquisition cards, located also in the tunnel and thus implemented by radiation-tolerant electronics, acquire and digitize the data and transmit the digitized data to the surface above the tunnel using optical links.
+At the surface, data processing cards named \emph{BLETCs} receive the data and decide whether or not the beam should be permitted to be injected or continue circulating.
+Each acquisition card receives data from eight ionization chambers, and each BLETC receives data from two acquisition cards.
+A BLETC provides data to the Logging, Post Mortem, and Collimation systems that drive on-line displays in the control room, perform long-term storage for offline analysis, and setup the collimators automatically.
+Due to demanding performance requirements, BLETCs are implemented on FPGAs, which include the resources needed to implement complex processing and can be reprogrammed making them ideal for future upgrades or system specification changes.
+
 Figure~\ref{fig:BLETC} shows a block diagram of the processes on a BLETC FPGA.
@@
 
-\emph{(a) Receive, Check, and Compare (RCC)}: The RCC block receives the data direct from the tunnel, and attempts to detect and correct erroneous transmissions by using Cyclic Redundancy Check and 8B/10B algorithms.{\bf references for these algorithms?}
+\emph{(a) Receive, Check, and Compare (RCC)}: The RCC block receives the data directly from the acquisition card, and attempts to detect and correct erroneous transmissions by using Cyclic Redundancy Check and 8B/10B algorithms~\cite{CRC,8B10B}.
 
 \emph{(b) Data Processing}: Whether or not a quench results from particle loss depends on the loss duration and the beam energy.
-Given the tolerance acceptable for quench prevention, the loss level versus loss duration is approximated by sliding integration windows (called \emph{Running Sums}) fulfilling the tolerance.{\bf not entirely sure what this sentence means}
-In order to achieve the \emph{dynamic range} (domain of variation of losses) requested by the specification, the detectors use both Current-to-Frequency converters and ADC circuitries.{\bf ADC?}
-The data processing block merges these two types of data coming from the detector so as to send a single count value to the SRS block.
+Given the tolerance acceptable for quench prevention, the quench threshold versus loss duration is approximated by a minimum number of sliding integration windows (called \emph{Running Sums}) fulfilling the tolerance.
+In order to achieve the \emph{dynamic range} (domain of variation of losses) requested by the specification, the detectors use both Current-to-Frequency converter and Analogue to Digital convertor circuitries.
+The data processing block merges these two types of data coming from a detector so as to send a single count value to the SRS block.
 The implementation of the SRS block is described in Section~\ref{sec-SRS}.
 
-\emph{(c) Threshold Comparator}: Every running sum needs to be compared to the threshold for the beam energy reading at that moment.
+\emph{(c) Threshold Comparator}: Every running sum needs to be compared to the threshold that was chosen by the beam energy reading at that moment.
 The comparator initiates a beam dump request if any of the running sums is higher than its corresponding threshold.
@@
 The beam energy information is scaled to 32 levels (0.45 to 7 TeV) and each processing module holds data only for those 16 connected detectors.
-Thus a total of 6,144 threshold values need to be held on each card.
+Thus, a total of 6,144 threshold values need to be held on each card.
 
 \emph{(d) Logging, Post Mortem and Collimation}:  To be able to trace back the loss signal development, the BLMS stores the loss measurement data.
-This data is sent over the VME-bus {\bf what's the VME-bus?} for online viewing and storage by the Logging and Post-Mortem systems.
+This data is sent to Logging and Post-Mortem systems for online viewing and storage.
 For the purpose of supervision, the BLMS drives an online event display to show error and status information recorded by the tunnel electronics and the RCC process as well as the maximum loss rates seen by the running sums.
@@
 The number of values in the window is its \emph{integration time}.
-Ideally, we would have an infinite number of windows with lengths covering the whole spectrum of times from 40 micro-seconds (the rate at which count values enter a BLETC card) to 100 seconds, for detecting all losses from ultra-fast up to steady.
-To approximate this ideal with finite resources, the BLMS is given the tolerance acceptable for quench prevention, and the quench threshold versus loss time curve is approximated by the minimum number of windows that meet the tolerance.
+Ideally, we would have an unbounded number of windows with lengths covering the whole spectrum of times from 40 micro-seconds (the rate at which count values enter a BLETC card) to 100 seconds, for detecting all losses from ultra-fast up to steady.
+To approximate this ideal with finite resources, the BLMS is given the tolerance acceptable for quench prevention, and the quench threshold versus loss duration curve is approximated by the minimum number of windows that meet the tolerance.
 
@@
 Long moving windows, that is, windows with large integration times, are required, which means keeping long histories of received count values.
-To accomplish this goal with relatively narrow shift registers, the SRS uses
-consecutive storage of sums of counts.
-Instead of storing all the values needed for a sum, the SRS accumulates many values as a partial sum, thereby using only a fraction of the memory space.
+To accomplish this goal with relatively narrow shift registers, the SRS uses consecutive storage of sums of counts.
+Instead of storing all the values needed for a sum, the SRS accumulates many values as a partial sum, thereby using only a fraction of the otherwise needed memory space.
 The partial sums for a window with a large integration time are chosen so that they also serve as the sums calculated by a window with a smaller intergration time.
@@
 In the SRS implementation, one shift register's sum is fed as input to another shift register.
-Therefore, the best achievable latency of each shift register is equal to the refreshing time of all preceding shift registers, i.e. the time needed to completely update their contents.
+Therefore, the best achievable latency of each shift register is equal to the refreshing time of its preceding shift register, i.e., the time needed to completely update its contents.
 The \emph{read delay} signal (see Figure~\ref{fig:SRS-basic}) of each shift register holds a delay equal to this latency to ensure correct operation.
@@
   \centering \scalebox{0.65}{ \includegraphics{srs-modified.eps}}
-   \caption{The block diagram showing the Successive Running Sums implementation in BLETC card.}
+   \caption{The block diagram showing the implementation of SRS in BLETC.}
   \label{fig:srs}
@@
 
-As shown in Table~\ref{fig:example-bakery}, cascading 6 slices is enough to reach the 100 second integration limit requested by the specifications.
+As shown in Table~\ref{fig:SRS-table}, cascading 6 slices is enough to reach the 100 second integration limit requested by the specifications given by the scientists and engineers who conceived the machine protection strategy for the LHC.
 
@@
   \centering \scalebox{0.8}{ \includegraphics{multi-rs.eps}}
+  \vspace{-0.1in}
    \caption{The block diagram showing the implementation of each slice.}
@@
 \end{tabular}
-\vspace{0.04in}
-\caption{{\bf<missing caption>}
-\label{fig:example-bakery}}
+\vspace{2ex}
+\caption{SRS configuration in BLETC.}
+\label{fig:SRS-table}
 \end{table}
@@
 We use the HOL4 open source software tool~\cite{HOL4,DBLP:conf/tphol/SlindN08}, which was developed initially at the University of Cambridge, but now by an international team.
-HOL4 enables the construction of theories in higher-order logic~\cite{DBLP:journals/jsyml/Church40}, a formal logic with a similar expressive power to set theory that is widely used for formalising hardware and software models and statements about them.
+HOL4 enables the construction of theories in Higher-Order Logic (HOL)~\cite{DBLP:journals/jsyml/Church40}, a formal logic with a similar expressive power to set theory that is widely used for formalising hardware and software models and statements about them.
 The implementation of HOL4 uses Milner's LCF approach~\cite{Milner:1972:LCF:891954}: a small ``kernel'' implementing the primitive rules of the logic, and convenient derived rules and tactics implemented in terms of the kernel.
@@
 \subsection{Formal Model of SRS}
-The first natural step to prove the correctness statement is to build a logical model of the SRS structure.
-We model each building block of the SRS structure that holds a count value -- for example, each cell in each shift register -- as a function in HOL that takes the input stream and current time step as inputs and returns the value the corresponding building block holds at that time.
+The first step to prove the correctness statement is to build a logical model of the SRS structure.
+We model each building block of the SRS structure that holds a count value -- for example, each cell in each shift register -- as a function in HOL\footnote{Here, use of acronym HOL refers to the formal notation that can be parsed by HOL4.
+In this paper, we use a slightly variant of this notation to improve readability.} that takes the input stream and current time step as inputs and returns the value the corresponding building block holds at that time.
 
@@
 This parameterization makes the model more likely to be applicable to future versions of the system, but also fits more naturally into HOL than would a finite model.
-We use natural numbers throughout, rather than, for example, finite words that would more accurately model count values.
+Our model of SRS is at a level of abstraction above the details of circuity that implements basic arithmetic operations.
+We use natural numbers throughout, rather than, for example, finite words that would more accurately model count values.{\bf Do we meed to give a reason why we chose this level of abstraction?}
 
 \begin{table}[t]
-\caption{
-Descriptions of the HOL functions comprising our model of the SRS structure.
-\label{tab:descriptions}
-}
 \begin{tabular}{lp{0.8\textwidth}}
@@
 \end{tabular}
+\vspace{2ex}
+\caption{
+Descriptions of the HOL functions comprising our model of the SRS structure.
+\label{tab:descriptions}
+}
 \end{table}
@@
 
-\begin{figure}
-\caption{
-Definitions of the HOL functions comprising our model of the SRS structure.
-\(\mathsf{tap}\) and \(\mathsf{input}\) are defined to match Figure~\ref{fig:srs}.
-Slice $0$ is a fake slice representing the SRS input; this enables a succinct definition of $\mathsf{source}$.
-\vspace{2ex}
-\label{fig:definitions}}
+Figure~\ref{fig:auxdefs} shows the definition of a set of additional functions required to prove our main results.
+The function $\mathsf{delay\_sum}\;n$ represents the cumulative delay of the preceding shift registers of a slice.
+For example, $\mathsf{delay\_sum}\;4 = \mathsf{delay}\;3 + \mathsf{delay}\;1 + \mathsf{delay}\;0 = 2 +1 +1 = 4$.
+The function $\mathsf{last\_update}\;n\;t$ specifies the time that the $n$-th slice last updated, and $\mathsf{exact}\;D\;n\;x\;t$ computes the exact sum of consecutive input counts, without delay, that $\mathsf{output}\;D\;n\;x\;t$ is supposed to approximate.
+
+\begin{figure}[t]
+\label{fig:definitions}
 \begin{tabular}{ll}
@@
 \end{tabular}
+\vspace{2ex}
+\caption{
+Definitions of the HOL functions comprising our model of the SRS structure.
+\(\mathsf{tap}\) and \(\mathsf{input}\) are defined to match Figure~\ref{fig:srs}.
+Slice $0$ is a virtual slice representing the SRS input; this enables a succinct definition of $\mathsf{source}$.}
 \end{figure}
 
-\begin{figure}
-\caption{
-Definitions of auxiliary HOL functions used while proving theorems about the model of the SRS structure.
+\begin{figure}[t]
 \label{fig:auxdefs}
-}
-\vspace{2ex}
 \begin{tabular}{l}
@@
 \end{tabular}
+\vspace{2ex}
+\caption{
+Definitions of auxiliary HOL functions used while proving theorems about the model of the SRS structure.}
 \end{figure}
-
-Figure~\ref{fig:auxdefs} defines a set of additional functions required to prove our main results.
-The function $\mathsf{delay\_sum}\;n$ represents the cumulative delay of the preceding shift registers of a slice.
-For example, $\mathsf{delay\_sum}\;4 = \mathsf{delay}\;3 + \mathsf{delay}\;1 + \mathsf{delay}\;0 = 2 +1 +1 = 4$.
-The function $\mathsf{last\_update}\;n\;t$ specifies the time that the $n$-th slice last updated, and $\mathsf{exact}\;D\;n\;x\;t$ computes the exact sum of consecutive input counts, without delay, that $\mathsf{output}\;D\;n\;x\;t$ is supposed to approximate.
 
@@
 
-Note that $\mathsf{true\_sum}$ is easily defined in terms of $\mathsf{exact}$, that is $\mathsf{true\_sum}\;D\;n\;t = \mathsf{exact}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)\;t$.
+The function $\mathsf{true\_sum}$ is easily defined in terms of $\mathsf{exact}$, that is $\mathsf{true\_sum}\;D\;n\;t = \mathsf{exact}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)\;t$.
 Since $\mathsf{RS}$ is also similarly defined in terms of the $\mathsf{output}$, Theorem~\ref{output_eq_exact} establishes a relationship between the values of the $\mathsf{RS}$ and $\mathsf{true\_sum}$.
@@
 One of these constraints is a bound on the difference between two consecutive count values.
-Such a bound can be derived from the highest quench level thresholds.
+Such a bound can be inferred from the highest quench level thresholds.
 
@@
 Our assumption on $t$ ensures that we are never in the $0$ case of either Theorem~\ref{output_eq_exact} or of the definition of $\mathsf{exact}$.
-Now, $\mathsf{exact}$ is defined as a sum of consecutive values of $D$, so we simply need to bound the difference between one such sum and another earlier one of the same size.
+The function $\mathsf{exact}$ is defined as a sum of consecutive values of $D$, so we simply need to bound the difference between one such sum and another earlier one of the same size.
 But if at each step the maximum difference is $k$, then the total difference is at most $k$ times the distance between the ends of the two sums, as required.
@@
 \end{proof}
+
+Our HOL4 proof script for verification of SRS consists of 800 lines of code.
+We proved 69 theorems, including 14 definitions.
+In addition, we had to prove approximately 30 generic theorems that were added to HOL4 libraries during this work.
 
@@
 In this case study, we built a parameterized model of the SRS structure and showed that its behavior is correct with respect to its specification.
-It is likely that the configuration of the slices and shift registers will change in the future to accommodate higher precision.
+It is likely that the configuration of the slices and shift registers will need to change in the future as the understanding of the LHC and its weaknesses increases to accommodate more targeted protections
 Thus, the parameterization in our model was crucial to make it applicable to the future upgrades.
@@
 
-The use of mechanized theorem proving to verify the correctness of the SRS behavior complements an intensive verification based on simulation performed at CERN.
-The simulation was mainly based on two scenarios one for recording fast losses and  another for the situation when a loss gradually builds up in the accelerator.
-The results of the simulation showed that the current implementation of SRS, given in Figure~\ref{fig:srs}, satisfies the specification, that is the running sums have  the maximum 20\% relative error.
-On the contrary, the results in this paper apply to all possible input steams, not just a limited number of input streams simulating fast or slow loss scenarios.
+The use of mechanized theorem proving to verify the correctness of the SRS behavior complements an intensive verification effort based on simulation performed at CERN.
+This effort targeted the validity of the SRS as an accurate and fast enough method by analyzing its behavior (a) in the boundaries of the threshold limits and (b) in expected types of beam losses, e.g., fast and steady losses.
+Its results showed that the current implementation of the SRS, given in Figure~\ref{fig:srs}, satisfies the specification, that is the running sums have the maximum 20\% relative error.
+On the contrary, the results in this paper apply to all possible input steams, not just a limited number of input stream scenarios performed at CERN.
+
 But, the error bound given in Theorem~\ref{max_error} does not satisfy the maximum 20\% relative error.
@@
 Therefore, the SRS structure in LHC is an interesting example in which the formal verification does not completely replace conventional verification approaches such as simulation.
+
+HOL4 has highly developed tools  for reasoning about bitstrings (words, bytes etc).
+One possible future research direction os to refine the natural number model to a more hardware-realistic one and even verifying a model extracted from the HDL used to synthesize the FPGA.
 

### 2011-03-31 fmics/paper.tex
revise footnote about HOL

@@
 The first step to prove the correctness statement is to build a logical model of the SRS structure.
-We model each building block of the SRS structure that holds a count value -- for example, each cell in each shift register -- as a function in HOL\footnote{Here, use of acronym HOL refers to the formal notation that can be parsed by HOL4.
-In this paper, we use a slightly variant of this notation to improve readability.} that takes the input stream and current time step as inputs and returns the value the corresponding building block holds at that time.
+We model each building block of the SRS structure that holds a count value -- for example, each cell in each shift register -- as a function in HOL\footnote{The acronym HOL refers to higher-order logic rather than the software tool HOL4.
+We use the tool to define a function in the logic.}
 

### 2011-03-31 fmics/paper.tex
minor tweaks to text in sec 2

@@
 There are approximately 4000 ionization chambers strategically placed on the sides of the magnets all around the LHC tunnel underground.
-The ionization chambers produce an electrical signal, based on the recording of shower particles, which are readout by \emph{acquisition cards}.
-Acquisition cards, located also in the tunnel and thus implemented by radiation-tolerant electronics, acquire and digitize the data and transmit the digitized data to the surface above the tunnel using optical links.
-At the surface, data processing cards named \emph{BLETCs} receive the data and decide whether or not the beam should be permitted to be injected or continue circulating.
+The ionization chambers produce electrical signals, based on the recording of shower particles, which are read out by \emph{acquisition cards}.
+Acquisition cards, also located in the tunnel and therefore implemented by radiation-tolerant electronics, acquire and digitize the data and transmit the digitized data to the surface above the tunnel using optical links.
+At the surface, data processing cards named \emph{BLETCs} receive the data and decide whether or not the beam should be permitted to be injected or to continue circulating.
 Each acquisition card receives data from eight ionization chambers, and each BLETC receives data from two acquisition cards.
@@
 \emph{(b) Data Processing}: Whether or not a quench results from particle loss depends on the loss duration and the beam energy.
-Given the tolerance acceptable for quench prevention, the quench threshold versus loss duration is approximated by a minimum number of sliding integration windows (called \emph{Running Sums}) fulfilling the tolerance.
+Given the tolerance acceptable for quench prevention, the quench threshold versus loss duration is approximated by the minimum number of sliding integration windows (called \emph{Running Sums}) fulfilling the tolerance.
 In order to achieve the \emph{dynamic range} (domain of variation of losses) requested by the specification, the detectors use both Current-to-Frequency converter and Analogue to Digital convertor circuitries.
@@
 
-As shown in Table~\ref{fig:SRS-table}, cascading 6 slices is enough to reach the 100 second integration limit requested by the specifications given by the scientists and engineers who conceived the machine protection strategy for the LHC.
+As shown in Table~\ref{fig:SRS-table}, cascading 6 slices is enough to reach the 100 second integration limit requested by the specifications given by the scientists and engineers who designed the machine protection strategy for the LHC.
 

### 2011-03-31 fmics/paper.tex
mention parenthetically that the first tap position is 0
move some labels into captions, to fix the cross-references

@@
 \hline
-\(\mathsf{tap}\;n\;x\)&The position of tap $x$ of slice $n$.\\
+\(\mathsf{tap}\;n\;x\)&The position of tap $x$ of slice $n$. (The first position is $0$.)\\
 \(\mathsf{input}\;n\)&A pair $(n',x)$ indicating that the input to slice $n$ is output $x$ of slice $n'$.\\
@@
 \begin{figure}[t]
-\label{fig:definitions}
 \begin{tabular}{ll}
@@
 \caption{
+\label{fig:definitions}
 Definitions of the HOL functions comprising our model of the SRS structure.
@@
 \begin{figure}[t]
-\label{fig:auxdefs}
 \begin{tabular}{l}
@@
 \caption{
+\label{fig:auxdefs}
 Definitions of auxiliary HOL functions used while proving theorems about the model of the SRS structure.}

### 2011-03-31 fmics/paper.tex
minor tweak fig 4,5 captions
tweak wordings in a couple of sentences

@@
   \centering \scalebox{0.65}{ \includegraphics{srs-modified.eps}}
-   \caption{The block diagram showing the implementation of SRS in BLETC.}
+   \caption{Block diagram showing the implementation of SRS in a BLETC.}
   \label{fig:srs}
@@
   \vspace{-0.1in}
-   \caption{The block diagram showing the implementation of each slice.}
+   \caption{Block diagram showing the implementation of each slice.}
   \label{fig:slice}
@@
 
-The function $\mathsf{true\_sum}$ is easily defined in terms of $\mathsf{exact}$, that is $\mathsf{true\_sum}\;D\;n\;t = \mathsf{exact}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)\;t$.
-Since $\mathsf{RS}$ is also similarly defined in terms of the $\mathsf{output}$, Theorem~\ref{output_eq_exact} establishes a relationship between the values of the $\mathsf{RS}$ and $\mathsf{true\_sum}$.
+The function $\mathsf{true\_sum}$ is easily defined in terms of $\mathsf{exact}$, namely, $\mathsf{true\_sum}\;D\;n\;t = \mathsf{exact}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)\;t$.
+Since $\mathsf{RS}$ is similarly defined in terms of the $\mathsf{output}$, Theorem~\ref{output_eq_exact} establishes a relationship between the values of $\mathsf{RS}$ and of $\mathsf{true\_sum}$.
 

### 2011-03-31 fmics/paper.tex
add some sentences about difficulties, despite slick proof summaries

@@
 
-Our HOL4 proof script for verification of SRS consists of 800 lines of code.
+The proof of Theorem~\ref{max_error} is straightforward when summarized, as above.
+However, as for all of our theorems, there are some tricky details underlying the high-level summary.
+For example, \emph{proving} that the maximum difference between the end terms in a sum is the maximum difference between consecutive elements multiplied by the number of terms requires an inductive argument.
+Our HOL4 proof script for verification of SRS consists of 750 lines of code.
 We proved 69 theorems, including 14 definitions.

### 2011-03-31 fmics/paper.tex
minor tweaks in section 5

@@
 
-We described a case study in which HOL4 theorem prover is used to formally verify properties of the SRS structure within the LHC machine protection system.
+We described a case study in which the HOL4 theorem prover is used to formally verify properties of the SRS structure within the LHC machine protection system.
 In this case study, we built a parameterized model of the SRS structure and showed that its behavior is correct with respect to its specification.
@@
 However, some sanity tests related to Theorem 1 led us astray when we did not know the correct formula for that theorem.
-Because the formulae we conjectured worked for small values of $n$, and simulating large values of $n$ was expensive, but the counterexamples were only on large values of $n$.
-The most important use of mechanized logical deduction in this effort was its role as an ``implacable skeptic" that insisted on us providing the justification and compelled us to clarify the intricate details of the SRS structure and behavior.
+The formulae we conjectured worked for small values of $n$, and simulating large values of $n$ was expensive, but the counterexamples were only on large values of $n$.
+The most important use of mechanized logical deduction in this effort was its role as an ``implacable skeptic" that insisted we provide justification and compelled us to clarify the intricate details of the SRS structure and behavior.
 
@@
 This effort targeted the validity of the SRS as an accurate and fast enough method by analyzing its behavior (a) in the boundaries of the threshold limits and (b) in expected types of beam losses, e.g., fast and steady losses.
-Its results showed that the current implementation of the SRS, given in Figure~\ref{fig:srs}, satisfies the specification, that is the running sums have the maximum 20\% relative error.
+Its results showed that the current implementation of the SRS, given in Figure~\ref{fig:srs}, satisfies the specification, that is the running sums have a maximum 20\% relative error.
 On the contrary, the results in this paper apply to all possible input steams, not just a limited number of input stream scenarios performed at CERN.
 
-But, the error bound given in Theorem~\ref{max_error} does not satisfy the maximum 20\% relative error.
+But, the error bound given in Theorem~\ref{max_error} does not satisfy the maximum 20\% relative error bound.
 In order to make this bound tighter, we need to provide more characterization of the input as constraints on $D$.
 Due to the nature of the problem, such a characterization of input is not trivial.
-Therefore, the SRS structure in LHC is an interesting example in which the formal verification does not completely replace conventional verification approaches such as simulation.
+Therefore, the SRS structure in LHC is an interesting example in which formal verification does not completely replace conventional verification approaches such as simulation.
 

### 2011-03-31 fmics/paper.tex
merge dropbox changes

@@
 Given the tolerance acceptable for quench prevention, the quench threshold versus loss duration is approximated by the minimum number of sliding integration windows (called \emph{Running Sums}) fulfilling the tolerance.
-In order to achieve the \emph{dynamic range} (domain of variation of losses) requested by the specification, the detectors use both Current-to-Frequency converter and Analogue to Digital convertor circuitries.
+In order to achieve the \emph{dynamic range} (domain of variation of losses) required by the specification, the detectors use both Current-to-Frequency converter and Analogue to Digital convertor circuitries.
 The data processing block merges these two types of data coming from a detector so as to send a single count value to the SRS block.
@@
 Figure~\ref{fig:srs} shows the implementation of SRS in a BLETC.
-It consists of 6 \emph{slices}, where each slice computes two running sums (denoted by RS) with the use of a multipoint shift register, two subtractors and two accumulators (see Figure~\ref{fig:slice}).
+It consists of 6 \emph{slices}, where each slice computes two running sums (e.g., slice 4 computes the fourth and fifth running sums: RS6 and RS7) with the use of a multipoint shift register, two subtractors and two accumulators (see Figure~\ref{fig:slice}).
 
-As shown in Table~\ref{fig:SRS-table}, cascading 6 slices is enough to reach the 100 second integration limit requested by the specifications given by the scientists and engineers who designed the machine protection strategy for the LHC.
+As shown in Table~\ref{fig:SRS-table}, cascading 6 slices is enough to reach the 100 second integration limit required by the specifications given by the scientists and engineers who designed the machine protection strategy for the LHC.
 
@@
 The multi-layered structure of the SRS and the read delay of each shift register result in the outputs of SRS being delayed from the $\mathsf{true\_sum}$ values.
-A sketch of the desired correctness statement is \[\forall{n} \cdot \, \mathsf{RS}\;n= \mathsf{true\_sum}\; n \pm\text{acceptable error}\]
+A sketch of the desired correctness statement is: \[\forall{n} \cdot \, \mathsf{RS}\;n= \mathsf{true\_sum}\; n \pm\text{acceptable error}\]
 Although Figure~\ref{fig:srs} suggests that the SRS structure has only twelve outputs  (i.e.
@@
 This parameterization makes the model more likely to be applicable to future versions of the system, but also fits more naturally into HOL than would a finite model.
-Our model of SRS is at a level of abstraction above the details of circuity that implements basic arithmetic operations.
-We use natural numbers throughout, rather than, for example, finite words that would more accurately model count values.{\bf Do we meed to give a reason why we chose this level of abstraction?}
+We defined our formal model to be at a level of abstraction above the details of circuity that implements basic arithmetic operations.
+We use natural numbers throughout, rather than, for example, finite words that would more accurately model count values.
+This level of abstraction is sufficient for the SRS correctness properties we are interested in.
+A separate verification effort can focus on showing that a more realistic model of hardware circuitry accurately implements natural number arithmetic.
 
@@
 We formulate the difference between the $\mathsf{RS}$ values and their corresponding $\mathsf{true\_sum}$s, caused by the delay, as $\mathsf{error}$.
-In order to define a maximum bound on the $\mathsf{error}$, we had to formulate some characterization of the input stream to the SRS structure as constraints on $D$.
-One of these constraints is a bound on the difference between two consecutive count values.
-Such a bound can be inferred from the highest quench level thresholds.
+To meet the tolerance acceptable for quench prevention, the specification of SRS requires a bound on the value of this $\mathsf{error}$.
+Considering any possible input stream to SRS, such a bound on the $\mathsf{error}$ value will be infinite.
+However, according to extensive experimental analysis, the beam loss over time follows some patterns.
+Thus, formulating some characterization of the input stream to SRS structure as constraints on $D$, will help defining a tighter bound on the $\mathsf{error}$.
+One of these constraints is a maximum value on the difference between two consecutive count values.
+This value can be inferred from the highest quench level thresholds.
 
@@
 The proof of Theorem~\ref{max_error} is straightforward when summarized, as above.
-However, as for all of our theorems, there are some tricky details underlying the high-level summary.
+However, as for all of our theorems, there are several non-trivial details underlying the high-level summary provided.
 For example, \emph{proving} that the maximum difference between the end terms in a sum is the maximum difference between consecutive elements multiplied by the number of terms requires an inductive argument.
@@
 One possible future research direction os to refine the natural number model to a more hardware-realistic one and even verifying a model extracted from the HDL used to synthesize the FPGA.
+
+The general method (I think Mike mentioned) of subjecting a formal
+model to "obviousness" tests, collecting small theorems about it along
+the way, seems to be a good one.
 

### 2011-03-31 fmics/paper.tex
Reword paragraph before Theorem 3

@@
 We formulate the difference between the $\mathsf{RS}$ values and their corresponding $\mathsf{true\_sum}$s, caused by the delay, as $\mathsf{error}$.
-To meet the tolerance acceptable for quench prevention, the specification of SRS requires a bound on the value of this $\mathsf{error}$.
-Considering any possible input stream to SRS, such a bound on the $\mathsf{error}$ value will be infinite.
+To meet the tolerance acceptable for quench prevention, the SRS specification requires a bound on this $\mathsf{error}$.
+Without restricting the input stream, the $\mathsf{error}$ is unbounded\footnote{For any bound, we can construct an input stream that causes the bound to be exceeded by, for example, having a sequence of zeroes followed by a sequence of count values much higher than the bound.}.
 However, according to extensive experimental analysis, the beam loss over time follows some patterns.
-Thus, formulating some characterization of the input stream to SRS structure as constraints on $D$, will help defining a tighter bound on the $\mathsf{error}$.
-One of these constraints is a maximum value on the difference between two consecutive count values.
-This value can be inferred from the highest quench level thresholds.
+By formulating some characterization of the input stream as constraints on $D$, we can recover a bound on the $\mathsf{error}$.
+One of these constraints is a maximum value for the difference between two consecutive count values.
+This ``maximum jump size'' can be inferred from the highest quench level thresholds.
 

### 2011-03-31 fmics/paper.tex
more dropbox changes

--- 2011-03-31 fmics/paper.tex
@@
 In this case study, we built a parameterized model of the SRS structure and showed that its behavior is correct with respect to its specification.
-It is likely that the configuration of the slices and shift registers will need to change in the future as the understanding of the LHC and its weaknesses increases to accommodate more targeted protections
+It is likely that the configuration of the slices and shift registers will need to change in the future as the understanding of the LHC and its weaknesses increases to accommodate more targeted protections.
 Thus, the parameterization in our model was crucial to make it applicable to the future upgrades.
@@
 This effort targeted the validity of the SRS as an accurate and fast enough method by analyzing its behavior (a) in the boundaries of the threshold limits and (b) in expected types of beam losses, e.g., fast and steady losses.
-Its results showed that the current implementation of the SRS, given in Figure~\ref{fig:srs}, satisfies the specification, that is the running sums have a maximum 20\% relative error.
-On the contrary, the results in this paper apply to all possible input steams, not just a limited number of input stream scenarios performed at CERN.
+Its results showed that the current implementation of the SRS, given in Figure~\ref{fig:srs}, satisfies the specification, that is the running sums have a maximum 20\% relative error ($|\mathsf{RS}\;D\;n\;t - \mathsf{true\_sum}\;D\;n\;t| < 20\%(\mathsf{true\_sum}))$.
 
-But, the error bound given in Theorem~\ref{max_error} does not satisfy the maximum 20\% relative error bound.
-In order to make this bound tighter, we need to provide more characterization of the input as constraints on $D$.
-Due to the nature of the problem, such a characterization of input is not trivial.
-Therefore, the SRS structure in LHC is an interesting example in which formal verification does not completely replace conventional verification approaches such as simulation.
+On the contrary, the results in this paper apply to a larger set of input streams, not just a limited number of sample inputs considered at CERN.
+However, the error bound proved in Theorem~\ref{max_error} is not tight enough to satisfy the maximum 20\% relative error.
+One of the potential reasons for this problem is the lack of knowledge to characterize the input stream to the SRS.
+Due to the nature of this problem, such a characterization of input is not trivial.
+While testing and simulation of the SRS on a limited set of input streams offer a level of confidence that the behavior of SRS satisfies this particular specification, constructing a formal proof for all possible behaviors requires a better understanding of the characteristics of the input stream.
+One future research direction is to investigate whether constructing such a formal proof is feasible for the SRS component of BLMS.
 
-HOL4 has highly developed tools  for reasoning about bitstrings (words, bytes etc).
-One possible future research direction os to refine the natural number model to a more hardware-realistic one and even verifying a model extracted from the HDL used to synthesize the FPGA.
-
-The general method (I think Mike mentioned) of subjecting a formal
-model to "obviousness" tests, collecting small theorems about it along
-the way, seems to be a good one.
-
-%Future direction, go with lower level implementation rather than finite words (closer  to vhdl coed) and improve the bound
+Another future research direction is to refine our formal model to a lower level of abstraction and closer to the hardware model (e.g., representing the count values by words instead of natural number).
+We are also considering verifying a model extracted from the Hardware Description Language (HDL) used to synthesize the BLETC FPGA.
 

### 2011-03-31 fmics/paper.tex
more dropbox changes

--- 2011-03-31 fmics/paper.tex
@@
 One of our main challenges in this effort was building the formal model and formulating the correctness statements.
+There are three different sources of complexity that are inherent elements of understanding why the SRS structure implements its intended behavior:
+(a) the structure of the SRS; although it features considerable regularity, it includes exceptions to this regularity that complicate reasoning, e.g., the input of one slice may not be the output of the immediately previous slice, but may be the first or second output of any previous slice or even the global input;
+(b) the non-trivial timing relationships between different elements of the SRS; e.g., the frequency of updates to the contents of a shift register depends on its position of the shift register in the layered structure of the SRS;
+(c) arithmetic relationships that are not always intuitively apparent at first glance.
+While each of these sources of complexity is manageable on its own, reasoning about the correctness of the SRS is matter of grappling with all three sources of complexity at once.
+{\bf Ramana can you add an example here?}
+
 To understand the structure of the SRS, we started with a model of a simplified structure, which had a more regular arrangements of the slices and ignored the middle taps, and proceeded to a model which is closer to the SRS implementation in BLMS.
@@
 The most important use of mechanized logical deduction in this effort was its role as an ``implacable skeptic" that insisted we provide justification and compelled us to clarify the intricate details of the SRS structure and behavior.
+{\bf not sure if we should keep this paragraph after adding the previous one}
 

### 2011-04-01 fmics/paper.tex
merge dropbox changes to paper.tex

@@
 
-In this paper, we describe a formal verification approach based on logical deduction for analyzing the SRS implementation.
+In this paper, we describe a formal verification approach based on logical deduction, using HOL4 theorem prover~\cite{HOL4,DBLP:conf/tphol/SlindN08}, for analyzing the SRS implementation.
 Our high level proof strategy takes advantage of the regular structure of the SRS, which consists of multiple layers of shift registers and some simple arithmetic hardware.
@@
 Figure~\ref{fig:BLETC} shows a block diagram of the processes on a BLETC FPGA.
-In the following, we briefly describe each of the four processing blocks on a BLETC card.
+In the following, we briefly describe each of the main four processing blocks on a BLETC card.
 
@@
 In order to achieve the \emph{dynamic range} (domain of variation of losses) required by the specification, the detectors use both Current-to-Frequency converter and Analogue to Digital convertor circuitries.
-The data processing block merges these two types of data coming from a detector so as to send a single count value to the SRS block.
+The data processing block merges these two types of data coming from a detector so as to send a single value, referred to as \emph{count}, to the SRS block.
 The implementation of the SRS block is described in Section~\ref{sec-SRS}.
@@
 
-Processing the data collected by the detectors, referred to as \emph{count} values, involves an analysis of the loss pattern over time, accounting for the energy of the beam.
+Processing the data collected by the detectors involves an analysis of the loss pattern over time, accounting for the energy of the beam.
 The processing procedure is based on the idea that a constantly updated moving window can be maintained in an accumulator by adding the incoming (newest) value and subtracting the oldest value (see Figure~\ref{fig:RS-basic}).
 The number of values in the window is its \emph{integration time}.
-Ideally, we would have an unbounded number of windows with lengths covering the whole spectrum of times from 40 micro-seconds (the rate at which count values enter a BLETC card) to 100 seconds, for detecting all losses from ultra-fast up to steady.
+Ideally, we would have an unbounded number of windows with lengths covering the whole spectrum of times from 40 micro-seconds (the rate at which data from detectors enter a BLETC card) to 100 seconds, for detecting all losses from ultra-fast up to steady.
 To approximate this ideal with finite resources, the BLMS is given the tolerance acceptable for quench prevention, and the quench threshold versus loss duration curve is approximated by the minimum number of windows that meet the tolerance.
@@
 Figure~\ref{fig:srs} shows the implementation of SRS in a BLETC.
-It consists of 6 \emph{slices}, where each slice computes two running sums (e.g., slice 4 computes the fourth and fifth running sums: RS6 and RS7) with the use of a multipoint shift register, two subtractors and two accumulators (see Figure~\ref{fig:slice}).
+It consists of 6 \emph{slices}, where each slice computes two running sums (e.g., slice 4 computes the sixth and seventh running sums: RS6 and RS7) with the use of a multipoint shift register, two subtractors and two accumulators (see Figure~\ref{fig:slice}).
 
-As shown in Table~\ref{fig:SRS-table}, cascading 6 slices is enough to reach the 100 second integration limit required by the specifications given by the scientists and engineers who designed the machine protection strategy for the LHC.
+As shown in Table~\ref{fig:SRS-table}, cascading 6 slices is enough to reach the approximately 100 second integration limit required by the specifications given by the scientists and engineers who designed the machine protection strategy for the LHC.
 
@@
 
+In this section, we describe our approach to formal verification of the SRS component in BLETC and present the formal model and results.
+
 \subsection{Introduction}
@@
 
+Prior to defining a formal model of the SRS and the theorems on the model, we developed an informal ``white board"-level argument for why the SRS implements its intended behavior.
+The essence of this argument relies on four facts which can be established from the structure of the SRS and some details about timing relationship between layers:
+\begin{enumerate}
+\item The shift register of each slice (except the first) is updated once the content of the shift register of its previous slice is completely updated.
+\item The integration window of a given shift register in a slice (except the first slice) can be decomposed into a sequence of  non-overlapping consecutive segments, $S_1, S_2, \ldots, S_w$ (where $w$ is the width of the shift register) each of which is equal to the size of the integration window of the shift register of the previous slice.
+\item After a period of initialization, the values stored in the shift register of a given slice (except the first) are the $w$ outputs of the previous slice, where $w$ is the width of the shift register of this slice.
+\item After a period of initialization, the output of each slice is always equal to the sum of the contents of its shift register.
+\end{enumerate}
+
+Using Facts 1, 2 and 3 in an inductive argument, we show that each cell of the shift register of a given slice, after a period of initialization, always contains the sum of the SRS inputs for one of the non-overlapping consecutive segments that make up the integration window of this slice.
+Then using Fact 4 and arithmetic reasoning, we can show that the output of this slice, after a period of initialization, contains the sum of the SRS inputs over its integration window.
+
+While the above paragraph gives the appearance of a straightforward argument for the correctness of the SRS (corresponding roughly to Theorem~\ref{output_input_at_update_times} in Section~\ref{subsec:theorems}), in fact the argument involves consideration of many details that arise from the formal model of the SRS presented in the next section.
+{\bf one possibility is to move the paragraph about sources of complexities currently in section 5 to here.}
+
 \subsection{Formal Model of SRS}
@@
 We model each building block of the SRS structure that holds a count value -- for example, each cell in each shift register -- as a function in HOL\footnote{The acronym HOL refers to higher-order logic rather than the software tool HOL4.
-We use the tool to define a function in the logic.}
+We use the tool to define a function in the logic.}.
 
@@
 We use natural numbers throughout, rather than, for example, finite words that would more accurately model count values.
-This level of abstraction is sufficient for the SRS correctness properties we are interested in.
+This level of abstraction is sufficient to prove our desired correctness properties of SRS.
 A separate verification effort can focus on showing that a more realistic model of hardware circuitry accurately implements natural number arithmetic.
@@
 \(\mathsf{source}\;D\;n\;m\;t\)&The value of the cell that is the direct input to cell $m$ of slice $n$, at time $t$, given input stream $D$.\\
-\(\mathsf{SR}\;D\;n\;m\;t\)&The value of cell $m$ of shift register $n$, at time $t$, given input stream $D$.\\
+\(\mathsf{SR}\;D\;n\;m\;t\)&The value of cell $m$ of slice $n$, at time $t$, given input stream $D$.\\
 \(\mathsf{output}\;D\;n\;x\;t\)&The output at tap $x$ of slice $n$, at time $t$, given input stream $D$.\\
@@
 \multicolumn{2}{l}{\;$\mathsf{source}\;D\;n\;0\;t=\mathsf{output}\;D\;n'\;x\;t\quad\text{where $(n',x)=\mathsf{input\;n}$}$}\\
-\multicolumn{2}{l}{$\;\mathsf{source}\;D\;n\;m\;t=\;\mathsf{SR}\;D\;n\;(m-1)\;\quad\text{(where $m>0$)}$}\\[1ex]
+\multicolumn{2}{l}{$\;\mathsf{source}\;D\;n\;m\;t=\;\mathsf{SR}\;D\;n\;(m-1)\;t\quad\text{(where $m>0$)}$}\\[1ex]
 
@@
 \multicolumn{2}{l}{$\;\mathsf{output}\;D\;n\;x\;t = \text{if $\mathsf{update\_time}\;n\;t$ then }$}\\
-\multicolumn{2}{l}{$\qquad\quad((\mathsf{output}\;D\;n\;x\;(t-1))+(\mathsf{source}\;D\;n\;0\;(t-1)))-{}$}\\
-&$\qquad\qquad\qquad(\mathsf{SR}\;D\;n\;(\mathsf{tap}\;n\;x)\;(t-1))$\\
+\multicolumn{2}{l}{$\qquad\quad((\mathsf{output}\;D\;n\;x\;(t-1))+(\mathsf{source}\;D\;n\;0\;(t-1)))-{}$}
+$(\mathsf{SR}\;D\;n\;(\mathsf{tap}\;n\;x)\;(t-1))$\\
 \multicolumn{2}{l}{$\qquad\qquad\qquad\qquad\text{else }\mathsf{output}\;D\;n\;x\;(t-1)\qquad\text{(where $n,t>0$)}$}\\[1ex]
@@
 \subsection{Theorems about the Model}
+\label{subsec:theorems}
 Our central result equates the output of a slice to a sum of consecutive input counts.
@@
 By formulating some characterization of the input stream as constraints on $D$, we can recover a bound on the $\mathsf{error}$.
-One of these constraints is a maximum value for the difference between two consecutive count values.
+One of these constraints is a maximum value for the difference between two consecutive input count values.
 This ``maximum jump size'' can be inferred from the highest quench level thresholds.
@@
 But if at each step the maximum difference is $k$, then the total difference is at most $k$ times the distance between the ends of the two sums, as required.
-(We use the fact that $\mathsf{last\_update}\;n\;t=t-t\operatorname{mod}\mathsf{delay}\;n$ and, in general, that $a\operatorname{mod}b<b$.)
+(We use the fact that $\mathsf{last\_update}\;n\;t=t-t\operatorname{mod}\mathsf{delay}\;n$ and, in general, that $a\operatorname{mod}b<b$.) \qed
 \end{proof}
@@
 (c) arithmetic relationships that are not always intuitively apparent at first glance.
+{\bf Ramana can you add an example here?}
 While each of these sources of complexity is manageable on its own, reasoning about the correctness of the SRS is matter of grappling with all three sources of complexity at once.
-{\bf Ramana can you add an example here?}
 
@@
 The most important use of mechanized logical deduction in this effort was its role as an ``implacable skeptic" that insisted we provide justification and compelled us to clarify the intricate details of the SRS structure and behavior.
-{\bf not sure if we should keep this paragraph after adding the previous one}
+{\bf not sure if we should keep this paragraph after adding the previous one, maybe should be combined with the previous}
 

### 2011-04-01 fmics/paper.tex
minor revisions in section 1

@@
 The SRS component has a complex structure and the correctness of its behaviour is critical for safe and productive use of the LHC.
-Any error in the SRS implementation would either compromise the availability of the LHC (unnecessary request for beam dump) or safety (not triggering a necessary beam dump).
-The current approach for analyzing the SRS implementation relies on the simulation of its behavior on sample streams of input for different loss scenarios~\cite{Chris-thesis}.
+Any error in the SRS implementation would compromise either the availability of the LHC (unnecessary request for a beam dump) or its safety (not triggering a necessary beam dump).
+The current approach for analyzing the SRS implementation is simulation of its behavior on sample streams of input for different loss scenarios~\cite{Chris-thesis}.
 
-In this paper, we describe a formal verification approach based on logical deduction, using HOL4 theorem prover~\cite{HOL4,DBLP:conf/tphol/SlindN08}, for analyzing the SRS implementation.
+In this paper, we describe a formal verification approach, based on logical deduction using HOL4 theorem prover~\cite{HOL4,DBLP:conf/tphol/SlindN08}, to analyzing the SRS implementation.
 Our high level proof strategy takes advantage of the regular structure of the SRS, which consists of multiple layers of shift registers and some simple arithmetic hardware.

### 2011-04-01 fmics/paper.tex
minor tweaks in sections 2 and 3

@@
 Figure~\ref{fig:BLETC} shows a block diagram of the processes on a BLETC FPGA.
-In the following, we briefly describe each of the main four processing blocks on a BLETC card.
+In the following, we briefly describe each of the four main processing blocks on a BLETC card.
 
-\emph{(a) Receive, Check, and Compare (RCC)}: The RCC block receives the data directly from the acquisition card, and attempts to detect and correct erroneous transmissions by using Cyclic Redundancy Check and 8B/10B algorithms~\cite{CRC,8B10B}.
+\emph{(a) Receive, Check, and Compare (RCC)}: The RCC block receives data directly from the acquisition cards, and attempts to detect and correct erroneous transmissions by using Cyclic Redundancy Check and 8B/10B algorithms~\cite{CRC,8B10B}.
 
@@
 Given the tolerance acceptable for quench prevention, the quench threshold versus loss duration is approximated by the minimum number of sliding integration windows (called \emph{Running Sums}) fulfilling the tolerance.
-In order to achieve the \emph{dynamic range} (domain of variation of losses) required by the specification, the detectors use both Current-to-Frequency converter and Analogue to Digital convertor circuitries.
-The data processing block merges these two types of data coming from a detector so as to send a single value, referred to as \emph{count}, to the SRS block.
+In order to achieve the required \emph{dynamic range} (domain of variation of losses), the detectors use both Current-to-Frequency converter and Analogue to Digital convertor circuitries.
+The data processing block merges these two types of data coming from a detector so as to send a single value, referred to as a \emph{count}, to the SRS block.
 The implementation of the SRS block is described in Section~\ref{sec-SRS}.
 
-\emph{(c) Threshold Comparator}: Every running sum needs to be compared to the threshold that was chosen by the beam energy reading at that moment.
+\emph{(c) Threshold Comparator}: Every running sum needs to be compared to the threshold determined by the beam energy reading at that moment.
 The comparator initiates a beam dump request if any of the running sums is higher than its corresponding threshold.
@@
 \emph{(d) Logging, Post Mortem and Collimation}:  To be able to trace back the loss signal development, the BLMS stores the loss measurement data.
-This data is sent to Logging and Post-Mortem systems for online viewing and storage.
+This data is sent to the Logging and Post-Mortem systems for online viewing and storage.
 For the purpose of supervision, the BLMS drives an online event display to show error and status information recorded by the tunnel electronics and the RCC process as well as the maximum loss rates seen by the running sums.
@@
 
-Beam losses can happen at different rates, compared to the number of cycles of the beam around the tunnel.
+Beam losses can happen at different rates, compared to the number of cycles of the beams around the tunnel.
 One-cycle failures are called \emph{ultra-fast} losses.
@@
 Figure~\ref{fig:srs} shows the implementation of SRS in a BLETC.
-It consists of 6 \emph{slices}, where each slice computes two running sums (e.g., slice 4 computes the sixth and seventh running sums: RS6 and RS7) with the use of a multipoint shift register, two subtractors and two accumulators (see Figure~\ref{fig:slice}).
+It consists of 6 \emph{slices}, where each slice computes two running sums (e.g., slice 4 computes running sums RS6 and RS7) with the use of a multipoint shift register, two subtractors and two accumulators (see Figure~\ref{fig:slice}).
 

### 2011-04-01 fmics/paper.tex
start minor revisions of sec 4

@@
 
-In this section, we describe our approach to formal verification of the SRS component in BLETC and present the formal model and results.
+In this section, we describe our approach to formal verification of the SRS component of a BLETC, and present the formal model and results.
 
@@
 Our goal is to build a generic model of the SRS structure, and to prove that it satisfies its specification, that it calculates approximate running sums of received count values within acceptable error margins.
-Let $\mathsf{RS}\;n$ denote, as in Figure~\ref{fig:srs}, the $n$-th output of the SRS structure, which is supposed to compute a sum of received count values, and let $\mathsf{true\_sum}\;n$ denote this sum.
-The multi-layered structure of the SRS and the read delay of each shift register result in the outputs of SRS being delayed from the $\mathsf{true\_sum}$ values.
+Let $\mathsf{RS}\;n$ denote, as in Figure~\ref{fig:srs}, output $n$ of the SRS structure, which is supposed to compute a sum of received count values, and let $\mathsf{true\_sum}\;n$ denote this sum.
+The multi-layered structure of the SRS and the read delay of each shift register result in the outputs being delayed from the $\mathsf{true\_sum}$ values.
 A sketch of the desired correctness statement is: \[\forall{n} \cdot \, \mathsf{RS}\;n= \mathsf{true\_sum}\; n \pm\text{acceptable error}\]
@@
 
-Prior to defining a formal model of the SRS and the theorems on the model, we developed an informal ``white board"-level argument for why the SRS implements its intended behavior.
-The essence of this argument relies on four facts which can be established from the structure of the SRS and some details about timing relationship between layers:
+Prior to defining a formal model of the SRS and the theorems on the model, we developed an informal ``white board-level'' argument for why the SRS implements its intended behavior.
+The essence of this argument relies on four facts which can be established from the structure of the SRS and some details about the timing relationship between layers:
 \begin{enumerate}

### 2011-04-01 fmics/paper.tex
some revisions in 4.1

@@
 Although Figure~\ref{fig:srs} suggests that the SRS structure has only twelve outputs  (i.e.
-$0\leq{n}\leq11$), we obtain a more generic result (that is useful in future upgrades of the system) by proving the correctness of the above statement for all values of $n$.
+$0\leq{n}\leq11$), we obtain a more generic result (that is useful in future upgrades of the system) by interpreting and proving the statement above for all values of $n$.
 
@@
 
-Prior to defining a formal model of the SRS and the theorems on the model, we developed an informal ``white board-level'' argument for why the SRS implements its intended behavior.
+Prior to defining a formal model of the SRS and proving theorems about it, we developed an informal ``white board-level'' argument for why the SRS implements its intended behavior.
 The essence of this argument relies on four facts which can be established from the structure of the SRS and some details about the timing relationship between layers:
 \begin{enumerate}
-\item The shift register of each slice (except the first) is updated once the content of the shift register of its previous slice is completely updated.
+\item The shift register of each slice (except the first) is updated once the shift register of its previous slice is completely updated.
 \item The integration window of a given shift register in a slice (except the first slice) can be decomposed into a sequence of non-overlapping consecutive segments, $S_1, S_2, \ldots, S_w$ (where $w$ is the width of the shift register) each of which is equal to the size of the integration window of the shift register of the previous slice.
@@
 While the above paragraph gives the appearance of a straightforward argument for the correctness of the SRS (corresponding roughly to Theorem~\ref{output_input_at_update_times} in Section~\ref{subsec:theorems}), in fact the argument involves consideration of many details that arise from the formal model of the SRS presented in the next section.
+Using a software theorem prover enables us to keep track of these details without losing sight of the overall goal.
 {\bf one possibility is to move the paragraph about sources of complexities currently in section 5 to here.}
@@
 The formal definitions of the functions listed in Table~\ref{tab:descriptions} are given in Figure~\ref{fig:definitions}.
-Note that the definition of $\mathsf{input}$ when $n=0$ or when $n>6$ does not change the structure presented.
+The definition of $\mathsf{input}$ when $n=0$ or when $n>6$ does not change the structure represented, since slice $0$ is a virtual slice and the real SRS has only six slices, so we give definitions convenient for theorem proving.
 A similar comment applies to other definitions when $n$, representing a slice number, is $0$, or when $x$, representing a shift register position, is greater than $1$.

### 2011-04-01 fmics/paper.tex
try to keep figs 6 and 7 on the same page
minor revision sec 4

@@
 
-\begin{figure}[t]
+\begin{figure}[p]
 \begin{tabular}{ll}
@@
 
-\begin{figure}[t]
+\begin{figure}[p]
 \begin{tabular}{l}
@@
 However, as for all of our theorems, there are several non-trivial details underlying the high-level summary provided.
-For example, \emph{proving} that the maximum difference between the end terms in a sum is the maximum difference between consecutive elements multiplied by the number of terms requires an inductive argument.
+For example, \emph{proving} that the maximum difference between the end terms in a sum is the maximum difference between consecutive terms multiplied by the number of terms requires an inductive argument.
 Our HOL4 proof script for verification of SRS consists of 750 lines of code.

### 2011-04-01 fmics/paper.tex
minor revisions in sec 5

@@
 (a) the structure of the SRS; although it features considerable regularity, it includes exceptions to this regularity that complicate reasoning, e.g., the input of one slice may not be the output of the immediately previous slice, but may be the first or second output of any previous slice or even the global input;
-(b) the non-trivial timing relationships between different elements of the SRS; e.g., the frequency of updates to the contents of a shift register depends on its position of the shift register in the layered structure of the SRS;
+(b) the non-trivial timing relationships between different elements of the SRS; e.g., the frequency of updates to the contents of a shift register depends on the position of the shift register in the layered structure of the SRS;
 (c) arithmetic relationships that are not always intuitively apparent at first glance.
@@
 
-To understand the structure of the SRS, we started with a model of a simplified structure, which had a more regular arrangements of the slices and ignored the middle taps, and proceeded to a model which is closer to the SRS implementation in BLMS.
+To understand the structure of the SRS, we started with a model of a simplified structure, which had a more regular arrangement of the slices and ignored the middle taps, and proceeded to a model which is closer to the SRS implementation in BLMS.
 In addition, we used a basic spreadsheet simulation model for a few slices for sanity testing the correctness formulae.
 However, some sanity tests related to Theorem 1 led us astray when we did not know the correct formula for that theorem.
-The formulae we conjectured worked for small values of $n$, and simulating large values of $n$ was expensive, but the counterexamples were only on large values of $n$.
-The most important use of mechanized logical deduction in this effort was its role as an ``implacable skeptic" that insisted we provide justification and compelled us to clarify the intricate details of the SRS structure and behavior.
+The formulae we conjectured worked for small values of $n$, and simulating large values of $n$ was expensive, but the counterexamples were only to be found at large values of $n$.
+%The most important use of mechanized logical deduction in this effort was its role as an ``implacable skeptic" that insisted we provide justification and compelled us to clarify the intricate details of the SRS structure and behavior.
 {\bf not sure if we should keep this paragraph after adding the previous one, maybe should be combined with the previous}
@@
 This effort targeted the validity of the SRS as an accurate and fast enough method by analyzing its behavior (a) in the boundaries of the threshold limits and (b) in expected types of beam losses, e.g., fast and steady losses.
-Its results showed that the current implementation of the SRS, given in Figure~\ref{fig:srs}, satisfies the specification, that is the running sums have a maximum 20\% relative error ($|\mathsf{RS}\;D\;n\;t - \mathsf{true\_sum}\;D\;n\;t| < 20\%(\mathsf{true\_sum}))$.
+Its results showed that the current implementation of the SRS, given in Figure~\ref{fig:srs}, satisfies the specification, that is the running sums have a maximum 20\% relative error ($|\mathsf{RS}\;D\;n\;t - \mathsf{true\_sum}\;D\;n\;t| < 20\%\times(\mathsf{true\_sum}))$.
 
-On the contrary, the results in this paper apply to a larger set of input streams, not just a limited number of sample inputs considered at CERN.
-However, the error bound proved in Theorem~\ref{max_error} is not tight enough to satisfy the maximum 20\% relative error.
-One of the potential reasons for this problem is the lack of knowledge to characterize the input stream to the SRS.
-Due to the nature of this problem, such a characterization of input is not trivial.
-While testing and simulation of the SRS on a limited set of input streams offer a level of confidence that the behavior of SRS satisfies this particular specification, constructing a formal proof for all possible behaviors requires a better understanding of the characteristics of the input stream.
+By contrast, the results in this paper apply to all possible input streams, not just the sample inputs considered at CERN.
+However, the bound proved in Theorem~\ref{max_error} is not tight enough to satisfy the maximum 20\% relative error bound.
+One of the potential reasons for this problem is not knowing how to characterize the input stream to the SRS.
+Due to the physical nature of this problem, such a characterization of the input stream is not trivial.
+While testing and simulation of the SRS on a limited set of input streams offers a level of confidence that the behavior of SRS satisfies this particular specification, constructing a formal proof for all possible behaviors requires a better understanding of the characteristics of the input stream.
 One future research direction is to investigate whether constructing such a formal proof is feasible for the SRS component of BLMS.

### 2011-04-01 fmics/paper.tex
merge dropbox changes to paper.tex

@@
 
-In this paper, we describe a formal verification approach, based on logical deduction using HOL4 theorem prover~\cite{HOL4,DBLP:conf/tphol/SlindN08}, to analyzing the SRS implementation.
+In this paper, we describe a formal verification approach, based on logical deduction using HOL4 theorem prover~\cite{HOL4,DBLP:conf/tphol/SlindN08}, to analyze the SRS implementation.
 Our high level proof strategy takes advantage of the regular structure of the SRS, which consists of multiple layers of shift registers and some simple arithmetic hardware.
@@
 HOL4 is an \emph{interactive} theorem prover: the user provides the high level proof strategy by composing functions that automate common chains of logical deduction.
-The work described here could have been done using other systems such as HOL Light~\cite{HOLLight,DBLP:conf/tphol/Harrison09a}, Isabelle/HOL~\cite{Isabelle}, ProofPower~\cite{ProofPower}, PVS~\cite{PVS,DBLP:conf/tphol/OwreS08} or Coq~\cite{Coq}.
+The work described here could have been done using other systems such as HOL Light~\cite{HOLLight}, Isabelle/HOL~\cite{Isabelle}, ProofPower~\cite{ProofPower}, PVS~\cite{PVS,DBLP:conf/tphol/OwreS08} or Coq~\cite{Coq}.
 The first three use essentially the same higher-order logic as HOL4, whilst PVS and Coq support more powerful logics.
@@
 \begin{enumerate}
-\item The shift register of each slice (except the first) is updated once the shift register of its previous slice is completely updated.
+\item The shift register of each slice (except the first) is updated once the shift register of its previous\footnote{Here, previous slice refers to the slice whose output is used as input to this slice.} slice is completely updated, that is, it flushes out all of the values entered its shift register since the previous update.
 \item The integration window of a given shift register in a slice (except the first slice) can be decomposed into a sequence of non-overlapping consecutive segments, $S_1, S_2, \ldots, S_w$ (where $w$ is the width of the shift register) each of which is equal to the size of the integration window of the shift register of the previous slice.
@@
 While the above paragraph gives the appearance of a straightforward argument for the correctness of the SRS (corresponding roughly to Theorem~\ref{output_input_at_update_times} in Section~\ref{subsec:theorems}), in fact the argument involves consideration of many details that arise from the formal model of the SRS presented in the next section.
-Using a software theorem prover enables us to keep track of these details without losing sight of the overall goal.
-{\bf one possibility is to move the paragraph about sources of complexities currently in section 5 to here.}
+Using a theorem proving tool enables us to keep track of these details without losing sight of the overall goal.
 
@@
 Therefore\footnote{For presentation convenience, here, we do not provide the details regarding the case when $t$ is small, for example $t<m\times\mathsf{delay}\;n$.
-For complete theorem statements and the proof script, please refer to \url{https://github.com/xrchz/CERN-LHC-BLMTC-SRS/blob/master/hol/srsScript.sml}.},
+For complete theorem statements and the proof script, please refer to {\scriptsize \url{https://github.com/xrchz/CERN-LHC-BLMTC-SRS/blob/master/hol/srsScript.sml}.}},
 \begin{equation*}
@@
 \begin{proof}
-By applying Theorem~\ref{output_eq_exact}, we are left with an equation involving $\mathsf{exact}$ but not $\mathsf{output}$, specifically between $\mathsf{exact}\;D\;n\;x\;t$ and $\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)$.
+By applying Theorem~\ref{output_eq_exact}, we are left with an equation involving the function $\mathsf{exact}$ but not $\mathsf{output}$, specifically between $\mathsf{exact}\;D\;n\;x\;t$ and $\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)$.
 Our assumption on $t$ ensures that we are never in the $0$ case of either Theorem~\ref{output_eq_exact} or of the definition of $\mathsf{exact}$.
@@
 
-We described a case study in which the HOL4 theorem prover is used to formally verify properties of the SRS structure within the LHC machine protection system.
+We described a case study in which the HOL4 theorem prover is used to verify properties of the SRS structure within the LHC machine protection system.
 In this case study, we built a parameterized model of the SRS structure and showed that its behavior is correct with respect to its specification.
-It is likely that the configuration of the slices and shift registers will need to change in the future as the understanding of the LHC and its weaknesses increases to accommodate more targeted protections.
+It is likely that the configuration of the slices and shift registers will need to change in the future as the understanding of the LHC and its possible weaknesses increases to accommodate more targeted protections.
 Thus, the parameterization in our model was crucial to make it applicable to the future upgrades.
@@
 (c) arithmetic relationships that are not always intuitively apparent at first glance.
-{\bf Ramana can you add an example here?}
 While each of these sources of complexity is manageable on its own, reasoning about the correctness of the SRS is matter of grappling with all three sources of complexity at once.
@@
 %The most important use of mechanized logical deduction in this effort was its role as an ``implacable skeptic" that insisted we provide justification and compelled us to clarify the intricate details of the SRS structure and behavior.
-{\bf not sure if we should keep this paragraph after adding the previous one, maybe should be combined with the previous}
 

### 2011-04-01 fmics/paper.tex
fix grammar error

@@
 
-In this paper, we describe a formal verification approach, based on logical deduction using HOL4 theorem prover~\cite{HOL4,DBLP:conf/tphol/SlindN08}, to analyze the SRS implementation.
+In this paper, we describe a formal verification approach, based on logical deduction using HOL4 theorem prover~\cite{HOL4,DBLP:conf/tphol/SlindN08}, to analyzing the SRS implementation.
 Our high level proof strategy takes advantage of the regular structure of the SRS, which consists of multiple layers of shift registers and some simple arithmetic hardware.

### 2011-04-01 fmics/paper.tex
reword fact 1

@@
 \begin{enumerate}
-\item The shift register of each slice (except the first) is updated once the shift register of its previous\footnote{Here, previous slice refers to the slice whose output is used as input to this slice.} slice is completely updated, that is, it flushes out all of the values entered its shift register since the previous update.
+\item The shift register of each slice (except the first) is updated once the shift register of its previous\footnote{Here, previous slice refers to the slice whose output is used as input to this slice.} slice is completely updated; an update means refreshing the value of every cell in the shift register.
 \item The integration window of a given shift register in a slice (except the first slice) can be decomposed into a sequence of non-overlapping consecutive segments, $S_1, S_2, \ldots, S_w$ (where $w$ is the width of the shift register) each of which is equal to the size of the integration window of the shift register of the previous slice.

### 2011-04-01 fmics/paper.tex
minor revisions in Section 5

@@
 One of our main challenges in this effort was building the formal model and formulating the correctness statements.
-There are three different sources of complexity that are inherent elements of understanding why the SRS structure implements its intended behavior:
+There are three different sources of complexity that are inherent in understanding why the SRS structure implements its intended behavior:
 (a) the structure of the SRS; although it features considerable regularity, it includes exceptions to this regularity that complicate reasoning, e.g., the input of one slice may not be the output of the immediately previous slice, but may be the first or second output of any previous slice or even the global input;
@@
 (c) arithmetic relationships that are not always intuitively apparent at first glance.
-While each of these sources of complexity is manageable on its own, reasoning about the correctness of the SRS is matter of grappling with all three sources of complexity at once.
+While each of these sources of complexity is manageable on its own, reasoning about the correctness of the SRS is a matter of grappling with all three sources of complexity at once.
 
@@
 
-Another future research direction is to refine our formal model to a lower level of abstraction and closer to the hardware model (e.g., representing the count values by words instead of natural number).
+Another future research direction is to refine our formal model to a lower level of abstraction, closer to the hardware level (e.g., by representing count values as finite words instead of natural numbers).
 We are also considering verifying a model extracted from the Hardware Description Language (HDL) used to synthesize the BLETC FPGA.

### 2011-04-01 fmics/paper.tex
Revise Fact 1 wording again

@@
 \begin{enumerate}
-\item The shift register of each slice (except the first) is updated once the shift register of its previous\footnote{Here, previous slice refers to the slice whose output is used as input to this slice.} slice is completely updated; an update means refreshing the value of every cell in the shift register.
+\item The shift register of each slice (except the first) is updated once the shift register of its previous\footnote{Here, previous slice refers to the slice whose output is used as input to this slice.} slice is completely updated, that is, when a count value has propagated down the full length of the previous slice's shift register.
 \item The integration window of a given shift register in a slice (except the first slice) can be decomposed into a sequence of non-overlapping consecutive segments, $S_1, S_2, \ldots, S_w$ (where $w$ is the width of the shift register) each of which is equal to the size of the integration window of the shift register of the previous slice.

### 2011-04-01 fmics/paper.tex
merge dropbox changes

--- 2011-04-01 fmics/paper.tex
@@
 In particular, our use of mechanized logical deduction was a highly iterative process that incrementally refined our understanding of (1) the implementation (2) the intended behavior and (3) the ``white board level" argument or explanation for why the implementation achieves the intended behaviors.
-The most important use of HOL4  was its role as an ``implacable skeptic" that compelled us to really understand the details~\cite{rushby}.
+The most important use of HOL4  was its role as an ``implacable skeptic" that insisted we provide justification and compelled us to clarify the details~\cite{rushby}.
 
@@
 Given the tolerance acceptable for quench prevention, the quench threshold versus loss duration is approximated by the minimum number of sliding integration windows (called \emph{Running Sums}) fulfilling the tolerance.
-In order to achieve the required \emph{dynamic range} (domain of variation of losses), the detectors use both Current-to-Frequency converter and Analogue to Digital convertor circuitries.
-The data processing block merges these two types of data coming from a detector so as to send a single value, referred to as a \emph{count}, to the SRS block.
+In order to achieve the required \emph{dynamic range} (domain of variation of losses), the detectors use both Current-to-Frequency converter and Analogue-to-Digital converter circuitries.
+The data combine block merges these two types of data coming from a detector so as to send a single value, referred to as a \emph{count}, to the SRS block.
 The implementation of the SRS block is described in Section~\ref{sec-SRS}.
@@
 HOL4 is an \emph{interactive} theorem prover: the user provides the high level proof strategy by composing functions that automate common chains of logical deduction.
-The work described here could have been done using other systems such as HOL Light~\cite{HOLLight}, Isabelle/HOL~\cite{Isabelle}, ProofPower~\cite{ProofPower}, PVS~\cite{PVS,DBLP:conf/tphol/OwreS08} or Coq~\cite{Coq}.
+The work described here could have been done using other systems such as HOL Light~\cite{HOLLight}, Isabelle/HOL~\cite{Isabelle}, ProofPower~\cite{ProofPower}, PVS~\cite{PVS} or Coq~\cite{Coq}.
 The first three use essentially the same higher-order logic as HOL4, whilst PVS and Coq support more powerful logics.
@@
 \begin{enumerate}
-\item The shift register of each slice (except the first) is updated once the shift register of its previous\footnote{Here, previous slice refers to the slice whose output is used as input to this slice.} slice is completely updated; an update means refreshing the value of every cell in the shift register.
-\item The integration window of a given shift register in a slice (except the first slice) can be decomposed into a sequence of non-overlapping consecutive segments, $S_1, S_2, \ldots, S_w$ (where $w$ is the width of the shift register) each of which is equal to the size of the integration window of the shift register of the previous slice.
+\item The shift register of each slice (except the first) is updated once the shift register of its previous\footnote{Here, the phrase ``previous slice" refers to the slice whose output is used as input to this slice.} slice is completely updated, that is, when a count value has propagated down the full length of the previous slice's shift register.
+\item The integration window of a given shift register in a slice (except the first slice) can be decomposed into a sequence of non-overlapping segments, $S_1, S_2, \ldots, S_w$ (where $w$ is the width of the shift register) each of which is equal to the size of the integration window of the shift register of the previous slice.
 \item After a period of initialization, the values stored in the shift register of a given slice (except the first) are the $w$ outputs of the previous slice, where $w$ is the width of the shift register of this slice.
@@
 Our model is both a simplification and a generalization of the actual structure of SRS in BLMS in the following sense.
-We model an infinite number of slices (rather than six slices), each with an infinite number of shift register cells and taps (rather than fixed width shift registers and only two taps), simply by letting indices range over the natural numbers without explicitly giving limits.
-This parameterization makes the model more likely to be applicable to future versions of the system, but also fits more naturally into HOL than would a finite model.
+We model an unbounded number of slices (rather than six slices), each with an unbounded number of shift register cells and taps (rather than fixed width shift registers and only two taps), simply by letting indices range over the natural numbers without explicitly giving limits.
+This parameterization makes the model more likely to be applicable to future versions of the system, but also fits more naturally into HOL than would a bounded model.
 We defined our formal model to be at a level of abstraction above the details of circuity that implements basic arithmetic operations.
 We use natural numbers throughout, rather than, for example, finite words that would more accurately model count values.
-This level of abstraction is sufficient to prove our desired correctness properties of SRS.
+This level of abstraction is sufficient to answer the questions that originally motivated this work.
 A separate verification effort can focus on showing that a more realistic model of hardware circuitry accurately implements natural number arithmetic.
@@
 \end{equation*}
-By induction on $t$, we can show that $\mathsf{output}\;D\;n\;x\;t$ is a sum of values of consecutive shift register cells
+By induction on $t$, we can show that $\mathsf{output}\;D\;n\;x\;t$ is a sum of values of consecutive shift register cells,
 \begin{equation*}
@@
 There are three different sources of complexity that are inherent in understanding why the SRS structure implements its intended behavior:
-(a) the structure of the SRS; although it features considerable regularity, it includes exceptions to this regularity that complicate reasoning, e.g., the input of one slice may not be the output of the immediately previous slice, but may be the first or second output of any previous slice or even the global input;
+(a) the structure of the SRS; although it features considerable regularity, it includes exceptions to this regularity that complicate reasoning, e.g., the input of one slice may not be the output of the immediately previous slice, but may be the first or second output of any earlier slice or even the global input;
 (b) the non-trivial timing relationships between different elements of the SRS; e.g., the frequency of updates to the contents of a shift register depends on the position of the shift register in the layered structure of the SRS;

### 2011-04-01 fmics/paper.tex
minor revisions throughout

@@
 Our confidence in the SRS design as a result of this effort ultimately rests upon our deep understanding of why the design is correct rather than the fact that we obtained ``Theorem Proved" as the final output of a software tool.
-In particular, our use of mechanized logical deduction was a highly iterative process that incrementally refined our understanding of (1) the implementation (2) the intended behavior and (3) the ``white board level" argument or explanation for why the implementation achieves the intended behaviors.
+In particular, our use of mechanized logical deduction was a highly iterative process that incrementally refined our understanding of (1) the implementation (2) the intended behavior and (3) the ``whiteboard-level" argument or explanation for why the implementation achieves the intended behaviors.
 The most important use of HOL4  was its role as an ``implacable skeptic" that insisted we provide justification and compelled us to clarify the details~\cite{rushby}.
@@
 \emph{(b) Data Processing}: Whether or not a quench results from particle loss depends on the loss duration and the beam energy.
-Given the tolerance acceptable for quench prevention, the quench threshold versus loss duration is approximated by the minimum number of sliding integration windows (called \emph{Running Sums}) fulfilling the tolerance.
+Given the tolerance acceptable for quench prevention, the quench threshold versus loss duration is approximated by the minimum number of sliding integration windows (called \emph{running sums}) fulfilling the tolerance.
 In order to achieve the required \emph{dynamic range} (domain of variation of losses), the detectors use both Current-to-Frequency converter and Analogue-to-Digital converter circuitries.
-The data combine block merges these two types of data coming from a detector so as to send a single value, referred to as a \emph{count}, to the SRS block.
+The Data Combine block merges these two types of data coming from a detector so as to send a single value, referred to as a \emph{count}, to the SRS block.
 The implementation of the SRS block is described in Section~\ref{sec-SRS}.
@@
 
-Prior to defining a formal model of the SRS and proving theorems about it, we developed an informal ``white board-level'' argument for why the SRS implements its intended behavior.
+Prior to defining a formal model of the SRS and proving theorems about it, we developed an informal ``whiteboard-level'' argument for why the SRS implements its intended behavior.
 The essence of this argument relies on four facts which can be established from the structure of the SRS and some details about the timing relationship between layers:
@@
 \(\mathsf{output}\;D\;n\;x\;t\)&The output at tap $x$ of slice $n$, at time $t$, given input stream $D$.\\
-\(\mathsf{RS}\;D\;n\;t\)&The $n$-th running sum at time $t$, given input stream $D$.\\
+\(\mathsf{RS}\;D\;n\;t\)&The value of running sum $n$ at time $t$, given input stream $D$.\\
 \(\mathsf{update\_time}\;n\;t\)&A boolean indicating whether $t$ is an update time for slice $n$.
@@
 %Some text description for Figure 7, we can remove it later on
-As explained in Section~\ref{sec-SRS}, the delay of each slice is equal to the delay of the slice it receives its input from multiplied by the number of the elements in the input slice used for the sum.
+As explained in Section~\ref{sec-SRS}, the delay of each slice is equal to the delay of the slice it receives its input from multiplied by the number of elements in the input slice used for the sum.
 For example in Figure~\ref{fig:srs}, $\mathsf{delay}\;4 = \mathsf{delay}\;3 \times \mathsf{tap}\;3\;0 = 2 \times 32$.
@@
 For example, $\mathsf{delay\_sum}\;4 = \mathsf{delay}\;3 + \mathsf{delay}\;1 + \mathsf{delay}\;0 = 2 +1 +1 = 4$.
-The function $\mathsf{last\_update}\;n\;t$ specifies the time that the $n$-th slice last updated, and $\mathsf{exact}\;D\;n\;x\;t$ computes the exact sum of consecutive input counts, without delay, that $\mathsf{output}\;D\;n\;x\;t$ is supposed to approximate.
+The function $\mathsf{last\_update}\;n\;t$ returns the latest time not after $t$ at which slice $n$ updates, and $\mathsf{exact}\;D\;n\;x\;t$ computes the exact sum of consecutive input counts, without delay, that $\mathsf{output}\;D\;n\;x\;t$ is supposed to approximate.
 
@@
 When updated, the values in its cells are shifted one cell.
-Therefore\footnote{For presentation convenience, here, we do not provide the details regarding the case when $t$ is small, for example $t<m\times\mathsf{delay}\;n$.
+Therefore\footnote{For presentational convenience, here, we do not provide details for the case when $t$ is small (e.g., when $t<m\times\mathsf{delay}\;n$).
 For complete theorem statements and the proof script, please refer to {\scriptsize \url{https://github.com/xrchz/CERN-LHC-BLMTC-SRS/blob/master/hol/srsScript.sml}.}},
@@
 
-We formulate the difference between the $\mathsf{RS}$ values and their corresponding $\mathsf{true\_sum}$s, caused by the delay, as $\mathsf{error}$.
-To meet the tolerance acceptable for quench prevention, the SRS specification requires a bound on this $\mathsf{error}$.
-Without restricting the input stream, the $\mathsf{error}$ is unbounded\footnote{For any bound, we can construct an input stream that causes the bound to be exceeded by, for example, having a sequence of zeroes followed by a sequence of count values much higher than the bound.}.
+We call the difference between the $\mathsf{RS}$ values and their corresponding $\mathsf{true\_sum}$s, caused by the delay, the \emph{error}.
+To meet the tolerance acceptable for quench prevention, the SRS specification requires a bound on this error.
+Without restricting the input stream, the error is unbounded\footnote{For any bound, we can construct an input stream that causes the bound to be exceeded by, for example, having a sequence of zeroes followed by a sequence of count values much higher than the bound.}.
 However, according to extensive experimental analysis, the beam loss over time follows some patterns.
-By formulating some characterization of the input stream as constraints on $D$, we can recover a bound on the $\mathsf{error}$.
+By formulating some characterization of the input stream as constraints on $D$, we can recover a bound on the error.
 One of these constraints is a maximum value for the difference between two consecutive input count values.
@@
 
-We described a case study in which the HOL4 theorem prover is used to verify properties of the SRS structure within the LHC machine protection system.
+We have described a case study in which we used HOL4 theorem prover to verify properties of the SRS structure within the LHC machine protection system.
 In this case study, we built a parameterized model of the SRS structure and showed that its behavior is correct with respect to its specification.
 It is likely that the configuration of the slices and shift registers will need to change in the future as the understanding of the LHC and its possible weaknesses increases to accommodate more targeted protections.
-Thus, the parameterization in our model was crucial to make it applicable to the future upgrades.
+Thus, the parameterization in our model is crucial to make it applicable to the future upgrades.
 
@@
 There are three different sources of complexity that are inherent in understanding why the SRS structure implements its intended behavior:
-(a) the structure of the SRS; although it features considerable regularity, it includes exceptions to this regularity that complicate reasoning, e.g., the input of one slice may not be the output of the immediately previous slice, but may be the first or second output of any earlier slice or even the global input;
+(a) the structure of the SRS: although it features considerable regularity, it includes exceptions to this regularity that complicate reasoning, e.g., the input of one slice may not be the output of the immediately previous slice, but may be the first or second output of any earlier slice or even the global input;
 (b) the non-trivial timing relationships between different elements of the SRS; e.g., the frequency of updates to the contents of a shift register depends on the position of the shift register in the layered structure of the SRS;
@@
 This effort targeted the validity of the SRS as an accurate and fast enough method by analyzing its behavior (a) in the boundaries of the threshold limits and (b) in expected types of beam losses, e.g., fast and steady losses.
-Its results showed that the current implementation of the SRS, given in Figure~\ref{fig:srs}, satisfies the specification, that is the running sums have a maximum 20\% relative error ($|\mathsf{RS}\;D\;n\;t - \mathsf{true\_sum}\;D\;n\;t| < 20\%\times(\mathsf{true\_sum}))$.
+Its results showed that the current implementation of the SRS, given in Figure~\ref{fig:srs}, satisfies the specification, that is, the running sums have a maximum 20\% relative error ($|\mathsf{RS}\;D\;n\;t - \mathsf{true\_sum}\;D\;n\;t| < 20\%\times(\mathsf{true\_sum}))$.
 

### 2011-05-26 fmics/paper.tex
Add Naghmeh's responses to minor comments from Dropbox

@@
 
+Our interest in using formal methods was originally motivated by questions about the SRS that arose in the course of an external technical audit of the BLMS performed by two of the co-authors, Ghafari and Joyce, and their colleagues.
 Compared to test-based methods, like simulation, formal methods not only offer much higher confidence in the correctness of a system's behavior, but also help improve our understanding of its specification.
@@
 
-\emph{(a) Receive, Check, and Compare (RCC)}: The RCC block receives data directly from the acquisition cards, and attempts to detect and correct erroneous transmissions by using Cyclic Redundancy Check and 8B/10B algorithms~\cite{CRC,8B10B}.
+\emph{(a) Receive, Check, and Compare (RCC)}: The RCC block receives data directly from the acquisition cards, and attempts to detect erroneous transmissions by using Cyclic Redundancy Check and 8B/10B algorithms~\cite{CRC,8B10B}.
 
@@
 There are 12 running sums calculated for each 16-detector channel allocated to a BLETC card.
-The beam energy information is scaled to 32 levels (0.45 to 7 TeV) and each processing module holds data only for those 16 connected detectors.
+There are 32 levels (0.45 to 7 TeV) of beam energy and each processing module holds data only for those 16 connected detectors.
+%The beam energy information is scaled to 32 levels (0.45 to 7 TeV) and each processing module holds data only for those 16 connected detectors.
 Thus, a total of 6,144 threshold values need to be held on each card.
@@
 %<What about ACL2?>
-While offering less ``push-button'' automation than other kinds of formal verification such as model-checking, machine-assisted theorem proving is appropriate for verifying the SRS, since it is unlikely that the correctness theorems, which need lemmas proved by manually guided mathematical induction, could be generated automatically.
+While offering less ``push-button'' automation than other kinds of formal verification such as model-checking, machine-assisted theorem proving using HOL4 is appropriate for verifying the SRS, 
+since it generates correctness theorems; in addition, it gives a way to very explicitly parameterize the model.
+%since it is unlikely that the correctness theorems, which need lemmas proved by manually guided mathematical induction, could be generated automatically.
 
@@
 $\;\mathsf{tap}\;2\;0=8-1\quad$ &  $\mathsf{tap}\;2\;x=16-1$\\
-$\;\dots\quad$ & $\mathsf{tap}\;6\;x=128-1\quad\text{(where each $x>0$)}$\\[1ex]
+$\;\mathsf{tap}\;3\;0=32-1\quad$ &  $\mathsf{tap}\;3\;x=128-1$\\
+$\;\mathsf{tap}\;4\;0=32-1\quad$ &  $\mathsf{tap}\;4\;x=256-1$\\
+$\;\mathsf{tap}\;5\;0=16-1\quad$ &  $\mathsf{tap}\;5\;x=64-1$\\
+$\;\mathsf{tap}\;6\;0=32-1\quad$ &  $\mathsf{tap}\;6\;x=128-1\quad\text{(where each $x>0$)}$\\
+$\;\dots\quad$ & $\mathsf{tap}\;n\;x=0\quad\text{(where $x\geq0$ and $n>6$)}$\\[1ex]
 
@@
 Since the outputs of a slice stay constant between update times for the slice, Theorem~\ref{output_input_at_update_times} suffices to characterize all outputs at all times.
-Thus, the SRS structure's outputs are equal to the exact running sums of the input counts over windows whose sizes depend on the the width of the slice's shift register and position of the taps.
+Thus, the SRS structure's outputs are equal to the exact running sums of the input counts over windows whose sizes depend on the width of the slice's shift register and position of the taps.
 However, the sums are delayed by the total delay across all previous slices (represented by $\mathsf{delay\_sum}$).
@@
 We are also considering verifying a model extracted from the Hardware Description Language (HDL) used to synthesize the BLETC FPGA.
+\\
+
+{\bf Acknowledgment:} The authors thank Mike Gordon at the University of Cambridge for his insightful and helpful feedback during the course of this work.
 

### 2011-05-26 fmics/paper.tex
A tight bound version of Theorem 3

@@
 t>(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay}\;n)+\mathsf{delay}\;n+\mathsf{delay\_sum}\;n\implies\\
-|\mathsf{output}\;D\;n\;x\;t-\mathsf{exact}\;D\;n\;x\;t|\leq(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay\;n})\times k\times(\mathsf{delay}\;n+\mathsf{delay\_sum}\;n)
+|\mathsf{output}\;D\;n\;x\;t-\mathsf{exact}\;D\;n\;x\;t|\leq(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay\;n})\times k\times(\mathsf{delay\_sum}\;n-1+t\operatorname{mod}\mathsf{delay}\;n)
 \end{multline*}
+Furthermore, for all $k$ there exists an input stream $D_k$ satisfying $|D_k\;(t'+1)-D_k\;t'|\leq k$ for all $t'$ and $|\mathsf{output}\;D_k\;n\;x\;t-\mathsf{exact}\;D_k\;n\;x\;t|=(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay\;n})\times k\times(\mathsf{delay\_sum}\;n-1+t\operatorname{mod}\mathsf{delay}\;n)$ for all $n>0$, $x$, and $t>(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay}\;n)+\mathsf{delay}\;n+\mathsf{delay\_sum}\;n$.
 \end{theorem}
 \begin{proof}
-By applying Theorem~\ref{output_eq_exact}, we are left with an equation involving the function $\mathsf{exact}$ but not $\mathsf{output}$, specifically between $\mathsf{exact}\;D\;n\;x\;t$ and $\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)$.
+By applying Theorem~\ref{output_eq_exact}, we are left with an inequation involving the function $\mathsf{exact}$ but not $\mathsf{output}$, specifically between $\mathsf{exact}\;D\;n\;x\;t$ and $\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)$.
 Our assumption on $t$ ensures that we are never in the $0$ case of either Theorem~\ref{output_eq_exact} or of the definition of $\mathsf{exact}$.
-The function $\mathsf{exact}$ is defined as a sum of consecutive values of $D$, so we simply need to bound the difference between one such sum and another earlier one of the same size.
+The function $\mathsf{exact}$ is defined as a sum of consecutive values of $D$; we need to bound the difference between one such sum and another earlier one of the same size.
 But if at each step the maximum difference is $k$, then the total difference is at most $k$ times the distance between the ends of the two sums, as required.
-(We use the fact that $\mathsf{last\_update}\;n\;t=t-t\operatorname{mod}\mathsf{delay}\;n$ and, in general, that $a\operatorname{mod}b<b$.) \qed
+(We use the fact that $\mathsf{last\_update}\;n\;t=t-t\operatorname{mod}\mathsf{delay}\;n$.)
+The input stream achieving this bound is given by $D_k\;t=k\times t$. \qed
 \end{proof}
+
+Theorem~\ref{max_error} gives a bound on the error of an SRS output at a given time in terms of the time step, the slice and tap numbers, and the assumed bound on the difference between consecutive counts.
+This bound is tight for the conditions we assume, namely that the time step is sufficiently high and that the difference between consecutive counts is bounded, in the sense that it is achievable by an input stream satisfying those conditions.
+To determine whether an error bound is safe, however, it turns out to be more useful to know the relative size of the error as a fraction of the true sum.
+We have not yet characterized the relationship between our error bound at a given time and the true sum at that time.
 

### 2011-05-28 fmics/paper.tex
Naghmeh's final changes from Dropbox

@@
 
+\begin{figure}[t]
+  \centering \includegraphics{rs.eps}
+   \caption{Block diagram showing how to produce and maintain a continuous running sum of arriving values.}
+  \label{fig:RS-basic}
+\end{figure}
+
 Figure~\ref{fig:BLETC} shows a block diagram of the processes on a BLETC FPGA.
@@
 \label{sec-SRS}
-
-\begin{figure}[t]
-  \centering \includegraphics{rs.eps}
-   \caption{Block diagram showing how to produce and maintain a continuous running sum of arriving values.}
-  \label{fig:RS-basic}
-\end{figure}
 
@@
 
-Under such a constraint, we have proved a preliminary result bounding the error of $\mathsf{output}\;D\;n\;x\;t$ compared to $\mathsf{exact}\;D\;n\;x\;t$.
+Under such a constraint, we have proved the follwoing result bounding the error of $\mathsf{output}\;D\;n\;x\;t$ compared to $\mathsf{exact}\;D\;n\;x\;t$.
 \begin{theorem}\label{max_error}
@@
 t>(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay}\;n)+\mathsf{delay}\;n+\mathsf{delay\_sum}\;n\implies\\
-|\mathsf{output}\;D\;n\;x\;t-\mathsf{exact}\;D\;n\;x\;t|\leq(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay\;n})\times k\times(\mathsf{delay\_sum}\;n-1+t\operatorname{mod}\mathsf{delay}\;n)
+|\mathsf{output}\;D\;n\;x\;t-\mathsf{exact}\;D\;n\;x\;t|\leq(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay\;n})\times k\times\\
+(\mathsf{delay\_sum}\;n-1+t\operatorname{mod}\mathsf{delay}\;n)
 \end{multline*}
-Furthermore, for all $k$ there exists an input stream $D_k$ satisfying $|D_k\;(t'+1)-D_k\;t'|\leq k$ for all $t'$ and $|\mathsf{output}\;D_k\;n\;x\;t-\mathsf{exact}\;D_k\;n\;x\;t|=(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay\;n})\times k\times(\mathsf{delay\_sum}\;n-1+t\operatorname{mod}\mathsf{delay}\;n)$ for all $n>0$, $x$, and $t>(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay}\;n)+\mathsf{delay}\;n+\mathsf{delay\_sum}\;n$.
+Furthermore, for all $k$ there exists an input stream $D_k$ satisfying $|D_k\;(t'+1)-D_k\;t'|\leq k$ for all $t'$ and 
+\begin{multline*}
+|\mathsf{output}\;D_k\;n\;x\;t-\mathsf{exact}\;D_k\;n\;x\;t|=(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay\;n})\times k\times\\
+(\mathsf{delay\_sum}\;n-1+t\operatorname{mod}\mathsf{delay}\;n)
+\end{multline*}
+
+for all $n>0$, $x$, and $t>(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay}\;n)+\mathsf{delay}\;n+\mathsf{delay\_sum}\;n$.
+
 \end{theorem}
 \begin{proof}
-By applying Theorem~\ref{output_eq_exact}, we are left with an inequation involving the function $\mathsf{exact}$ but not $\mathsf{output}$, specifically between $\mathsf{exact}\;D\;n\;x\;t$ and $\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)$.
+By applying Theorem~\ref{output_eq_exact}, we are left with an inequation involving the function $\mathsf{exact}$ 
+%but not $\mathsf{output}$
+specifically between $\mathsf{exact}\;D\;n\;x\;t$ and $\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)$.
 Our assumption on $t$ ensures that we are never in the $0$ case of either Theorem~\ref{output_eq_exact} or of the definition of $\mathsf{exact}$.
@@
 
-Theorem~\ref{max_error} gives a bound on the error of an SRS output at a given time in terms of the time step, the slice and tap numbers, and the assumed bound on the difference between consecutive counts.
+Theorem~\ref{max_error} gives a bound on the error of a running sum at a given time in terms of the time step, the slice and tap numbers, and the assumed bound on the difference between consecutive counts.
 This bound is tight for the conditions we assume, namely that the time step is sufficiently high and that the difference between consecutive counts is bounded, in the sense that it is achievable by an input stream satisfying those conditions.
-To determine whether an error bound is safe, however, it turns out to be more useful to know the relative size of the error as a fraction of the true sum.
+To determine whether an error bound is acceptable with respect to the specification of SRS, however, it turns out to be more useful to know the relative size of the error as a fraction of the true sum.
+According to the specification of SRS, the running sums should have a maximum 20\% relative error (($|\mathsf{RS}\;D\;n\;t - \mathsf{true\_sum}\;D\;n\;t| < 20\%\times(\mathsf{true\_sum}))$.
+%To determine whether an error bound is safe, however, it turns out to be more useful to know the relative size of the error as a fraction of the true sum.
 We have not yet characterized the relationship between our error bound at a given time and the true sum at that time.
@@
 This effort targeted the validity of the SRS as an accurate and fast enough method by analyzing its behavior (a) in the boundaries of the threshold limits and (b) in expected types of beam losses, e.g., fast and steady losses.
-Its results showed that the current implementation of the SRS, given in Figure~\ref{fig:srs}, satisfies the specification, that is, the running sums have a maximum 20\% relative error ($|\mathsf{RS}\;D\;n\;t - \mathsf{true\_sum}\;D\;n\;t| < 20\%\times(\mathsf{true\_sum}))$.
+Its results showed that the current implementation of the SRS, given in Figure~\ref{fig:srs}, satisfies the specification, that is, the running sums have a maximum 20\% relative error.
 
-By contrast, the results in this paper apply to all possible input streams, not just the sample inputs considered at CERN.
-However, the bound proved in Theorem~\ref{max_error} is not tight enough to satisfy the maximum 20\% relative error bound.
+By contrast, the results in this paper apply to all possible input streams, not just the sample inputs considered at CERN -- our main contribution, as stated in Theorem 2, showed that the behavior of the SRS is correct for all possible input streams.
+However, we do not know if the constraint on input stream given in Theorem 3 is sufficient to satisfy the maximum 20\% relative error bound.
+%However, the bound proved in Theorem~\ref{max_error} is not tight enough to satisfy the maximum 20\% relative error bound.
 One of the potential reasons for this problem is not knowing how to characterize the input stream to the SRS.

### 2011-05-28 fmics/paper.tex
fix typo

@@
 
-Under such a constraint, we have proved the follwoing result bounding the error of $\mathsf{output}\;D\;n\;x\;t$ compared to $\mathsf{exact}\;D\;n\;x\;t$.
+Under such a constraint, we have proved the following result bounding the error of $\mathsf{output}\;D\;n\;x\;t$ compared to $\mathsf{exact}\;D\;n\;x\;t$.
 \begin{theorem}\label{max_error}
